### 2015-03-23 manuscript/draft.tex
latex manuscript

\documentclass[preprint,authoryear]{elsarticle}

\usepackage[utf8]{inputenc}
\usepackage{xcolor}

\newcommand{\memo}[2]{\textcolor{#1}{#2}}
\newcommand{\xavi}[1]{\memo{orange}{xavi: #1\\}}
\newcommand{\enrico}[1]{\memo{blue}{enrico: #1\\}}

\journal{Evolution and Human Behavior}

\begin{document}

\begin{frontmatter}

\title{Allee Effect and Cultural Evolution}

\author[label1]{Enrico R. Crema\corref{cor1}}
\cortext[cor1]{Enrico R. Crema, enrico.crema@upf.edu}
\author[label2]{Xavier Rubio-Campillo}

\address[label1]{CaSEs - Complexity and Socio-Ecological Dynamics Research Group, Barcelona}
\address[label2]{BSC - Barcelona Supercomputing Center}
\begin{abstract}
Social learning often relies on the evaluation of the fitness of cultural traits through a comparative assessment of alternatives observed in a population. This process is shaped by the details of the learning strategy, as well as the context where the trait is displayed. The latter can modify the fitness of the cultural trait, ultimately changing the evolutionary fate of each adoption or loss. Here we explore a particular situation where the fitness of a trait is dependent to its frequency in the population.  We argue that many cultural traits pertaining subsistence activities exhibit both positive and negative frequency selection, derived by the beneficial effect of cooperation and the detrimental role played by interference and limited resources. A too small or too large population of individuals possessing the same variant will be subject to a decline in fitness, while an intermediate size will be optimal. We explore this core assumption, which closely resembles the “Allee effect” in ecology, using a modified Lotka-Volterra model of payoff-biased social learning, and show that this can generate a large variety of evolutionary dynamics observed in socially transmitted subsistence traits,  including episodes of reversion to suboptimal variants, long-term co-existence of multiple variants, and fixation to a single optimal or suboptimal strategy.
\end{abstract}

\begin{keyword}
Social Learning\sep Subsistence\sep Allee Effect\sep Cooperation\sep Carrying Capacity
\end{keyword}

\end{frontmatter}

\section{Introduction}

Major transitions in human history have been often marked by the fast spread of key innovations, which subsequently acted as catalysts for further changes in economy, society, and culture. Shifts in subsistence economy has often been regarded as a prime example of these transitions, and archaeologists and anthropologists have long focused on historical process behind these events (van der Veen 2010?). Key innovations such as the know-how of the domestication process \citep{barker2006}; the use of new subsistence tools\footnote{citation needed}; the adoption of processing techniques \citep{molleson1993}; the inclusion/exclusion and extensification/intensification of prey species\footnote{citation needed} are indeed often regarded as milestones in the human evolutionary history. Yet the archaeological record denies a simple progressive model of adoption. Instead, evidence showcase a variety of trajectories, including the long-term coexistence of multiple subsistence strategies, or even reversions to older cultural variants \citep{rowley2001}.

Formal models portraying changes in human subsistence have been mostly offered in human behavioral ecology (Smith and Winterhalder 1992, Bird and O'Connell 2006, Kennet and Winterhalder 2006), where several mathematical models inspired from evolutionary ecology offered precise prediction based on expected patterns under the assumption of optimality (Belovsky 1988)\footnote{citation needed}. While earliest models relied almost exclusively on this assumption, most recent works are increasing discussing how the knowledge of alternative subsistence strategies are acquired, and whether this process of learning can drive the evolutionary trajectory of a given population (Henrich and Boyd, 1998), partly as a synergy with the dual inheritance theory (Boyd and Richerson 1985).

The results of these works often exhibits counter-intuitive outcomes that cannot be predicted by standard assumptions of optimality. For instance, extreme reliance on social learning can promote information parasitism (Giraldeau et al 2002), leading to a decline in the proportion of asocial learners that in certain environmental context can have strongly detrimental effects at the population level (Whitehead and Richerson 2009). Other examples include instances where a reduced effective population size promotes an increase in random drift, and the ultimate loss of adaptive traits (Henrich 2004), or instances where neutral traits are jointly transmitted together with adaptive ones (Ackland et al 2007). These examples illustrate how theoretical models based on the dual inheritance theory can extend predictions of the standard optimal foraging theory, integrating dynamics that, at least at a first glance,  are regarded as counter-intuitive. These include episodes of reversions (e.g. \footnote{citation needed}), where a more recent optimal strategies are abandoned for older suboptimal variants; stable coexistence (e.g. \footnote{citation needed}), where optimal and suboptimal strategies co-exist within a population; and even the complete loss of optimal variants (e.g. Henrich 2004). The evolutionary processes leading to these different outcomes are undoubtedly contingent to local historical dynamics, nonetheless we argue that a simple model of social learning in combination with a positive and negative frequency dependent selection can lead all these equilibria.

\section{Payoff-biased Learning and Allee Effect}

A key feature that underlies many models of cultural transmission is that the social learner is often guided by some form of evaluation of characteristics associated with potential teachers or demonstrators. To put it simply, one can decide to copy traits possessed by individuals showing some signal that suggests higher payoff, success, or fitness (e.g. number of offspring, income, presence/absence of key traits, etc.). This learning strategy has been broadly referred to as directed social learning (Coussi-Korbel and Fragaszy 1995), model-biased transmission (Boyd and Richerson 1985, Henrich and Gil-White 2001), attraction model (Acerbi and Bentley 2014), and payoff-based transmission\footnote{citation needed}. The exact details of these model differ to some extent, as some might imply that the target trait of interest and the observed payoff-signal is only weakly correlated (e.g. prestige biased transmission), while other entails a more direct link between the two. In one sense one can argue that they are part of a broader category of social learning strategy (see also Crema and Lake, under review).

The degree of correlation between the payoff-signal and the target trait can also be viewed as a form of uncertainty or amount of information content offered by the signal. Previous studies (Schlag 1998,1999; Baldini 2013, Lake and Crema 2012) have explored how different learning strategies (e.g. copy-the-best, or copy-the-best-average, or copy-proportionally) show different dynamics, and are differently biased by this uncertainty. In some cases these might produce counter-intuitive outcomes. For instance, Baldini (2013) has shown that when the learning strategy is  copy-the-best suboptimal variants with higher payoff uncertainty and lower mean (equivalent to a lower correlation between signal payoff-signal and the target trait) can be favored over variants with higher mean and lower uncertainty. In other contexts, the uncertainty might be in the identification of the correct target trait, which could lead to the adoption of variants that have minimal or no correlation with the observed cue (cf. “evolutionary hitchhiking”,  Ackland et al 2007).

The uncertainty of the payoff signal is partly derived by the presence of interacting cultural and physical traits, as well as the environment where the trait is manifest. For example, a hunter might evaluate the performance of an arrowhead type (the target cultural trait) using as a cue the number of preys captured (the payoff signal). The correlation between the two is partly due to the efficiency of the arrowhead, but also by the bow, the prey species being capture and their availability,  the physical properties of the hunter, and chance. In many cases we should expect that the social teacher and learner share many of these interacting traits as well as the environment where these are displayed. In other words, cultural and ecological inheritance can generally reduce the uncertainty of the payoff-signal. There are however exceptions. One of these are contexts where the success of a variant is strongly correlated with the number of individuals possessing it within a population. This density dependence implies the adoption (or abandonment) of a trait by any individual can alter the expected fitness, and hence the payoff-signal, of all individuals within the population.

Several cultural traits are expected to show such density dependence. For example, behavior linked to the exploitation or consumption of limited resources are expected to exhibit a negative frequency dependence.  Novel optimal variants are likely to spread rapidly when the trait is rare, but once a threshold frequency is exceeded,  beneficial effect will start to decline,  the payoff will become smaller,  and eventually rarer suboptimal variants can be preferred (see Lake and Crema 2012 for an extensive analysis).  Negative dependency is however not the only form frequency dependence we should expect. In some cases, cultural variants that are strongly based on direct or indirect cooperation might require a minimum frequency to provide sufficient fitness, or more broadly show a positive correlation with the number of individuals possessing the trait. This positive frequency dependence, known in ecology as Allee Effect \citep{allee1958}, can be expected at different meta-population levels.  For example foraging groups are known to show higher performance when the number of individuals is not too small (Smith 1981), while certain activity such as anthropogenic fire (Bird et al 2013) or selective hunting (Dods 2002) might require a sufficient population density to promote sufficient niche enhancement and a consequent increase in payoff (Rowley-Conwy 2011). More\footnote{citation needed!} Language? Bettencourt Cities? \enrico{Ideally we need more example, not necessarily within the domain of subsistence activities. Any Allee Effect? Add Ostrom 2000 mention collective good problems here}.

Here we explore the combination of frequency dependent selection and payoff-based transmission with a modified version of the Lotka-Volterra model. We will demonstrate that degree by which a population relies on social learning can strongly affect the long-term equilibria of the system, with changes in the size of basins of different basins of attraction. Section 2 will illustrate our model, section 3 will explore three different scenarios, and section 4 will discuss our results and their broader implication.

[RESULTS HERE...]

\section{The Model}
We consider the evolutionary dynamics of two mutually exclusive subsistence strategies which fitness have both positive and negative frequency dependence. In particular we examine how two population of individuals, m and n, change over time if we allow for a social learning strategy guided by the difference in the fitness of the individuals. The model can theoretically be extended to more than two strategies, but here we decide to explore a two trait version for simplicity. The core model can be depicted with the following coupled difference equation:

\xavi{equation1}

where the population of each trait at time t+1 is the given by the number of individuals during the previous time-step (mt and nt), the number of offspring generated by each population (R(m,t) and R(n,t)) and the number of individuals shifting from one strategy to the other (C(m,n,t)).

The intrinsic growth rate of each population can be obtained from the following paired equation:

\xavi{equation2}

where r is the basic growth rate, A is the allee threshold, and K is the carrying capacity, with A<K.  The intrinsic growth rate will thus become negative when the population size is below A or above K, and will reach its maximum value Rmax at (K-A)/2, with Rmax equivalent to r(A-K)2/4AK.

Cultural transmission is portrayed here as a migration from the population with lower payoff to the population with higher payoff:

\xavi{equation3}

where delta defines the proportion of individuals shifting strategy. This is given by the following equation:

\xavi{equation4}

where z is the maximum transmission rate (i.e. the highest proportion of individuals shifting strategy), DELTA XX is the maximum between Rmax(m,t) and Rmax(n,t) , and b is a calibration parameter that measure the perception of the difference in the payoffs (i.e. small values of b determines a higher rate by which the theoretical maximum transmission rate is reached).

\section{Experiments}

We investigate different initial conditions of m and n for 6 different values of z (0,0.05,0.1,0.2,0.3, and 0.5). For each parameter combination we apply equations 1~4 for 50,000 iterations and record the final values of m and n, identifying the following equilibria

\enrico{We might need to use a rounding factor. And in fact these introduces to the problem of scale and difference between discrete and continuos  models.}
\enrico{We can analytical solve equations 1~4 to shows that the E1, E2 and E3 are indeed equilibria. Perhaps in an appendix? (although this is quite obvious...)}

We are aware that condition U can potentially become one of the other stable equilibria after the 50,000 simulations, hence we computed the Lyapunov exponent to define the stability of the system. We devised three core scenarios reflecting fixing b=2 and equating Rn to gRm, where g is given by the following equation:

\xavi{equation5}

so that when Rn=gRm,  Rmax(m,t) and Rmax(n,t) becomes equal. In all scenarios we considered Allee thresholds and carrying capacities as proportions by ensuring that max(Km,Kn)=1.

\enrico{Justify choice of arbitrary values}

\subsection{Scenario 1: Traits with identical payoffs (i.e. Am = An   and Km=Kn)}

The first scenario assumes two traits that are mathematically equivalent, so that  Am = An
, Km=Kn and rm=rn. Figure 1 shows the basin of attractions of each of the four equilibria for different values of z, setting r=0.005, .A=0.2 and K=1. When the z=0, i.e. when cultural transmission is disabled we have straight boundaries between the basins of E1, E2m, E2n, and E3 defined by Allee threshold. In other words when both one or both the population are below A we have some extinction events (E3 when mt=0<Am and nt=0<An, E2m when mt=0>Am and nt=0<An and E3n when mt=0<Am and nt=0>An). When z>0, the boundaries of these basins changes and with high z we see the emergence of unstable equilibria (U).

In all cases the E1 equilibria is maintained when $mt=0=nt=0 > Am=An$ as in this scenario differences arises only when m DIFFERENT OF n, and when the two populations are equally sized, C(m,n) is equal to 0 (see equation [3]).

\subsection{Scenario 2: Overlapping payoffs (i.e.  Am <  An  < Km < Kn)}

Am=0.2, An=0.5, Km=0.7,Kn=1

\subsection{Scenario 3: Internal overlap  (i.e.  Am <  An < Kn <  Km )}

An=0.2, Am=0.5, Km=0.7,Kn=1

\subsection{A note on competition}

The model described in equations [1]~[4] assumes that the resource pool of the two strategy are completely independent to each other. In reality many strategies have some degree of competition, whereby the frequency dependence is not exclusively internal. In other words, the frequency of one  trait might increase or decrease the fitness of the other. The following equation can substitute equation [2]:

\xavi{equation6}

where the four new parameters cAn, cKn, cAm, and cKm, measure respectively the negative (or positive) effect determined by the number of individuals of the “other” traits. When cA>0 the Allee threshold is essentially increased, while when cK>0 the carrying capacity is diminished as function of the other population., with larger values of c dictates a stronger influence of the other population.

One simple scenario is to assume that one strategy, say n, is dominant over its alternative m. We can  set cAn=0.1,  cKn=0.1,  cAm=-0.1, and  cKm=-0.1. Here negative values of c indicate a positive enhancement of one strategy over the other (in this case n gets benefit to the expense of m).

\section{Discussion}

\enrico{We need to mention that some behaviour might actually enhance the other community (or be detrimental). This can be modelled using a competition coefficient as in Jang 2013. I don't think we need to explore this but it is definitely worth mentioning!}

\enrico{Mutually exclusive trait, but does not necessarily imply HG vs FA mixture of strategy (say 50\%HG vs 50\%F) can be assumed to be a trait. Perhaps this needs to be mentioned earlier when the model is introduced.}

\section{Conclusion}

BlaBlaBla

\section{References}

\bibliographystyle{elsarticle-harv}
\bibliography{references}

\end{document}

### 2015-03-24 manuscript/draft.tex
added references

@@
 
-Formal models portraying changes in human subsistence have been mostly offered in human behavioral ecology (Smith and Winterhalder 1992, Bird and O'Connell 2006, Kennet and Winterhalder 2006), where several mathematical models inspired from evolutionary ecology offered precise prediction based on expected patterns under the assumption of optimality (Belovsky 1988)\footnote{citation needed}.
-While earliest models relied almost exclusively on this assumption, most recent works are increasing discussing how the knowledge of alternative subsistence strategies are acquired, and whether this process of learning can drive the evolutionary trajectory of a given population (Henrich and Boyd, 1998), partly as a synergy with the dual inheritance theory (Boyd and Richerson 1985).
+Formal models portraying changes in human subsistence have been mostly offered in human behavioral ecology \citep{smith1992,bird2006,kennett2006}, where several mathematical models inspired from evolutionary ecology offered precise prediction based on expected patterns under the assumption of optimality \citep{belovsky1988}.
+While earliest models relied almost exclusively on this assumption, most recent works are increasing discussing how the knowledge of alternative subsistence strategies are acquired, and whether this process of learning can drive the evolutionary trajectory of a given population \citep{henrich1998}, partly as a synergy with the dual inheritance theory \citep{boyd1985}.
 
 The results of these works often exhibits counter-intuitive outcomes that cannot be predicted by standard assumptions of optimality.
-For instance, extreme reliance on social learning can promote information parasitism (Giraldeau et al 2002), leading to a decline in the proportion of asocial learners that in certain environmental context can have strongly detrimental effects at the population level (Whitehead and Richerson 2009).
-Other examples include instances where a reduced effective population size promotes an increase in random drift, and the ultimate loss of adaptive traits (Henrich 2004), or instances where neutral traits are jointly transmitted together with adaptive ones (Ackland et al 2007).
+For instance, extreme reliance on social learning can promote information parasitism \citep{giraldeau2002}, leading to a decline in the proportion of asocial learners that in certain environmental context can have strongly detrimental effects at the population level \citep{whitehead2009}.
+Other examples include instances where a reduced effective population size promotes an increase in random drift, and the ultimate loss of adaptive traits \citep{henrich2004}, or instances where neutral traits are jointly transmitted together with adaptive ones \citep{ackland2007}.
 These examples illustrate how theoretical models based on the dual inheritance theory can extend predictions of the standard optimal foraging theory, integrating dynamics that, at least at a first glance, are regarded as counter-intuitive.
@@
 \footnote{citation needed}), where a more recent optimal strategies are abandoned for older suboptimal variants; stable coexistence (e.g.
-\footnote{citation needed}), where optimal and suboptimal strategies co-exist within a population; and even the complete loss of optimal variants (e.g.
-Henrich 2004).
+\footnote{citation needed}), where optimal and suboptimal strategies co-exist within a population; and even the complete loss of optimal variants \citep[e.g.][]{henrich2004}.
 The evolutionary processes leading to these different outcomes are undoubtedly contingent to local historical dynamics, nonetheless we argue that a simple model of social learning in combination with a positive and negative frequency dependent selection can lead all these equilibria.
@@
 number of offspring, income, presence/absence of key traits, etc.).
-This learning strategy has been broadly referred to as directed social learning (Coussi-Korbel and Fragaszy 1995), model-biased transmission (Boyd and Richerson 1985, Henrich and Gil-White 2001), attraction model (Acerbi and Bentley 2014), and payoff-based transmission\footnote{citation needed}.
+This learning strategy has been broadly referred to as directed social learning \citep{coussi1995}, model-biased transmission \citep{boyd1985,henrich2001}, attraction model \citep{acerbi2014}, and payoff-based transmission\footnote{citation needed}.
 The exact details of these model differ to some extent, as some might imply that the target trait of interest and the observed payoff-signal is only weakly correlated (e.g.
 prestige biased transmission), while other entails a more direct link between the two.
-In one sense one can argue that they are part of a broader category of social learning strategy (see also Crema and Lake, under review).
+In one sense one can argue that they are part of a broader category of social learning strategy\footnote{citation needed - see also Crema and Lake, under review}.
 
 The degree of correlation between the payoff-signal and the target trait can also be viewed as a form of uncertainty or amount of information content offered by the signal.
-Previous studies (Schlag 1998,1999; Baldini 2013, Lake and Crema 2012) have explored how different learning strategies (e.g.
+Previous studies \citep{schlag1998,baldini2013,lake2012} have explored how different learning strategies (e.g.
 copy-the-best, or copy-the-best-average, or copy-proportionally) show different dynamics, and are differently biased by this uncertainty.
 In some cases these might produce counter-intuitive outcomes.
-For instance, Baldini (2013) has shown that when the learning strategy is  copy-the-best suboptimal variants with higher payoff uncertainty and lower mean (equivalent to a lower correlation between signal payoff-signal and the target trait) can be favored over variants with higher mean and lower uncertainty.
-In other contexts, the uncertainty might be in the identification of the correct target trait, which could lead to the adoption of variants that have minimal or no correlation with the observed cue (cf.
-“evolutionary hitchhiking”,  Ackland et al 2007).
+For instance, \citet{baldini2013} has shown that when the learning strategy is  copy-the-best suboptimal variants with higher payoff uncertainty and lower mean (equivalent to a lower correlation between signal payoff-signal and the target trait) can be favored over variants with higher mean and lower uncertainty.
+In other contexts, the uncertainty might be in the identification of the correct target trait, which could lead to the adoption of variants that have minimal or no correlation with the observed cue \citep[cf.
+"evolutionary hitchhiking"][]{ackland2007}.
 
@@
 This positive frequency dependence, known in ecology as Allee Effect \citep{allee1958}, can be expected at different meta-population levels.
-For example foraging groups are known to show higher performance when the number of individuals is not too small (Smith 1981), while certain activity such as anthropogenic fire (Bird et al 2013) or selective hunting (Dods 2002) might require a sufficient population density to promote sufficient niche enhancement and a consequent increase in payoff (Rowley-Conwy 2011).
+For example foraging groups are known to show higher performance when the number of individuals is not too small\footnote{There was a reference on Smith 1981...is this JM Smith 'evolution and theory of games'?
+This is 1982}, while certain activity such as anthropogenic fire \citep{bird2013} or selective hunting (Dods 2002) might require a sufficient population density to promote sufficient niche enhancement and a consequent increase in payoff \citep{rowley2011}.
 More\footnote{citation needed!} Language?
@@
 Any Allee Effect?
-Add Ostrom 2000 mention collective good problems here}.
+Add Ostrom 2000 mention collective good problems here\footnote{what is ostrom 2000?}}.
 
@@
 \enrico{We need to mention that some behaviour might actually enhance the other community (or be detrimental).
-This can be modelled using a competition coefficient as in Jang 2013.
+This can be modelled using a competition coefficient as in \citep{jang2013}.
 I don't think we need to explore this but it is definitely worth mentioning!}

### 2015-03-25 manuscript/draft.tex
majority of equations defined

@@
 \usepackage{xcolor}
+\usepackage{amsmath}
 
@@
 The model can theoretically be extended to more than two strategies, but here we decide to explore a two trait version for simplicity.
-The core model can be depicted with the following coupled difference equation:
+The core model can be depicted with the coupled difference equation \eqref{eq1}:
 
-\xavi{equation1}
+\begin{align}
+\label{eq1}
+\begin{cases}
+m_{t+1}& = m_t + m_t R_{m,t} + C_{m,n,t} \\
+n_{t+1}& = n_t + n_t R_{n,t} + C_{m,n,t}
+\end{cases}
+\end{align}
 
-where the population of each trait at time t+1 is the given by the number of individuals during the previous time-step (mt and nt), the number of offspring generated by each population (R(m,t) and R(n,t)) and the number of individuals shifting from one strategy to the other (C(m,n,t)).
+where $m_{t+1}$ and $n_{t+1}$ is the population of each trait at time $t+1$ is the given by the number of individuals during the previous time-step ($m_t$ and $n_t$), $R_{m,t}$ and $R_{n,t}$ is the number of offspring generated by each population and $C_{m,n,t}$ is the number of individuals shifting from one strategy to the other.
 
-The intrinsic growth rate of each population can be obtained from the following paired equation:
+The intrinsic growth rate of each population can be obtained from the paired equation \eqref{eq2}:
 
-\xavi{equation2}
+\begin{align}
+\label{eq2}
+\begin{cases}
+R_{m,t}& = r_m (\frac{m_t}{A_m}-1)(1-\frac{m_t}{K_m})\\
+R_{n,t}& = r_n (\frac{n_t}{A_n}-1)(1-\frac{n_t}{K_n})
+\end{cases}
+\end{align}
 
-where r is the basic growth rate, A is the allee threshold, and K is the carrying capacity, with A<K.
-The intrinsic growth rate will thus become negative when the population size is below A or above K, and will reach its maximum value Rmax at (K-A)/2, with Rmax equivalent to r(A-K)2/4AK.
+where $r$ is the basic growth rate, $A$ is the allee threshold, and $K$ is the carrying capacity, with $A < K$.
+The intrinsic growth rate will thus become negative when the population size is below $A$ or above $K$, and will reach its maximum value $R_{max} = \frac{(K-A)}{2}$, with $R_{max}$ equivalent to $\frac{r(A-K)2}{4AK}$.
+\xavi{I think this last sentence is confusing}
 
-Cultural transmission is portrayed here as a migration from the population with lower payoff to the population with higher payoff:
+Cultural transmission is portrayed here as a migration from the population with lower payoff to the population with higher payoff, as seen in \eqref{eq3}:
 
-\xavi{equation3}
+\begin{align}
+\label{eq3}
+C_{m,n,t} = 
+\begin{cases}
+0& \text{if } R_{m,t} = R_{n,t}\\
+\zeta(m_t+m_tR_{m,t})& \text{if } R_{m,t} > R_{n,t}\\
+-\zeta(n_t+n_tR_{n,t})& \text{if } R_{m,t} < R_{n,t}
+\end{cases}
+\end{align}
 
-where delta defines the proportion of individuals shifting strategy. This is given by the following equation:
+where $\zeta$ defines the proportion of individuals shifting strategy, and given by \eqref{eq4}:
 
-\xavi{equation4}
+\begin{align}
+\label{eq4}
+\zeta = 
+\begin{cases}
+z& \text{if }|R_{m,t}-R_{n,t}| > b\Delta\\
+z\frac{|R_{m,t}-R{n,t}|}{b\Delta}& \text{if }|R_{m,t}-R_{n,t}| \leq b\Delta\\
+\end{cases}
+\end{align}
 
-where z is the maximum transmission rate (i.e.
-the highest proportion of individuals shifting strategy), DELTA XX is the maximum between Rmax(m,t) and Rmax(n,t) , and b is a calibration parameter that measure the perception of the difference in the payoffs (i.e.
+where $\zeta$ is the maximum transmission rate (i.e.
+the highest proportion of individuals shifting strategy), $\Delta$ is the maximum between $R_{max(m,t)}$ and $R_{max(n,t)}$, and $b$ is a calibration parameter that measure the perception of the difference in the payoffs (i.e.
 small values of b determines a higher rate by which the theoretical maximum transmission rate is reached).
@@
 
-We investigate different initial conditions of m and n for 6 different values of z (0,0.05,0.1,0.2,0.3, and 0.5).
-For each parameter combination we apply equations 1~4 for 50,000 iterations and record the final values of m and n, identifying the following equilibria
+We investigate different initial conditions of $m$ and $n$ for 6 different values of z (0,0.05,0.1,0.2,0.3, and 0.5).
+For each parameter combination we apply equations (\ref{eq1},~\ref{eq4}) for 50,000 iterations and record the final values of $m$ and $n$, identifying the following equilibria.
 
@@
 
-\xavi{equation5}
+\xavi{equation5, lyapunov, where this does go?}
 
@@
 
-The first scenario assumes two traits that are mathematically equivalent, so that  Am = An
-, Km=Kn and rm=rn.
+The first scenario assumes two traits that are mathematically equivalent, so that $A_m = A_m, K_m=K_n and r_m=r_n$.
 Figure 1 shows the basin of attractions of each of the four equilibria for different values of z, setting r=0.005, .A=0.2 and K=1.
@@
 
-The model described in equations [1]~[4] assumes that the resource pool of the two strategy are completely independent to each other.
+The model described in equations  (\ref{eq1},~\ref{eq4}) assumes that the resource pool of the two strategy are completely independent to each other.
 In reality many strategies have some degree of competition, whereby the frequency dependence is not exclusively internal.

### 2015-03-27 src/basin.R
Added Function for counting different equilibria in the basin

@@ -100,7 +100,39 @@ basePlot<-function(...)
     }
 
 
+basinCount<-function(res)
+    {
+        #Identify Stable Attractor Node
+        Em=ExpStableNode(Kn,Km,cKn,cKm)$m
+        En=ExpStableNode(Kn,Km,cKn,cKm)$n
+        
+        #deal with floating point arithmethic problem#
+        res$m.final[which(as.logical(sapply(res$m.final,all.equal,target=Km,USE.NAMES=F,simplify=TRUE)))]=Km
+        res$n.final[which(as.logical(sapply(res$n.final,all.equal,target=Kn,USE.NAMES=F,simplify=TRUE)))]=Kn
+        res$m.final[which(as.logical(sapply(res$m.final,all.equal,target=Em,USE.NAMES=F,simplify=TRUE)))]=Em
+        res$n.final[which(as.logical(sapply(res$n.final,all.equal,target=En,USE.NAMES=F,simplify=TRUE)))]=En
+        res$m.final[which(as.logical(sapply(res$m.final,all.equal,target=0,USE.NAMES=F,simplify=TRUE)))]=0
+        res$n.final[which(as.logical(sapply(res$n.final,all.equal,target=0,USE.NAMES=F,simplify=TRUE)))]=0
 
+        #Define Classes
+        res$class=1 #Default Class (Unstable Equilibria)
+        # E1: Stable Coexistence
+        res$class[which(res$m.final==Em&res$n.final==En)]=2
+        # E2m: Dominance m
+        res$class[which(res$m.final==Km&res$n.final==0)]=3   
+        # E2n: Dominance n
+        res$class[which(res$m.final==0&res$n.final==Kn)]=4      
+        # E3: Extinction 
+        res$class[which(res$m.final==0&res$n.final==0)]=5
+
+        UE=sum(res$class==1)/nrow(res)
+        SE=sum(res$class==2)/nrow(res)
+        Sm=sum(res$class==3)/nrow(res)
+        Sn=sum(res$class==4)/nrow(res)
+        E=sum(res$class==5)/nrow(res)
+        
+        return(list(UE=UE,SE=E,Sm=Sm,Sn=Sn,E=E))
+    }
 
 
 

### 2015-03-27 src/basin.R
Minor Bug Fixes

@@ -1,8 +1,6 @@
 basin<-function(x.lim,y.lim,points,multicore=TRUE,r.n,r.m,Am,Km,An,Kn,z,timeSteps,b=2,cores=1,cAn,cAm,cKn,cKm,verbose=TRUE)
 {
-    require(utils)
-    require(foreach)
-    require(doParallel)
+
 
     x <- seq(from = x.lim[1], to = x.lim[2], length = points)
     y <- seq(from = y.lim[1], to = y.lim[2], length = points)
@@ -14,6 +12,9 @@ basin<-function(x.lim,y.lim,points,multicore=TRUE,r.n,r.m,Am,Km,An,Kn,z,timeStep
 
     if (multicore==TRUE)
         {
+            require(utils)
+            require(foreach)
+            require(doParallel)
             registerDoParallel(cores=cores)
             tmp=foreach(i=1:nrow(res), .combine=rbind) %dopar%
             {
@@ -131,7 +132,7 @@ basinCount<-function(res)
         Sn=sum(res$class==4)/nrow(res)
         E=sum(res$class==5)/nrow(res)
         
-        return(list(UE=UE,SE=E,Sm=Sm,Sn=Sn,E=E))
+        return(list(UE=UE,SE=SE,Sm=Sm,Sn=Sn,E=E))
     }
 
 

### 2015-03-27 LegionSubmit/send.sh
Legion Submission Codes

@@ -1,0 +1,46 @@
+#!/bin/bash -l
+
+# Batch script to run an array job on Legion with the upgraded
+# software stack under SGE.
+
+# 1. Force bash
+#$ -S /bin/bash
+
+# 2. Request ten minutes of wallclock time (format hours:minutes:seconds).
+#$ -l h_rt=1:30:00
+
+# 3. Request 1 gigabyte of RAM.
+#$ -l mem=4G
+
+# 4 Request TMPDIR space                                     
+#$ -l tmpfs=15G 
+
+# 5. Set up the job array.  In this instance we have requested 10000 tasks
+# numbered 1 to 10000.
+#$ -t 1-2601
+
+# 6. Set the name of the job.
+#$ -N allee
+
+# 7. Select the project that this job will run under.
+# Find <your_project_id> by running the command "groups"
+
+# 8. Set the working directory to somewhere in your scratch space.  This is
+# a necessary step with the upgraded software stack as compute nodes cannot
+# write to $HOME.
+# Replace "<your_UCL_id>" with your UCL user ID :)
+#$ -wd /home/tcrnerc/Scratch/models/allee/temp
+
+# 9. Run the application.
+cd $TMPDIR
+
+#export R_SESSION_TMPDIR=$TMPDIR
+module unload compilers/intel/11.1/072
+module unload mpi/qlogic/1.2.7/intel
+module unload mkl/10.2.5/035
+module load recommended/r
+
+# 10. Run the script. 
+Rscript  /home/tcrnerc/Scratch/models/allee/submit.R $SGE_TASK_ID
+# 11. Move the file
+mv *.csv $HOME/Scratch/output/allee/overlap

### 2015-03-27 LegionSubmit/submit.R

@@ -1,0 +1,34 @@
+
+
+place<-getwd()
+source("/home/tcrnerc/Scratch/models/allee/src/allee.R")
+source("/home/tcrnerc/Scratch/models/allee/src/basin.R")
+source("/home/tcrnerc/Scratch/models/allee/src/utility.R")
+setwd(place)
+
+lambda=seq(0,0.5,0.01)
+z=seq(0,0.5,0.01)
+param=expand.grid(lambda=lambda,z=z)
+Args<-as.numeric(commandArgs(TRUE))
+thisparam<-param[Args,]
+Am=0.1
+An=Am+thisparam$lambda
+Km=0.6
+Kn=Km+thisparam$lambda
+z=thisparam$z
+cAn=0
+cAm=0
+cKn=0
+cKm=0
+r.n=0.05
+r.m=EqualiseMaxGrowthRate(An=An,Am=Am,Kn=Kn,Km=Km)*r.n
+x.lim=c(0,1.2)
+y.lim=c(0,1.2)
+    
+res=basin(x.lim=x.lim,y.lim=x.lim,points=40,timeSteps=100000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=TRUE,multicore=FALSE)
+
+resCount=basinCount(res)
+finalres=cbind(thisparam,as.data.frame(resCount))
+
+name<-paste("./res",Args,".csv",sep="")
+write.csv(finalres,file=name)

### 2015-05-04 manuscript/draft.tex
Updated Abstract

@@
 
-\journal{Evolution and Human Behavior}
+\journal{Theoretical Population Biology}
 
@@
 
-\author[label1]{Enrico R. Crema\corref{cor1}}
-\cortext[cor1]{Enrico R. Crema, enrico.crema@upf.edu}
-\author[label2]{Xavier Rubio-Campillo}
+\author[label1,label2]{Enrico R. Crema\corref{cor1}}
+\cortext[cor1]{Corresponding Author: enrico.crema@upf.edu}
+\author[label3]{Xavier Rubio-Campillo}
 
 \address[label1]{CaSEs - Complexity and Socio-Ecological Dynamics Research Group, Barcelona}
-\address[label2]{BSC - Barcelona Supercomputing Center}
+\address[label2]{UCL Institute of Archaeology}
+\address[label3]{BSC - Barcelona Supercomputing Center}
+
 \begin{abstract}
-Social learning often relies on the evaluation of the fitness of cultural traits through a comparative assessment of alternatives observed in a population.
-This process is shaped by the details of the learning strategy, as well as the context where the trait is displayed.
-The latter can modify the fitness of the cultural trait, ultimately changing the evolutionary fate of each adoption or loss.
-Here we explore a particular situation where the fitness of a trait is dependent to its frequency in the population.
-We argue that many cultural traits pertaining subsistence activities exhibit both positive and negative frequency selection, derived by the beneficial effect of cooperation and the detrimental role played by interference and limited resources.
-A too small or too large population of individuals possessing the same variant will be subject to a decline in fitness, while an intermediate size will be optimal.
-We explore this core assumption, which closely resembles the “Allee effect” in ecology, using a modified Lotka-Volterra model of payoff-biased social learning, and show that this can generate a large variety of evolutionary dynamics observed in socially transmitted subsistence traits,  including episodes of reversion to suboptimal variants, long-term co-existence of multiple variants, and fixation to a single optimal or suboptimal strategy.
+Several forms of social learning rules are guided by the amount of payoff received by social teachers adopting specific cultural variants.
+The rationale, is that such payoff will be correlated to the beneficial or detrimental properties of a given variant.
+Here we assume that the payoff directly affects the reproductive fitness of individuals, and that it is determined by the absolute frequency of individuals possessing the same cultural variant.
+More specifically we consider a  form of frequency dependence where a too small or too large population of individuals possessing the same variant will be subject to a decline in fitness, while an intermediate size is optimal.
+This non-linear relationship, known as Allee effect in ecology, portrays a large number of cultural variants, in particular subsistence strategies that are enhanced  by the effect of cooperation but are constrained by interference and limited resources.
+This paper explores the evolutionary dynamics of a two trait model where each variant exhibit Allee effect.
+We used a modified Lotka-Volterra model of payoff-biased social learning, and show that that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion to suboptimal variants, long-term co-existence of multiple variants, and fixation to a single optimal or suboptimal strategy.
 \end{abstract}

### 2015-05-06 manuscript/draft.tex
Updated Manuscript

@@
 \title{Allee Effect and Cultural Evolution}
+%\title{The evolution of payoff-biased learning under Allee Effect} %Perhaps a
 
@@
 The rationale, is that such payoff will be correlated to the beneficial or detrimental properties of a given variant.
-Here we assume that the payoff directly affects the reproductive fitness of individuals, and that it is determined by the absolute frequency of individuals possessing the same cultural variant.
-More specifically we consider a  form of frequency dependence where a too small or too large population of individuals possessing the same variant will be subject to a decline in fitness, while an intermediate size is optimal.
+Here we assume that the payoff directly affects the reproductive fitness of individuals, and that it is determined by the density of individuals possessing the same cultural variant.
+More specifically we consider a  form of density-dependence where a too small or too large population of individuals possessing the same variant will be subject to a decline in fitness, while an intermediate size is optimal.
 This non-linear relationship, known as Allee effect in ecology, portrays a large number of cultural variants, in particular subsistence strategies that are enhanced  by the effect of cooperation but are constrained by interference and limited resources.
@@
 
-Major transitions in human history have been often marked by the fast spread of key innovations, which subsequently acted as catalysts for further changes in economy, society, and culture.
-Shifts in subsistence economy has often been regarded as a prime example of these transitions, and archaeologists and anthropologists have long focused on historical process behind these events (van der Veen 2010?).
-Key innovations such as the know-how of the domestication process \citep{barker2006}; the use of new subsistence tools\footnote{citation needed}; the adoption of processing techniques \citep{molleson1993}; the inclusion/exclusion and extensification/intensification of prey species\footnote{citation needed} are indeed often regarded as milestones in the human evolutionary history.
-Yet the archaeological record denies a simple progressive model of adoption.
-Instead, evidence showcase a variety of trajectories, including the long-term coexistence of multiple subsistence strategies, or even reversions to older cultural variants \citep{rowley2001}.
+A key element in all social learning strategies is the availability of some heuristics and learning bias that guides the selection of a given variant over another.
+This information can be based on some properties of the variant itself (\emph{content-bias}) or derived from contextual cues (\emph{context-bias}), such as the rarity or the commonness or the rareness of a trait, or some features associated with the model (i.e.
+the social teacher) such as prestige, success, or the similarity to the learner \citep{henrich_mcelreath2003}.
+An expected consequence of \emph{context-bias} is the possibility of a feedback process, whereby the attractiveness of a given might change as a consequence of the learning strategy \citep{kendal_etal_2009}.
+For example a strong anti-conformist bias might initially favour rare traits, but as they spread into a population their selective advantage will decline as the result of a negative feedback.
+In a similar, but in a opposite direction, both conformist-bias and homophily-guided learning strategies would exercise a positive feedback loop.
+%citation needed?
 
-Formal models portraying changes in human subsistence have been mostly offered in human behavioral ecology \citep{smith1992,bird2006,kennett2006}, where several mathematical models inspired from evolutionary ecology offered precise prediction based on expected patterns under the assumption of optimality \citep{belovsky1988}.
-While earliest models relied almost exclusively on this assumption, most recent works are increasing discussing how the knowledge of alternative subsistence strategies are acquired, and whether this process of learning can drive the evolutionary trajectory of a given population \citep{henrich1998}, partly as a synergy with the dual inheritance theory \citep{boyd1985}.
+This paper examines a group of learning strategies generally referred to as \emph{payoff-based transmission} \citep{schlag1998,kendal_etal_2009,lake_and_crema_2012,baldini2013,kandler_and_laland_2013,crema_lake_inpress}.
+The core assumption shared by these models is that the probability of copying a given cultural variant is positively correlated with some payoff signal expressed by individuals possessing such a variant.
+Several variants of this learning strategies have been proposed and explored, including: "imitate the best", where the social learner identifies and copies the trait possessed by the individual with the highest payoff; "compare means" where the social learner compare the average payoff of different strategies and chooses the highest variant; and the "proportional imitation rule", where the probability of adopting a variant is proportional to the difference in payoff between the demonstrator and the learner \citep{schlag1998,baldini2013}.
 
-The results of these works often exhibits counter-intuitive outcomes that cannot be predicted by standard assumptions of optimality.
-For instance, extreme reliance on social learning can promote information parasitism \citep{giraldeau2002}, leading to a decline in the proportion of asocial learners that in certain environmental context can have strongly detrimental effects at the population level \citep{whitehead2009}.
-Other examples include instances where a reduced effective population size promotes an increase in random drift, and the ultimate loss of adaptive traits \citep{henrich2004}, or instances where neutral traits are jointly transmitted together with adaptive ones \citep{ackland2007}.
-These examples illustrate how theoretical models based on the dual inheritance theory can extend predictions of the standard optimal foraging theory, integrating dynamics that, at least at a first glance, are regarded as counter-intuitive.
-These include episodes of reversions (e.g.
-\footnote{citation needed}), where a more recent optimal strategies are abandoned for older suboptimal variants; stable coexistence (e.g.
-\footnote{citation needed}), where optimal and suboptimal strategies co-exist within a population; and even the complete loss of optimal variants \citep[e.g.][]{henrich2004}.
-The evolutionary processes leading to these different outcomes are undoubtedly contingent to local historical dynamics, nonetheless we argue that a simple model of social learning in combination with a positive and negative frequency dependent selection can lead all these equilibria.
+Whilst most studies focused on how the difference in these imitation rules can drive the dynamics of the system, less attention has been paid on the payoff signal.
+The exact nature of the signal is most likely some indirect proxy that is assumed, by the social learner, to be linked with the target cultural trait.
+Thus one might be guided by a variety of cues (e.g.
+number of offpsring, income, presence/absence of other traits associated with success or prestige, etc.) that are differently correlated with the target trait resulting into different degrees of uncertainty in the signal.
+The magnitude of this uncertainty can have strong impact to efficiency of specific learning rules.
+For example, Crema and Lake (In press) have shown that "imitate the best" is negatively correlated with payoff uncertainty, When this is high novel traits, which are expected to be initially rare in a population, have a decreased probability of being selected even if their average payoff is higher than the extant suboptimal variants.
 
-\section{Payoff-biased Learning and Allee Effect}
-
-A key feature that underlies many models of cultural transmission is that the social learner is often guided by some form of evaluation of characteristics associated with potential teachers or demonstrators.
-To put it simply, one can decide to copy traits possessed by individuals showing some signal that suggests higher payoff, success, or fitness (e.g.
-number of offspring, income, presence/absence of key traits, etc.).
-This learning strategy has been broadly referred to as directed social learning \citep{coussi1995}, model-biased transmission \citep{boyd1985,henrich2001}, attraction model \citep{acerbi2014}, and payoff-based transmission\footnote{citation needed}.
-The exact details of these model differ to some extent, as some might imply that the target trait of interest and the observed payoff-signal is only weakly correlated (e.g.
-prestige biased transmission), while other entails a more direct link between the two.
-In one sense one can argue that they are part of a broader category of social learning strategy\footnote{citation needed - see also Crema and Lake, under review}.
-
-The degree of correlation between the payoff-signal and the target trait can also be viewed as a form of uncertainty or amount of information content offered by the signal.
-Previous studies \citep{schlag1998,baldini2013,lake2012} have explored how different learning strategies (e.g.
-copy-the-best, or copy-the-best-average, or copy-proportionally) show different dynamics, and are differently biased by this uncertainty.
-In some cases these might produce counter-intuitive outcomes.
-For instance, \citet{baldini2013} has shown that when the learning strategy is  copy-the-best suboptimal variants with higher payoff uncertainty and lower mean (equivalent to a lower correlation between signal payoff-signal and the target trait) can be favored over variants with higher mean and lower uncertainty.
-In other contexts, the uncertainty might be in the identification of the correct target trait, which could lead to the adoption of variants that have minimal or no correlation with the observed cue \citep[cf.
-"evolutionary hitchhiking"][]{ackland2007}.
-
-The uncertainty of the payoff signal is partly derived by the presence of interacting cultural and physical traits, as well as the environment where the trait is manifest.
-For example, a hunter might evaluate the performance of an arrowhead type (the target cultural trait) using as a cue the number of preys captured (the payoff signal).
-The correlation between the two is partly due to the efficiency of the arrowhead, but also by the bow, the prey species being capture and their availability, the physical properties of the hunter, and chance.
+Given the same cultural trait, the variability in the payoff signal is partly derived by the presence of interacting cultural and physical traits, as well as the environment where the trait is manifest.
+For example, a fisherman might evaluate the performance of a hook (the target cultural trait) using as a cue the number of fish captured (the payoff signal).
+The correlation between the two is partly due to the efficiency of the hook, but also by the rod, the choice of the bait, the availability of the prey species, the choice of the fishing spot, the physical properties of the fisherman, and chance.
 In many cases we should expect that the social teacher and learner share many of these interacting traits as well as the environment where these are displayed.
-In other words, cultural and ecological inheritance can generally reduce the uncertainty of the payoff-signal.
+In other words, cultural and ecological inheritance can generally reduce the uncertainty and the variability in the payoff-signal.
 There are however exceptions.
-One of these are contexts where the success of a variant is strongly correlated with the number of individuals possessing it within a population.
-This density dependence implies the adoption (or abandonment) of a trait by any individual can alter the expected fitness, and hence the payoff-signal, of all individuals within the population.
+One of these are contexts where the success of a variant is strongly correlated with the absolute number of individuals possessing the same trait within a population.
+This density dependence implies the adoption (or abandonment) of a trait can alter the expected payoff-signal, possibly generating a variety of feedback mechanism.
 
 Several cultural traits are expected to show such density dependence.
-For example, behavior linked to the exploitation or consumption of limited resources are expected to exhibit a negative frequency dependence.
+For example, behaviour linked to the exploitation of limited resources are expected to exhibit a negative density dependence.
 Novel optimal variants are likely to spread rapidly when the trait is rare, but once a threshold frequency is exceeded,  beneficial effect will start to decline, the payoff will become smaller, and eventually rarer suboptimal variants can be preferred (see Lake and Crema 2012 for an extensive analysis).
-Negative dependency is however not the only form frequency dependence we should expect.
-In some cases, cultural variants that are strongly based on direct or indirect cooperation might require a minimum frequency to provide sufficient fitness, or more broadly show a positive correlation with the number of individuals possessing the trait.
-This positive frequency dependence, known in ecology as Allee Effect \citep{allee1958}, can be expected at different meta-population levels.
-For example foraging groups are known to show higher performance when the number of individuals is not too small\footnote{There was a reference on Smith 1981...is this JM Smith 'evolution and theory of games'?
-This is 1982}, while certain activity such as anthropogenic fire \citep{bird2013} or selective hunting (Dods 2002) might require a sufficient population density to promote sufficient niche enhancement and a consequent increase in payoff \citep{rowley2011}.
-More\footnote{citation needed!} Language?
-Bettencourt Cities?
-\enrico{Ideally we need more example, not necessarily within the domain of subsistence activities.
-Any Allee Effect?
-Add Ostrom 2000 mention collective good problems here\footnote{what is ostrom 2000?}}.
+Density dependence can also have a positive direction, whereby an increase in the number of individuals possessing a trait might have a selective advantage additional to what is already expected from its frequency.
+Variants that are enhanced by some form of cooperation, or more in general traits exhibiting a positive niche construction are expected to show such a pattern, known in ecology as Allee Effect \citep{allee1958}.
+For example foraging groups are known to show higher performance when the number of individuals is not too small \citep{hill_and_hawkes_1983,janssen_and_hill_2014}, while certain activity such as anthropogenic fire \citep{bird2013} or selective hunting \citep{dods_2002} might require a population density beyond a certain threshold in order to promote sufficient niche enhancement and a consequent increase in payoff \citep{rowley-conwy_and_layton_2011}.
+Other traits that might be promoted by increased density includes language \citep{kandler2009}, organisational populations \citep{caroll_and_hannan_1989}, % and prosocial behaviour.
+%I think Ostrom 2000 was added here in relation to this. In any case we need one extra example here.
 
-Here we explore the combination of frequency dependent selection and payoff-based transmission with a modified version of the Lotka-Volterra model.
-We will demonstrate that degree by which a population relies on social learning can strongly affect the long-term equilibria of the system, with changes in the size of basins of different basins of attraction.
-Section 2 will illustrate our model, section 3 will explore three different scenarios, and section 4 will discuss our results and their broader implication.
-
-[RESULTS HERE...]
+Here we explore the combined effect of positive and negative density-dependent selection, assuming that: 1) the social learning strategy is payoff-based transmission; and 2)  payoff is equivalent to the reproductive fitness of each individual.
+We will demonstrate, by exploring a a two-trait difference equation, that the degree by which a population relies on social learning can strongly affect the evolutionary trajectory of the system, leading to the emergence of stable and unstable equilibria, as well as the occurrence of temporary or permanent episodes of reversion to suboptimal variants.
 
 \section{The Model}
-We consider the evolutionary dynamics of two mutually exclusive subsistence strategies which fitness have both positive and negative frequency dependence.
-In particular we examine how two population of individuals, m and n, change over time if we allow for a social learning strategy guided by the difference in the fitness of the individuals.
-The model can theoretically be extended to more than two strategies, but here we decide to explore a two trait version for simplicity.
-The core model can be depicted with the coupled difference equation \eqref{eq1}:
+We consider the evolutionary dynamics of two mutually exclusive subsistence strategies which fitness have both positive and negative density dependence.
+In particular we examine how two population of individuals, $a$ and $b$, change over time if we allow for a social learning strategy guided by a ``proportional imitation rule'', that is the probability of an individual with lower payoff to adopt a the trait of an individual with higher payoff is proportional to the difference in the payoff.
+The model can theoretically be extended to $n$ strategies, but here we decide to explore a two trait version for simplicity.
+The core model can be depicted with the following coupled difference equation:
 
@@
 \begin{cases}
-m_{t+1}& = m_t + m_t R_{m,t} + C_{m,n,t} \\
-n_{t+1}& = n_t + n_t R_{n,t} + C_{m,n,t}
+a_{t+1}& = a_t + a_t R_{a,t} + C_{a,b,t} \\
+b_{t+1}& = b_t + b_t R_{b,t} + C_{a,b,t}
 \end{cases}
@@
 
-where $m_{t+1}$ and $n_{t+1}$ is the population of each trait at time $t+1$ is the given by the number of individuals during the previous time-step ($m_t$ and $n_t$), $R_{m,t}$ and $R_{n,t}$ is the number of offspring generated by each population and $C_{m,n,t}$ is the number of individuals shifting from one strategy to the other.
+where $a_{t+1}$ and $b_{t+1}$ are the population of each trait at time $t+1$.
+This is given by the number of individuals during the previous time-step ($a_t$ and $b_t$), the reproductive rate of each population ($R_{a,t}$ and $R_{b,t}$), and the shift from one population to another dictated by cultural transmission ($C_{a,b,t}$).
 
-The intrinsic growth rate of each population can be obtained from the paired equation \eqref{eq2}:
+The reproductive growth rate of each population is defined by the following paired equation:
 
@@
 \begin{cases}
-R_{m,t}& = r_m (\frac{m_t}{A_m}-1)(1-\frac{m_t}{K_m})\\
-R_{n,t}& = r_n (\frac{n_t}{A_n}-1)(1-\frac{n_t}{K_n})
+R_{a,t}& = r_a (\frac{a_t}{A_a}-1)(1-\frac{a_t}{K_a})\\
+R_{b,t}& = r_b (\frac{b_t}{A_b}-1)(1-\frac{b_t}{K_b}) 
 \end{cases}
@@
 
-where $r$ is the basic growth rate, $A$ is the allee threshold, and $K$ is the carrying capacity, with $A < K$.
-The intrinsic growth rate will thus become negative when the population size is below $A$ or above $K$, and will reach its maximum value $R_{max} = \frac{(K-A)}{2}$, with $R_{max}$ equivalent to $\frac{r(A-K)2}{4AK}$.
-\xavi{I think this last sentence is confusing}
+where $r$ is the basic growth rate, $A$ is the \emph{Allee} threshold,  $K$ is the carrying capacity, and the condition $A < K$.
+Equation \eqref{eq2} implies that the reproductive rate is negative when density is below $A$ or above $K$, and will reach its maxima $R_{max}=\frac{r(A-K)2}{4AK}$ when the population density is equal to $\frac{(K-A)}{2}$.
 
-Cultural transmission is portrayed here as a migration from the population with lower payoff to the population with higher payoff, as seen in \eqref{eq3}:
+Cultural transmission is portrayed in \eqref{eq3}:
 
@@
 \label{eq3}
-C_{m,n,t} = 
+C_{a,b,t} = 
 \begin{cases}
-0& \text{if } R_{m,t} = R_{n,t}\\
-\zeta(m_t+m_tR_{m,t})& \text{if } R_{m,t} > R_{n,t}\\
--\zeta(n_t+n_tR_{n,t})& \text{if } R_{m,t} < R_{n,t}
+0& \text{if } R_{a,t} = R_{b,t}\\
+\zeta(a_t+a_tR_{a,t})& \text{if } R_{a,t} > R_{b,t}\\
+-\zeta(b_t+b_tR_{b,t})& \text{if } R_{a,t} < R_{b,t}
 \end{cases}
@@
 
-where $\zeta$ defines the proportion of individuals shifting strategy, and given by \eqref{eq4}:
+where $\zeta$ defines the proportion of individuals changing strategy, and given by \eqref{eq4}:
 
@@
 \begin{cases}
-z& \text{if }|R_{m,t}-R_{n,t}| > b\Delta\\
-z\frac{|R_{m,t}-R{n,t}|}{b\Delta}& \text{if }|R_{m,t}-R_{n,t}| \leq b\Delta\\
+z& \text{if }|R_{a,t}-R_{b,t}| > b\Delta\\
+z\frac{|R_{a,t}-R{b,t}|}{m\Delta}& \text{if }|R_{a,t}-R_{b,t}| \leq b\Delta\\
 \end{cases}
@@
 
-where $\zeta$ is the maximum transmission rate (i.e.
-the highest proportion of individuals shifting strategy), $\Delta$ is the maximum between $R_{max(m,t)}$ and $R_{max(n,t)}$, and $b$ is a calibration parameter that measure the perception of the difference in the payoffs (i.e.
-small values of b determines a higher rate by which the theoretical maximum transmission rate is reached).
+where $z$ is a measure of reliance in social learning (effectively equivalent to highest possible transmission rate),  $\Delta$ is the maximum between $R_{max(a,t)}$ and $R_{max(b,t)}$, and $m$ is a calibration parameter that measure the perception of the difference in the payoffs (i.e.
+small values of $m$ determines a higher rate by which the theoretical maximum transmission rate is reached).
+Thus \eqref{eq3} and \eqref{eq4}  conforms with the ``proportional imitation rule''  \citep{schlag1998}, although here the payoff is equivalent to the reproduction rate.
+
+\section{Results}
+
+We investigate the equilibrium conditions for different initial population sizes of  $a$ and $b$ and different settings of $z$, $A$, and $K$, assuming that the $r_a=r_b$, and for convenience holding the condition $K_b>K_a$.
+Given that equations \ref{eq1},~\ref{eq4} have no analytical solution, we identified equilibrium conditions via simulation (see appendix 1 for details; source codes can be found on the following repository %put link here
+as well as in the ESM).  
+%if the points we need to reiterate are short perhaps a footnote will suffice instead of an appendix?
+
+\subsection{Equilibrium conditions with no social learning, $z=0$}
+
+\subsection{Effect of social learning $z>0$}
+
+\subsection{Effect of competition}
 
@@
 
+%General Intro on Human History%
+Major transitions in human history have been marked by the fast spread of key innovations, which laid the foundations of further cumulative development in economy, society, and culture.
+Shifts in subsistence economy has often been regarded as a prime example of these transitions, triggered by the spread of a suite of behavioural traits such as the know-how of the domestication process \citep{barker2006}; the use of new subsistence tools %citation needed
+and processing techniques \citep{molleson1993}, and the inclusion/exclusion and extensification/intensification of different prey species %citation needed
+Archaeological evidence does  not support a simple progressive adoption of these key cultural variants, and instead a showcase a variety of trajectories, including the long-term coexistence of multiple subsistence strategies, or even reversions to older traits \citep{rowley2001}.
+  %add long list of citations. or perhaps one or two concrete examples (probably better)
+
+\enrico{General Intro on HBE? Not sure if needed for TPB? Maybe move to discussion?}
+Formal models portraying changes in human subsistence have been mostly offered in human behavioral ecology \citep{smith1992,bird2006,kennett2006}, where several mathematical models inspired from evolutionary ecology %we need a better one-line description of HBE.
+ used to predict  based on expected patterns under the assumption of optimality \citep{belovsky1988}. 
+While earliest models relied almost exclusively on this assumption, most recent works are increasing discussing how the knowledge of alternative subsistence strategies are acquired %citation needed?
+Are there such works?
+Need to check
+and whether this process of learning can drive the evolutionary trajectory of a given population \citep{henrich1998}, partly as a synergy with the dual inheritance theory \citep{boyd1985}.
+
+%Also this bits might be moved to discussion?%
+The results of these works often exhibits counter-intuitive outcomes that cannot be predicted by standard assumptions of optimality.
+For instance, extreme reliance on social learning can promote information parasitism \citep{giraldeau2002}, leading to a decline in the proportion of asocial learners that in certain environmental context can have strongly detrimental effects at the population level \citep{whitehead2009}.
+Other examples include instances where a reduced effective population size promotes an increase in random drift, and the ultimate loss of adaptive traits \citep{henrich2004}, or instances where neutral traits are jointly transmitted together with adaptive ones \citep{ackland2007}.
+These examples illustrate how theoretical models based on the dual inheritance theory can extend predictions of the standard optimal foraging theory, integrating dynamics that, at least at a first glance, are regarded as counter-intuitive.
+
+These include episodes of reversions (e.g.
+\footnote{citation needed}), where a more recent optimal strategies are abandoned for older suboptimal variants; stable coexistence (e.g.
+\footnote{citation needed}), where optimal and suboptimal strategies co-exist within a population; and even the complete loss of optimal variants \citep[e.g.][]{henrich2004}.
+The evolutionary processes leading to these different outcomes are undoubtedly contingent to local historical dynamics, nonetheless we argue that a simple model of social learning in combination with a positive and negative frequency dependent selection can lead all these equilibria.
+
 \enrico{We need to mention that some behaviour might actually enhance the other community (or be detrimental).
@@
 
+\section{Appendix I}
+
+-50,000 iterations
+-floating point arithmetic
+
 \end{document}

### 2015-05-11 bifurcationAnalysis/bifurcation.R
Bifurcation Analysis

@@ -1,0 +1,84 @@
+
+allee<-function(r,A=0.2,K=1,start=0.21,timesteps=10000)
+{
+	pop<-numeric(length= timesteps)
+	pop[1]=start
+	for (t in 2: timesteps)
+	{
+		pop[t]=pop[t-1]+pop[t-1]*r*(1-pop[t-1]/A)*(pop[t-1]/K-1)
+	}
+	
+	return(pop[(timesteps-499):timesteps])	
+}
+
+
+bifurcationPlot<-function(rSEQ,A,K,start,...)
+    {
+        plot(x=1,y=1,xlim=c(min(rSEQ),max(rSEQ)),type="n",ylab="x",xlab=expression(r[x]),...)
+        for (r in rSEQ)
+            {
+                points(x=rep(r,500),y=allee(r=r,A=A,K=K,start=start),pch=20,cex=0.5)
+            }
+    }
+
+
+                                        # Figure 1 #
+
+
+par(mfcol=c(3,3))
+bifurcationPlot(rSEQ=seq(0.01,0.8,0.005),A=0.2,K=1,start=0.21,ylim=c(0.2,1.2))
+title("A=0.2; K=1; Initial x = 0.21")
+bifurcationPlot(rSEQ=seq(0.01,0.8,0.005),A=0.2,K=1,start=0.8,ylim=c(0.2,1.2))
+title("A=0.2; K=1; Initial x = 0.8")
+bifurcationPlot(rSEQ=seq(0.01,0.8,0.005),A=0.2,K=1,start=1.1,ylim=c(0.2,1.2))
+title("A=0.2; K=1; Initial x = 1.1")
+####
+bifurcationPlot(rSEQ=seq(0.9,2,0.005),A=0.4,K=1,start=0.41,ylim=c(0.2,1.2))
+title("A=0.4; K=1; Initial x = 0.41")
+bifurcationPlot(rSEQ=seq(0.9,2,0.005),A=0.4,K=1,start=0.8,ylim=c(0.2,1.2))
+title("A=0.4; K=1; Initial x = 0.8")
+bifurcationPlot(rSEQ=seq(0.9,2,0.005),A=0.4,K=1,start=1.1,ylim=c(0.2,1.2))
+title("A=0.4; K=1; Initial x = 1.1")
+####
+bifurcationPlot(rSEQ=seq(0.2,1.2,0.005),A=0.4,K=1.5,start=0.41,ylim=c(0.2,2))
+title("A=0.4; K=1.5; Initial x = 0.41")
[…]

### 2015-05-11 manuscript/draft.tex
Updated Draft
-This is a temp. update so I can work from home in the weekend
Updated manuscript

@@
 
-We investigate the equilibrium conditions for different initial population sizes of  $a$ and $b$ and different settings of $z$, $A$, and $K$, assuming that the $r_a=r_b$, and for convenience holding the condition $K_b>K_a$.
+We investigate the equilibrium conditions for different initial population sizes of  $a$ and $b$ and different settings of $z$, $A$, and $K$, assuming that  $r_a=r_b$, and for convenience holding the condition $K_b \geq K_a$.
 Given that equations \ref{eq1},~\ref{eq4} have no analytical solution, we identified equilibrium conditions via simulation (see appendix 1 for details; source codes can be found on the following repository %put link here
 as well as in the ESM).  
-%if the points we need to reiterate are short perhaps a footnote will suffice instead of an appendix?
+%if the points we need to reiterate are short, perhaps a footnote will suffice instead of an appendix? 
+% 100,000 iterations 
+% basin plots resolution
 
@@
 
-\subsection{Effect of social learning $z>0$}
+When $z=0$, equations \ref{eq1},~\ref{eq4} \emph{de facto} portray a standard cubic growth model with Allee effect, where the success (or failure) of a given focal trait is simply given by its own density, and the density of alternative cultural traits has no influence.
+For a generic cultural trait $x$, the model predicts three point attractors at $x=0$ (extinction), $x=A_x$, and $x=K_x$.
+The equilibrium $x=A_x$ is unstable in all cases, whereas the stability of the node at $x=K_x$ is a function of the per capita growth rate $r_x$.
+Increasing the latter decreases the stability of fixed points \citep{scheuring_1999}, leading to the transition from point attractors (with equilibrium at $x=K_x$), to limit cycles and chaotic dynamics with islands of stability (see fig.
+\ref{fig:bifurcationDiagram}).
+% Not sure if there are strange attractors.
+Reference  any  bifurcation analysis on the equation we used.
+Scheuring 1999 on JTB?
+
+\begin{figure}[h!]
+  \centering
+      \includegraphics[width=0.7\textwidth]{./figures/figure1.jpg}
+  \caption{Bifurcation diagram of a single population model with $A_x=0.2$ and $K_x=0.8$. For $0<r_x<0.6649747$, any initial value of $x$ greater than $A_x$ converge to $K_x$.}
+    \label{fig:bifurcationDiagram}
+\end{figure}
+
+Here we consider only scenarios where the node $x=K_x$ is stable.
+In this case the initial value of $x(t=0)$ will determine the ultimate equilibria of the system: extinction, if $x(t=0)<A_x$; unstable equilibria, if $x(t=0)=A_x$; and stable equilibrium, if $x(t=0)>A_x$.
+Thus in a 2-trait model, the coexistence of both variants is guaranteed only when the initial density of both traits are above their respective Allee threshold, otherwise we should expect that either (i.e.
+$a(t=0) \geq A_a$ and $b(t=0)<A_b$ or $a(t=0)<A_a$ and $b(t=0) \geq A_b$) or both (i.e.
+when $a(t=0)<A_a$ and $b(t=0)<A_b$) variants will be ultimately extinct (see fig.
+\ref{fig:NoTransmissionBasin}).
+
+\begin{figure}[h!]
+  \centering
+      \includegraphics[width=0.7\textwidth]{./figures/figure2.pdf}
+  \caption{Basins of attraction for the four equilibria (mutual extinction: grey; extinction of $a$: blue; extinction of $b$: red; stable coexistence: green).
+olid lines represent trajectories of the system for different initial conditions, represented by dots.
+Model parameters: $A_a=0.2$; $K_a=0.7$; $A_b=0.5$; $K_b=1.0$; $r_a=r_b=0.05$}.
+    \label{fig:NoTransmissionBasin}
+\end{figure}
+
+\subsection{Effects of social learning $z>0$}
+
+When social learning is enabled (i.e.
+when $z>0$), the stability of the nodes are perturbed by individuals switching from one strategy to another.
+The minimum implication of these perturbations is a change in the basins of attractions.
+Initial conditions closer to the boundaries between one basin and another will most profoundly be affected, potentially leading to a complete different equilibria with slightest change in the degree of reliance on social learning.
+Figure \ref{fig:basicEffectSocialLearning} shows an example of this.
+When social learning is absent, the initial conditions $i$, $j$, and $k$ respectively result into ...
+
+%Figure showing z=0 and z=something to illustrate the effect of social learning
+
+When the magnitude of social learning is sufficiently high, the system can exhibit exhibit limit-cycles
+
+%Stability nodes remain unchanged for low z, high z determines unstable equilibria
+%illustrate examples with different settings
+%overlap analysis and changes in basin size
+%For stability analysis show that the settings of r guarantees that the system is at a point equilibrium. First Bifurcation Point plot.
 
 \subsection{Effect of competition}
-
-\section{Experiments}
-
-We investigate different initial conditions of $m$ and $n$ for 6 different values of z (0,0.05,0.1,0.2,0.3, and 0.5).
-For each parameter combination we apply equations (\ref{eq1},~\ref{eq4}) for 50,000 iterations and record the final values of $m$ and $n$, identifying the following equilibria.
-
-\enrico{We might need to use a rounding factor. And in fact these introduces to the problem of scale and difference between discrete and continuos  models.}
-\enrico{We can analytical solve equations 1~4 to shows that the E1, E2 and E3 are indeed equilibria. Perhaps in an appendix? (although this is quite obvious...)}
-
-We are aware that condition U can potentially become one of the other stable equilibria after the 50,000 simulations, hence we computed the Lyapunov exponent to define the stability of the system.
-We devised three core scenarios reflecting fixing b=2 and equating Rn to gRm, where g is given by the following equation:
-
-\xavi{equation5, lyapunov, where this does go?}
-
-so that when Rn=gRm,  Rmax(m,t) and Rmax(n,t) becomes equal. In all scenarios we considered Allee thresholds and carrying capacities as proportions by ensuring that max(Km,Kn)=1.
-
-\enrico{Justify choice of arbitrary values}
-
-\subsection{Scenario 1: Traits with identical payoffs (i.e. Am = An   and Km=Kn)}
-
-The first scenario assumes two traits that are mathematically equivalent, so that $A_m = A_m, K_m=K_n and r_m=r_n$.
-Figure 1 shows the basin of attractions of each of the four equilibria for different values of z, setting r=0.005, .A=0.2 and K=1.
-When the z=0, i.e.
-when cultural transmission is disabled we have straight boundaries between the basins of E1, E2m, E2n, and E3 defined by Allee threshold.
-In other words when both one or both the population are below A we have some extinction events (E3 when mt=0<Am and nt=0<An, E2m when mt=0>Am and nt=0<An and E3n when mt=0<Am and nt=0>An).
-When z>0, the boundaries of these basins changes and with high z we see the emergence of unstable equilibria (U).
-
-In all cases the E1 equilibria is maintained when $mt=0=nt=0 > Am=An$ as in this scenario differences arises only when m DIFFERENT OF n, and when the two populations are equally sized, C(m,n) is equal to 0 (see equation [3]).
-
-\subsection{Scenario 2: Overlapping payoffs (i.e.  Am <  An  < Km < Kn)}
-
-Am=0.2, An=0.5, Km=0.7,Kn=1
-
-\subsection{Scenario 3: Internal overlap  (i.e.  Am <  An < Kn <  Km )}
-
-An=0.2, Am=0.5, Km=0.7,Kn=1
-
-\subsection{A note on competition}
+%what if competition
+%show new stable points
+%show dynamics?
 
@@
 The following equation can substitute equation [2]:
-
-\xavi{equation6}
 
@@
 
-\section{Conclusion}
-
-BlaBlaBla
-
 \section{References}

### 2015-05-11 manuscript/figures/figure1.R
Figure 1 code and plot

@@ -1,0 +1,30 @@
+
+allee<-function(r,A=0.2,K=1,start=0.21,timesteps=10000)
+{
+	pop<-numeric(length= timesteps)
+	pop[1]=start
+	for (t in 2: timesteps)
+	{
+		pop[t]=pop[t-1]+pop[t-1]*r*(1-pop[t-1]/A)*(pop[t-1]/K-1)
+	}
+	
+	return(pop[(timesteps-499):timesteps])	
+}
+
+
+bifurcationPlot<-function(rSEQ,A,K,start,...)
+    {
+        plot(x=1,y=1,xlim=c(min(rSEQ),max(rSEQ)),type="n",...)
+        for (r in rSEQ)
+            {
+                points(x=rep(r,500),y=allee(r=r,A=A,K=K,start=start),pch=20,cex=0.5)
+            }
+    }
+
+par(mar=c(5, 6, 4, 2) + 2)
+bifurcationPlot(rSEQ=seq(0.001,1.1,0.001),A=0.2,K=0.8,start=0.21,ylim=c(0.2,1),las=2,cex.lab=6,cex.axis=6,xlab="",ylab="")
+mtext(side=1,expression(r[x]),line=6,cex=6)
+mtext(side=2,"x",line=6,cex=6)
+
+#axis(side=4,at=0.8,labels=expression(K[x]),las=3,cex.axis=3)
+dev.print(device=jpeg,"~/github/CultureAllee/manuscript/figures/figure1.jpg",width=4000,height=3000,pointsize = 30)

### 2015-05-11 manuscript/figures/figure2.R
Figure 2: Code and Plot

@@ -1,0 +1,39 @@
+source("./src/alleeIGC.R")
+source("./src/basin.R")
+source("./src/utility.R")
+
+lambda=0.3
+z=0
+Am=0.2
+An=Am+lambda
+Km=0.7
+Kn=Km+lambda
+cAn=0
+cAm=0
+cKn=0
+cKm=0
+r.n=0.05
+r.m=r.n
+x.lim=c(0,1.2)
+y.lim=c(0,1.2)
+
+
+
+figure=basin(x.lim=x.lim,y.lim=x.lim,points=41,timeSteps=5000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=FALSE,multicore=TRUE,cores=2)
+
+basinPlot(figure,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,xlab="a",ylab="b")
+abline(h=c(An,Kn),lty=2,col="white")
+abline(v=c(Am,Km),lty=2,col="white")
+axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
+axis(side=4,at=c(An,Kn),labels=c(expression(Ab),expression(Kb)),hadj=0.5,las=2)
+
+
+trajPlot(ini.m=0.1,ini.n=0.2,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
+trajPlot(ini.m=0.5,ini.n=0.2,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
+trajPlot(ini.m=0.1,ini.n=0.6,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
+trajPlot(ini.m=0.5,ini.n=0.7,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
[…]

### 2015-05-11 src/basin.R
Update basinPlot function
Axis label outside the core function.

@@ -79,7 +79,7 @@ basinPlot<-function(res,Am,Km,An,Kn,cKn,cKm,...)
         res$col[which(res$m.final==0&res$n.final==0)]=5  
         #plot(res[,1],res[,2],col=res$col,x.lab="m",y.lab="n",pch=15)
         tmp=rasterFromXYZ(res[,c(1,2,5)])
-        image(tmp,col=c("orange","darkgreen","indianred","royalblue","lightgrey"),zlim=c(1,5),xlab="m",ylab="n",...)
+        image(tmp,col=c("orange","darkgreen","indianred","royalblue","lightgrey"),zlim=c(1,5),...)
     }
 
 

### 2015-05-12 manuscript/draft.tex
Updated Manuscript

@@
   \caption{Basins of attraction for the four equilibria (mutual extinction: grey; extinction of $a$: blue; extinction of $b$: red; stable coexistence: green).
-olid lines represent trajectories of the system for different initial conditions, represented by dots.
+Solid lines represent trajectories of the system for different initial conditions, represented by dots.
 Model parameters: $A_a=0.2$; $K_a=0.7$; $A_b=0.5$; $K_b=1.0$; $r_a=r_b=0.05$}.
@@
 Figure \ref{fig:basicEffectSocialLearning} shows an example of this.
-When social learning is absent, the initial conditions $i$, $j$, and $k$ respectively result into ...
+When social learning is not present (fig.
+\ref{fig:basicEffectSocialLearning} -a), the initial conditions $i$, $j$, and $k$ result into a stable equilibrium at $a=K_a$ and $b=K_b$, whilst the starting point $l$ leads to the extinction of $b$.
+When $z=0.05$ (fig.\ref{fig:TransmissionBasin}-b), the basins of attractions are modified, resulting into different trajectories and possibly equilibria for the same initial conditions without social learning.
+Thus $i$ and $k$ are no longer leading to the coexistence of both traits but to the demise of one or the other, whilst $l$ now ensure the coexistence of both variants.
 
-%Figure showing z=0 and z=something to illustrate the effect of social learning
+\begin{figure}[h!]
+  \centering
+      \includegraphics[width=\textwidth]{./figures/figure3.pdf}
+  \caption{Basins of attraction for the four equilibria (mutual extinction: grey; extinction of $a$: blue; extinction of $b$: red; stable coexistence: green) with (right) and without (left) social learning.
+Model parameters: $A_a=0.2$; $K_a=0.8$; $A_b=0.4$; $K_b=1.0$; $r_a=r_b=0.05$; $z=0$ (left panel), $z=0.05$ (left panel) }
+    \label{fig:TransmissionBasin}
+\end{figure}
 
-When the magnitude of social learning is sufficiently high, the system can exhibit exhibit limit-cycles
+When the magnitude of social learning is sufficiently high, the system is subject to effects similar to effects to increase in the per-capita growth rate, with the consequent decrease in the stability of the fixed points.
+Figure \ref{fig:bifurcationWithTransmission} shows the bifurcation diagram for three different initial settings of $a$ and $b$.
+In the first two cases, increase in $z$ leads to a shift of the equilibria, that remains however constant for larger reliance on social learning.
+The third case shows, around $z=0.5$ a transition from a point attractor to a system characterised by repeated shifts between the two transmission strategies.
 
-%Stability nodes remain unchanged for low z, high z determines unstable equilibria
-%illustrate examples with different settings
-%overlap analysis and changes in basin size
+\begin{figure}[h!]
+  \centering
+      \includegraphics[width=\textwidth]{./figures/figure4.jpg}
+  \caption{Bifurcation diagram for $a$ (red) and $b$ (blue) as a function of social learning ($z$): left: $a=0.11$, $b=0.4$; Middle: $a=0.5$, $b=0.21$, $a=0.4$, $b=0.4$.
+Model parameters:  $A_a=0.1$; $K_a=0.6$; $A_b=0.2$; $K_b=0.7$; $r_a=r_b=0.05$}
+    \label{fig:bifurcationWithTransmission}
+\end{figure}
+
+%Overlap Analysis Here
 %For stability analysis show that the settings of r guarantees that the system is at a point equilibrium. First Bifurcation Point plot.

### 2015-05-12 manuscript/figures/figure3.R
Figure 3 Code, Plot, and Data

@@ -1,0 +1,75 @@
+source("./src/alleeIGC.R")
+source("./src/basin.R")
+source("./src/utility.R")
+
+                                        #Figure 3
+lambda=0.15
+z=0.05
+Am=0.2
+An=0.4
+Km=0.8
+Kn=1
+cAn=0
+cAm=0
+cKn=0
+cKm=0
+r.n=0.05
+r.m=r.n
+x.lim=c(0,1.2)
+y.lim=c(0,1.2)
+
+
+z=0
+
+figure3a=basin(x.lim=x.lim,y.lim=x.lim,points=41,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=TRUE,multicore=TRUE,cores=5)
+
+z=0.05
+
+figure3b=basin(x.lim=x.lim,y.lim=x.lim,points=150,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=TRUE,multicore=TRUE,cores=5)
+
+save.image("figure3.RData")
+
+
+load("/Users/enryu/github/CultureAllee/manuscript/figures/figure3.RData")
+setwd("~/github/CultureAllee/")
+source("./src/alleeIGC.R")
+source("./src/basin.R")
+source("./src/utility.R")
+
+par(mfrow=c(1,2))
+basinPlot(res=figure3a,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main="a",xlab="a",ylab="b")
+abline(h=c(An,Kn),lty=2,col="white")
+abline(v=c(Am,Km),lty=2,col="white")
+axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
+axis(side=4,at=c(An,Kn),labels=c(expression(Ab),expression(Kb)),hadj=0.5,las=2)
+z=0
[…]

### 2015-05-12 manuscript/figures/figure4.R
Figure 4: Code and Plot

@@ -1,0 +1,64 @@
+source("./src/alleeIGC.R")
+source("./src/basin.R")
+source("./src/utility.R")
+
+
+lambda=0.1
+z=0.5
+Am=0.1
+An=Am+lambda
+Km=0.6
+Kn=Km+lambda
+cAn=0
+cAm=0
+cKn=0
+cKm=0
+r.n=0.05
+r.m=r.n
+x.lim=c(0,1.2)
+y.lim=c(0,1.2)
+
+
+
+bifurcationPlotSingle<-function(Zseq=seq(0,0.7,0.005),timeSteps=10000,...)
+{
+    plot(x=1,y=1,xlim=c(min(Zseq),max(Zseq)),type="n",ylab="a",xlab="z",ylim=c(0,1.2))
+    for (z in Zseq)
+        {
+            tmp=unique(allee(timeSteps=timeSteps,storeFinalOnly=FALSE,z=z,ini.m=0.4,ini.n=0.4,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm)[[1]][(timeSteps-500):timeSteps])
+            points(x=rep(z,length(tmp)),y=tmp,pch=20,cex=0.5)
+               }
+        }
+
+bifurcationPlotSingle(Zseq=seq(0,0.7,0.005),timeSteps=10000,ini.m=0.4,ini.n=0.4,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm)
+bifurcationPlotSingle(Zseq=seq(0,0.7,0.005),timeSteps=10000,ini.m=0.7,ini.n=0.4,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm)
+
+
+
+
+cexSetting=10
+bifurcationPlotDouble<-function(Zseq=seq(0,0.7,0.005),timeSteps=10000,cexSetting,...)
+{
+    plot(x=1,y=1,xlim=c(min(Zseq),max(Zseq)),type="n",ylab="",xlab="z",ylim=c(0,1.2),cex.axis=cexSetting,cex.lab=cexSetting)
+    for (z in Zseq)
+        {
[…]

### 2015-05-15 manuscript/draft.tex
small edits and added citations

@@
 This paper explores the evolutionary dynamics of a two trait model where each variant exhibit Allee effect.
-We used a modified Lotka-Volterra model of payoff-biased social learning, and show that that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion to suboptimal variants, long-term co-existence of multiple variants, and fixation to a single optimal or suboptimal strategy.
+We used a modified Lotka-Volterra model of payoff-biased social learning, and show that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion to suboptimal variants, long-term co-existence of multiple variants, and fixation to a single optimal or suboptimal strategy.
+
+% XAVI: I'm confused by the use of trait and variant: shouldn't it be 'a single trait model with two variants'?
 \end{abstract}
@@
 This information can be based on some properties of the variant itself (\emph{content-bias}) or derived from contextual cues (\emph{context-bias}), such as the rarity or the commonness or the rareness of a trait, or some features associated with the model (i.e.
-the social teacher) such as prestige, success, or the similarity to the learner \citep{henrich_mcelreath2003}.
-An expected consequence of \emph{context-bias} is the possibility of a feedback process, whereby the attractiveness of a given might change as a consequence of the learning strategy \citep{kendal_etal_2009}.
+the social teacher) such as prestige, success, or the similarity to the learner~\citep{henrich_mcelreath2003}.
+An expected consequence of \emph{context-bias} is the possibility of a feedback process, whereby the attractiveness of a given variant might change as a consequence of the learning strategy~\citep{kendal_etal_2009}.
 For example a strong anti-conformist bias might initially favour rare traits, but as they spread into a population their selective advantage will decline as the result of a negative feedback.
@@
 
-This paper examines a group of learning strategies generally referred to as \emph{payoff-based transmission} \citep{schlag1998,kendal_etal_2009,lake_and_crema_2012,baldini2013,kandler_and_laland_2013,crema_lake_inpress}.
+This paper examines a group of learning strategies generally referred to as \emph{payoff-based transmission}~\citep{schlag1998,kendal_etal_2009,lake_and_crema_2012,baldini2013,kandler_and_laland_2013,crema_lake_inpress}.
 The core assumption shared by these models is that the probability of copying a given cultural variant is positively correlated with some payoff signal expressed by individuals possessing such a variant.
-Several variants of this learning strategies have been proposed and explored, including: "imitate the best", where the social learner identifies and copies the trait possessed by the individual with the highest payoff; "compare means" where the social learner compare the average payoff of different strategies and chooses the highest variant; and the "proportional imitation rule", where the probability of adopting a variant is proportional to the difference in payoff between the demonstrator and the learner \citep{schlag1998,baldini2013}.
+Several variants of this learning strategies have been proposed and explored, including: \emph{imitate the best}, where the social learner identifies and copies the trait possessed by the individual with the highest payoff; \emph{compare means}, where the social learner compare the average payoff of different strategies and chooses the highest variant; and the \emph{proportional imitation rule}, where the probability of adopting a variant is proportional to the difference in payoff between the demonstrator and the learner~\citep{schlag1998,baldini2013}.
 
@@
 The magnitude of this uncertainty can have strong impact to efficiency of specific learning rules.
-For example, Crema and Lake (In press) have shown that "imitate the best" is negatively correlated with payoff uncertainty, When this is high novel traits, which are expected to be initially rare in a population, have a decreased probability of being selected even if their average payoff is higher than the extant suboptimal variants.
+For example,~\citet{crema_lake_inpress} have shown that \emph{imitate the best} is negatively correlated with payoff uncertainty.
+In the case of high uncertainty, a novel trait with payoff higher than the average has a decreased probability of being selected, due to being initially rare in a population.
+This can generate scenarios dominated by extant suboptimal variants.
 
@@
 Several cultural traits are expected to show such density dependence.
-For example, behaviour linked to the exploitation of limited resources are expected to exhibit a negative density dependence.
+In particular, behaviour linked to the exploitation of limited resources are expected to exhibit a negative density dependence.
 Novel optimal variants are likely to spread rapidly when the trait is rare, but once a threshold frequency is exceeded,  beneficial effect will start to decline, the payoff will become smaller, and eventually rarer suboptimal variants can be preferred (see Lake and Crema 2012 for an extensive analysis).
 Density dependence can also have a positive direction, whereby an increase in the number of individuals possessing a trait might have a selective advantage additional to what is already expected from its frequency.
-Variants that are enhanced by some form of cooperation, or more in general traits exhibiting a positive niche construction are expected to show such a pattern, known in ecology as Allee Effect \citep{allee1958}.
-For example foraging groups are known to show higher performance when the number of individuals is not too small \citep{hill_and_hawkes_1983,janssen_and_hill_2014}, while certain activity such as anthropogenic fire \citep{bird2013} or selective hunting \citep{dods_2002} might require a population density beyond a certain threshold in order to promote sufficient niche enhancement and a consequent increase in payoff \citep{rowley-conwy_and_layton_2011}.
-Other traits that might be promoted by increased density includes language \citep{kandler2009}, organisational populations \citep{caroll_and_hannan_1989}, % and prosocial behaviour.
-%I think Ostrom 2000 was added here in relation to this. In any case we need one extra example here.
+Variants that are enhanced by some form of cooperation, or more in general traits exhibiting a positive niche construction are expected to show such a pattern, known in ecology as \emph{Allee Effect}~\citep{allee1958}.
+For example foraging groups are known to show higher performance when the number of individuals is not too small~\citep{hill_and_hawkes_1983,janssen_and_hill_2014}, while certain activities such as anthropogenic fire~\citep{bird2013} or selective hunting~\citep{dods_2002} might require a population density beyond a certain threshold in order to promote sufficient niche enhancement and a consequent increase in payoff~\citep{rowley-conwy_and_layton_2011}.
+Other traits that might be promoted by increased density includes language~\citep{kandler2009}, organisational populations~\citep{caroll_and_hannan_1989} and communication technologies~\citep{van_slyke_perceived_2007}.
 
 Here we explore the combined effect of positive and negative density-dependent selection, assuming that: 1) the social learning strategy is payoff-based transmission; and 2)  payoff is equivalent to the reproductive fitness of each individual.
-We will demonstrate, by exploring a a two-trait difference equation, that the degree by which a population relies on social learning can strongly affect the evolutionary trajectory of the system, leading to the emergence of stable and unstable equilibria, as well as the occurrence of temporary or permanent episodes of reversion to suboptimal variants.
+We will demonstrate, by exploring a two-trait difference equation, that the degree by which a population relies on social learning can strongly affect the evolutionary trajectory of the system, leading to the emergence of stable and unstable equilibria, as well as the occurrence of temporary or permanent episodes of reversion to suboptimal variants.
+% XAVI not sure if social learning or allee is more important...what about: "we will demonstrate...that the impact of Allee effect on social learning can strongly affect the evolutionary trajectory of the system
 
 \section{The Model}
-We consider the evolutionary dynamics of two mutually exclusive subsistence strategies which fitness have both positive and negative density dependence.
-In particular we examine how two population of individuals, $a$ and $b$, change over time if we allow for a social learning strategy guided by a ``proportional imitation rule'', that is the probability of an individual with lower payoff to adopt a the trait of an individual with higher payoff is proportional to the difference in the payoff.
+We consider the evolutionary dynamics of two mutually exclusive subsistence strategies whose fitness has both positive and negative density dependence.
+In particular we examine how two population of individuals, $a$ and $b$, change over time if we allow for a social learning strategy guided by a \emph{proportional imitation rule} (i.e.
+the probability of an individual with lower payoff to adopt the trait of an individual with higher payoff is proportional to the difference in the payoff).
 The model can theoretically be extended to $n$ strategies, but here we decide to explore a two trait version for simplicity.
-The core model can be depicted with the following coupled difference equation:
+The core model can be depicted with the coupled difference equation \eqref{eq1}:
 
 \begin{align}
-\label{eq1}
 \begin{cases}
@@
 \end{cases}
+\label{eq1}
 \end{align}
 
-where $a_{t+1}$ and $b_{t+1}$ are the population of each trait at time $t+1$.
+where $a_{t+1}$ and $b_{t+1}$ are the population using each trait at time $t+1$.
 This is given by the number of individuals during the previous time-step ($a_t$ and $b_t$), the reproductive rate of each population ($R_{a,t}$ and $R_{b,t}$), and the shift from one population to another dictated by cultural transmission ($C_{a,b,t}$).
-
-The reproductive growth rate of each population is defined by the following paired equation:
+% XAVI do you want to use cultural trasmission here or better stick with social learning?
+The reproductive growth rate $R$ of each population is defined by the following paired equation:
 
 \begin{align}
-\label{eq2}
 \begin{cases}
@@
 \end{cases}
+\label{eq2}
 \end{align}
 
-where $r$ is the basic growth rate, $A$ is the \emph{Allee} threshold,  $K$ is the carrying capacity, and the condition $A < K$.
+where $r$ is the basic growth rate, $A$ is the \emph{Allee} threshold and $K$ is the carrying capacity, following the condition $A < K$.
 Equation \eqref{eq2} implies that the reproductive rate is negative when density is below $A$ or above $K$, and will reach its maxima $R_{max}=\frac{r(A-K)2}{4AK}$ when the population density is equal to $\frac{(K-A)}{2}$.
@@
 
+% XAVI b is not explained?
 where $z$ is a measure of reliance in social learning (effectively equivalent to highest possible transmission rate), $\Delta$ is the maximum between $R_{max(a,t)}$ and $R_{max(b,t)}$, and $m$ is a calibration parameter that measure the perception of the difference in the payoffs (i.e.
 small values of $m$ determines a higher rate by which the theoretical maximum transmission rate is reached).
-Thus \eqref{eq3} and \eqref{eq4}  conforms with the ``proportional imitation rule''  \citep{schlag1998}, although here the payoff is equivalent to the reproduction rate.
+Thus \eqref{eq3} and \eqref{eq4}  conforms with the \emph{proportional imitation rule}~\citep{schlag1998}, although here the payoff is equivalent to the reproduction rate.
 
@@
 We investigate the equilibrium conditions for different initial population sizes of  $a$ and $b$ and different settings of $z$, $A$, and $K$, assuming that  $r_a=r_b$, and for convenience holding the condition $K_b \geq K_a$.
-Given that equations \ref{eq1},~\ref{eq4} have no analytical solution, we identified equilibrium conditions via simulation (see appendix 1 for details; source codes can be found on the following repository %put link here
+Given that equations \ref{eq1}-\ref{eq4} have no analytical solution, we identified equilibrium conditions via simulation (see appendix 1 for details; source code can be found on the following repository %put link here
 as well as in the ESM).  
@@
 
-\subsection{Equilibrium conditions with no social learning, $z=0$}
+\subsection{Equilibrium conditions without social learning}
 
-When $z=0$, equations \ref{eq1},~\ref{eq4} \emph{de facto} portray a standard cubic growth model with Allee effect, where the success (or failure) of a given focal trait is simply given by its own density, and the density of alternative cultural traits has no influence.
+When $z=0$, equations \ref{eq1}-\ref{eq4} \emph{de facto} portray a standard cubic growth model with Allee effect, where the success (or failure) of a given focal trait is simply given by its own density, and the density of alternative cultural traits has no influence.
 For a generic cultural trait $x$, the model predicts three point attractors at $x=0$ (extinction), $x=A_x$, and $x=K_x$.
 The equilibrium $x=A_x$ is unstable in all cases, whereas the stability of the node at $x=K_x$ is a function of the per capita growth rate $r_x$.
-Increasing the latter decreases the stability of fixed points \citep{scheuring_1999}, leading to the transition from point attractors (with equilibrium at $x=K_x$), to limit cycles and chaotic dynamics with islands of stability (see fig.
+Increasing the latter decreases the stability of fixed points~\citep{scheuring_1999}, leading to the transition from point attractors (with equilibrium at $x=K_x$), to limit cycles and chaotic dynamics with islands of stability (see fig.
 \ref{fig:bifurcationDiagram}).
@@
 
-\subsection{Effects of social learning $z>0$}
+\subsection{Effects of social learning}
 
@@
 when $z>0$), the stability of the nodes are perturbed by individuals switching from one strategy to another.
-The minimum implication of these perturbations is a change in the basins of attractions.
+The minimum implication of these perturbations is a change in the shape of the basins of attractions.
 Initial conditions closer to the boundaries between one basin and another will most profoundly be affected, potentially leading to a complete different equilibria with slightest change in the degree of reliance on social learning.
-Figure \ref{fig:basicEffectSocialLearning} shows an example of this.
-When social learning is not present (fig.
-\ref{fig:basicEffectSocialLearning} -a), the initial conditions $i$, $j$, and $k$ result into a stable equilibrium at $a=K_a$ and $b=K_b$, whilst the starting point $l$ leads to the extinction of $b$.
-When $z=0.05$ (fig.\ref{fig:TransmissionBasin}-b), the basins of attractions are modified, resulting into different trajectories and possibly equilibria for the same initial conditions without social learning.
+Figure \ref{fig:TransmissionBasin} shows an example of this.
+When social learning is not present (fig.~\ref{fig:TransmissionBasin}a), the initial conditions $i$, $j$, and $k$ result into a stable equilibrium at $a=K_a$ and $b=K_b$, whilst the starting point $l$ leads to the extinction of $b$.
+When $z=0.05$ (fig.~\ref{fig:TransmissionBasin}b), the basins of attractions are modified, resulting into different trajectories and possibly equilibria for the same initial conditions without social learning.
 Thus $i$ and $k$ are no longer leading to the coexistence of both traits but to the demise of one or the other, whilst $l$ now ensure the coexistence of both variants.
@@
 
-When the magnitude of social learning is sufficiently high, the system is subject to effects similar to effects to increase in the per-capita growth rate, with the consequent decrease in the stability of the fixed points.
+When the magnitude of social learning is sufficiently high, the system is subject to effects similar to an increase in the per capita growth rate $r$, with the consequent decrease in the stability of the fixed points.
 Figure \ref{fig:bifurcationWithTransmission} shows the bifurcation diagram for three different initial settings of $a$ and $b$.
@@
 The third case shows, around $z=0.5$ a transition from a point attractor to a system characterised by repeated shifts between the two transmission strategies.
-
+% XAVI shall we generate the pdf here?
 \begin{figure}[h!]
@@
 
-The model described in equations  (\ref{eq1},~\ref{eq4}) assumes that the resource pool of the two strategy are completely independent to each other.
+The model described in equations \ref{eq1}-\ref{eq4} assumes that the resource pool of the two strategy are completely independent to each other.
 In reality many strategies have some degree of competition, whereby the frequency dependence is not exclusively internal.
 In other words, the frequency of one trait might increase or decrease the fitness of the other.
-The following equation can substitute equation [2]:
+The following equation can substitute equation~\eqref{eq2}:
 
@@
 Major transitions in human history have been marked by the fast spread of key innovations, which laid the foundations of further cumulative development in economy, society, and culture.
-Shifts in subsistence economy has often been regarded as a prime example of these transitions, triggered by the spread of a suite of behavioural traits such as the know-how of the domestication process \citep{barker2006}; the use of new subsistence tools %citation needed
-and processing techniques \citep{molleson1993}, and the inclusion/exclusion and extensification/intensification of different prey species %citation needed
-Archaeological evidence does  not support a simple progressive adoption of these key cultural variants, and instead a showcase a variety of trajectories, including the long-term coexistence of multiple subsistence strategies, or even reversions to older traits \citep{rowley2001}.
+Shifts in subsistence economy has often been regarded as a prime example of these transitions, triggered by the spread of a suite of behavioural traits such as the know-how of the domestication process~\citep{barker2006} the use of new subsistence tools~\citep{petraglia_population_2009} and processing techniques~\citep{molleson1993}, and the inclusion/exclusion and extensification/intensification of different prey species %citation needed.
+Archaeological evidence does not support a simple progressive adoption of these key cultural variants, and instead a showcase a variety of trajectories, including the long-term coexistence of multiple subsistence strategies, or even reversions to older traits~\citep{rowley2001}.
   %add long list of citations. or perhaps one or two concrete examples (probably better)
-
 \enrico{General Intro on HBE? Not sure if needed for TPB? Maybe move to discussion?}
-Formal models portraying changes in human subsistence have been mostly offered in human behavioral ecology \citep{smith1992,bird2006,kennett2006}, where several mathematical models inspired from evolutionary ecology %we need a better one-line description of HBE.
- used to predict  based on expected patterns under the assumption of optimality \citep{belovsky1988}. 
+Formal models portraying changes in human subsistence have been mostly offered in human behavioral ecology~\citep{smith1992,bird2006,kennett2006}, where several mathematical models inspired from evolutionary ecology %we need a better one-line description of HBE.
+ used to predict  based on expected patterns under the assumption of optimality~\citep{belovsky1988}. 
 While earliest models relied almost exclusively on this assumption, most recent works are increasing discussing how the knowledge of alternative subsistence strategies are acquired %citation needed?
@@
 Need to check
-and whether this process of learning can drive the evolutionary trajectory of a given population \citep{henrich1998}, partly as a synergy with the dual inheritance theory \citep{boyd1985}.
+and whether this process of learning can drive the evolutionary trajectory of a given population~\citep{henrich1998}, partly as a synergy with the dual inheritance theory~\citep{boyd1985}.
 
-%Also this bits might be moved to discussion?%
 The results of these works often exhibits counter-intuitive outcomes that cannot be predicted by standard assumptions of optimality.
-For instance, extreme reliance on social learning can promote information parasitism \citep{giraldeau2002}, leading to a decline in the proportion of asocial learners that in certain environmental context can have strongly detrimental effects at the population level \citep{whitehead2009}.
-Other examples include instances where a reduced effective population size promotes an increase in random drift, and the ultimate loss of adaptive traits \citep{henrich2004}, or instances where neutral traits are jointly transmitted together with adaptive ones \citep{ackland2007}.
+For instance, extreme reliance on social learning can promote information parasitism~\citep{giraldeau2002}, leading to a decline in the proportion of asocial learners that in certain environmental context can have strongly detrimental effects at the population level~\citep{whitehead2009}.
+Other examples include instances where a reduced effective population size promotes an increase in random drift, and the ultimate loss of adaptive traits~\citep{henrich2004}, or instances where neutral traits are jointly transmitted together with adaptive ones~\citep{ackland2007}.
 These examples illustrate how theoretical models based on the dual inheritance theory can extend predictions of the standard optimal foraging theory, integrating dynamics that, at least at a first glance, are regarded as counter-intuitive.
@@
 \footnote{citation needed}), where a more recent optimal strategies are abandoned for older suboptimal variants; stable coexistence (e.g.
-\footnote{citation needed}), where optimal and suboptimal strategies co-exist within a population; and even the complete loss of optimal variants \citep[e.g.][]{henrich2004}.
+\footnote{citation needed}), where optimal and suboptimal strategies co-exist within a population; and even the complete loss of optimal variants~\citep[e.g.][]{henrich2004}.
 The evolutionary processes leading to these different outcomes are undoubtedly contingent to local historical dynamics, nonetheless we argue that a simple model of social learning in combination with a positive and negative frequency dependent selection can lead all these equilibria.
@@
 \enrico{We need to mention that some behaviour might actually enhance the other community (or be detrimental).
-This can be modelled using a competition coefficient as in \citep{jang2013}.
+This can be modelled using a competition coefficient as in~\citep{jang2013}.
 I don't think we need to explore this but it is definitely worth mentioning!}

### 2015-05-15 manuscript/draft.tex
overlap section

@@
   \caption{Basins of attraction for the four equilibria (mutual extinction: grey; extinction of $a$: blue; extinction of $b$: red; stable coexistence: green) with (right) and without (left) social learning.
-Model parameters: $A_a=0.2$; $K_a=0.8$; $A_b=0.4$; $K_b=1.0$; $r_a=r_b=0.05$; $z=0$ (left panel), $z=0.05$ (left panel) }
+Model parameters: $A_a=0.2$; $K_a=0.8$; $A_b=0.4$; $K_b=1.0$; $r_a=r_b=0.05$; $z=0$ (left panel) and $z=0.05$ (right panel) }
     \label{fig:TransmissionBasin}
@@
 
-%Overlap Analysis Here
-%For stability analysis show that the settings of r guarantees that the system is at a point equilibrium. First Bifurcation Point plot.
+\subsection{Effects of strategy overlap}
+% XAVI better title?
+
+Previous results suggest that the percentage of population that can switch strategies (i.e.
+the $z$ parameter) has major effects on the final equilibria.
+However, this will also depend on the interaction between the intervals where frequency dependence is $\geq 0$ (i.e.
+$[A_a,K_a]$ and $[A_b,K_b]$) .
+We have explored this relation introducing a new parameter $\lambda \in (0,1)$ and defining the positive interval for strategy $b$ as follows:
+
+\begin{align}
+\label{eqOverlap}
+\begin{cases}
+A_b = A_a + \lambda A_a\\
+K_b = K_a + \lambda K_a
+\end{cases}
+\end{align}
+
+The dynamics of two strategies are similar for low values of $\lambda$.
+High values indicate that the minimum number of individuals needed to achive positive fitness in $b$ gets closer to $K_a$, while the population size that $b$ can sustain also grows.
+As a consequence, the number of individuals switching strategies when $b$ has lower payoff is also higher and could potentially overload $a$ due to its lower $K_a$.
+
+Figure XX shows how the increase in $\lambda$ affects the equilibria of our model.
+
+% XAVI add results and figure
 
@@
 In other words, the frequency of one trait might increase or decrease the fitness of the other.
-The following equation can substitute equation~\eqref{eq2}:
+The following equation can substitute equation~\eqref{eqCompetition}:
 
-where the four new parameters cAn, cKn, cAm, and cKm, measure respectively the negative (or positive) effect determined by the number of individuals of the “other” traits.
-When cA>0 the Allee threshold is essentially increased, while when cK>0 the carrying capacity is diminished as function of the other population., with larger values of c dictates a stronger influence of the other population.
+\begin{align}
+\begin{cases}
+R_{a,t}& = r_a (\frac{a_t-c_{A_a}b_t}{A_a}-1)(1-\frac{a_t-c_{K_a}b_t}{K_a})\\
+R_{b,t}& = r_b (\frac{b_t-c_{A_b}a_t}{A_b}-1)(1-\frac{b_t-c_{K_b}a_t}{K_b}) 
+\end{cases}
+\label{eqCompetition}
+\end{align}
 
-One simple scenario is to assume that one strategy, say n, is dominant over its alternative m.
-We can  set cAn=0.1,  cKn=0.1,  cAm=-0.1, and  cKm=-0.1.
-Here negative values of c indicate a positive enhancement of one strategy over the other (in this case n gets benefit to the expense of m).
+where the four new parameters $c_{A_a}$, $c_{A_b}$, $c_{K_a}$ and $c_{K_b}$, measure respectively the negative (or positive) effect determined by the number of individuals of the “other” traits.
+When $c_{A}>0$ the Allee threshold is essentially increased, while when $c_{K}>0$ the carrying capacity is diminished as function of the other population.
+Larger values of any of the $c$ parameters dictates a stronger influence of the other population.
+
+Figure XX shows a scenario where strategy $a$ is dominant over strategy $b$.
+% XAVI add results and figure
 

### 2015-05-15 manuscript/figures/figure5.R
added new figures

@@ -1,0 +1,43 @@
+source("../../src/alleeIGC.R")
+source("../../src/basin.R")
+source("../../src/utility.R")
+
+z=0.05
+Am=0.2
+Km=0.8
+
+cAn=0
+cAm=0
+cKn=0
+cKm=0
+r.n=0.05
+r.m=r.n
+x.lim=c(0,1.2)
+y.lim=c(0,1.2)
+
+lambda=0
+diff = Km-Am
+An=Am + lambda*diff
+Kn=Km + lambda*diff
+
+
+figure5a <- basin(x.lim=x.lim,y.lim=x.lim,points=150,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=TRUE,multicore=TRUE,cores=4)
+
+lambda=0.5
+diff = Km-Am
+An=Am + lambda*diff
+Kn=Km + lambda*diff
+
+figure5b <- basin(x.lim=x.lim,y.lim=x.lim,points=150,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=TRUE,multicore=TRUE,cores=4)
+
+lambda=0.9
+diff = Km-Am
+An=Am + lambda*diff
+Kn=Km + lambda*diff
+
+figure5b <- basin(x.lim=x.lim,y.lim=x.lim,points=150,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=TRUE,multicore=TRUE,cores=4)
+
+basinPlot(res=figure5c,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main="b",xlab="a",ylab="b")
+basinPlot(res=figure5c,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main="b",xlab="a",ylab="b")
+basinPlot(res=figure5c,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main="b",xlab="a",ylab="b")
+

### 2015-05-15 manuscript/figures/figure6.R

@@ -1,0 +1,34 @@
+source("../../src/alleeIGC.R")
+source("../../src/basin.R")
+source("../../src/utility.R")
+
+lambda=0.15
+z=0.05
+Am=0.2
+An=0.4
+Km=0.8
+Kn=1
+r.n=0.05
+r.m=r.n
+x.lim=c(0,1.2)
+y.lim=c(0,1.2)
+
+cAn=0.25
+cAm=-0.25
+cKn=0.25
+cKm=-0.25
+
+figure6a <- basin(x.lim=x.lim,y.lim=x.lim,points=150,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=TRUE,multicore=TRUE,cores=4)
+
+cAn=-0.25
+cAm=0.25
+cKn=-0.25
+cKm=0.25
+
+figure6b <- basin(x.lim=x.lim,y.lim=x.lim,points=150,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=TRUE,multicore=TRUE,cores=4)
+
+basinPlot(res=figure6a,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main="b",xlab="a",ylab="b")
+basinPlot(res=figure6b,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main="b",xlab="a",ylab="b")
+
+
+

### 2015-05-15 manuscript/figures/figure6.R
removed parameter

@@ -2,7 +2,6 @@ source("../../src/alleeIGC.R")
 source("../../src/basin.R")
 source("../../src/utility.R")
 
-lambda=0.15
 z=0.05
 Am=0.2
 An=0.4

### 2015-05-18 manuscript/figures/figure6.R
pdf files

@@ -11,23 +11,48 @@ r.n=0.05
 r.m=r.n
 x.lim=c(0,1.2)
 y.lim=c(0,1.2)
+    
+cAn=0
+cAm=-0
+cKn=0
+cKm=-0
 
-cAn=0.25
-cAm=-0.25
-cKn=0.25
-cKm=-0.25
+figure60 <- basin(x.lim=x.lim,y.lim=x.lim,points=150,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=TRUE,multicore=TRUE,cores=4)
+
+
+cAn=0.1
+cAm=-0.1
+cKn=0.1
+cKm=-0.1
 
 figure6a <- basin(x.lim=x.lim,y.lim=x.lim,points=150,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=TRUE,multicore=TRUE,cores=4)
 
-cAn=-0.25
-cAm=0.25
-cKn=-0.25
-cKm=0.25
+cAn=-0.1
+cAm=0.1
+cKn=-0.1
+cKm=0.1
 
 figure6b <- basin(x.lim=x.lim,y.lim=x.lim,points=150,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=TRUE,multicore=TRUE,cores=4)
 
+pdf('figure_5a.pdf')    
+basinPlot(res=figure5a,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main="b",xlab="a",ylab="b")
+dev.off()
+pdf('figure_5b.pdf')    
+basinPlot(res=figure5b,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main="b",xlab="a",ylab="b")
+dev.off()
+pdf('figure_5c.pdf')    
+basinPlot(res=figure5c,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main="b",xlab="a",ylab="b")
+dev.off()
+
[…]

### 2015-05-18 experimentMN3/alleeIGC.R
mn3 job

@@ -1,0 +1,59 @@
+
+allee<-function(r.n=0.005,r.m=0.005,ini.m=0.6,ini.n=0.8,Am=0.2,Km=0.7,An=0.5,Kn=1,z=0,timeSteps=2000,b=2,storeFinalOnly=FALSE,cAm=-0.2,cAn=0.2,cKn=0.2,cKm=-0.2)
+{
+# Create Output Vectors:
+    m<-numeric(length=timeSteps)
+    n<-numeric(length=timeSteps)
+    m[1]=ini.m
+    n[1]=ini.n
+
+# Define DELTA
+    DELTA=max(r.n*((Am-Km)^2/(4*Am*Km)),r.m*((An-Kn)^2/(4*An*Kn)))
+# Compute Simulation
+for (t in 2:timeSteps)
+    {
+
+     #intrinsic growth rates
+     Rm=r.m*((m[t-1]-cAn*n[t-1])/Am-1)*(1-(m[t-1]+cKn*n[t-1])/Km)
+     Rn=r.n*((n[t-1]-cAm*m[t-1])/An-1)*(1-(n[t-1]+cKm*m[t-1])/Kn)
+     
+#transmission rates
+
+     if (m[t-1]>0|n[t-1]>0)
+         {
+             if (abs(Rm-Rn)>=(b*DELTA)) {zeta=z}
+             if (abs(Rm-Rn)<(b*DELTA)) {zeta=z*abs(Rm-Rn)/(b*DELTA)}
+         }
+     else
+         {
+             zeta=0
+         }
+     
+#switch function
+     if (Rm==Rn) {C=0} #No transmission
+     if(Rm>Rn) {C=zeta*(n[t-1]+ Rn*n[t-1])} # The proportion moving is a function of the updated population size
+     if (Rn>Rm) {C=-zeta*(m[t-1]+ Rm*m[t-1])} # The proportion moving is a function of the updated population size
+                 
+#transmission(migration)
+# This is essentially population of the previous steps + population increasing through reproduction, and population moving after transmission (post-reproduction).      
+     
+     m[t]=m[t-1]+ Rm*m[t-1] + C
+     n[t]=n[t-1]+ Rn*n[t-1] - C
+
+# This is to avoid unecessary computational time     
+     if((abs(m[t]-m[t-1])==0)&(abs(n[t]-n[t-1])==0))
+         {
+             #print(t)
+             m[t:timeSteps]=m[t]
[…]

### 2015-05-18 experimentMN3/alleeTasks.py

@@ -1,0 +1,22 @@
+#!/usr/bin/python3
+
+from scipy import arange
+
+def main():   
+    competitionValues = [0.0]
+    zValues = [0.05, 0.2]
+    lambdaValues = [0.5, 0.75, 1]
+
+    tasksFile = open('allee.tasks', 'w')
+    numRun = 0
+    for i in competitionValues:
+        for j in zValues:
+            for z in lambdaValues:
+                # params: run competition z lambda
+                tasksFile.write('/apps/R/3.0.2/bin/Rscript run.R '+str(numRun)+' '+str(i)+' '+str(j)+' '+str(z)+'\n')
+                numRun += 1
+    tasksFile.close()
+
+if __name__ == "__main__":
+    main()
+ 

### 2015-05-18 experimentMN3/basin.R

@@ -1,0 +1,140 @@
+basin<-function(x.lim,y.lim,points,multicore=TRUE,r.n,r.m,Am,Km,An,Kn,z,timeSteps,b=2,cores=1,cAn,cAm,cKn,cKm,verbose=TRUE)
+{
+
+
+    x <- seq(from = x.lim[1], to = x.lim[2], length = points)
+    y <- seq(from = y.lim[1], to = y.lim[2], length = points)
+    coordgrid=expand.grid(x,y)
+    res=data.frame(ini.m=coordgrid[,1],ini.n=coordgrid[,2],m.final=NA,n.final=NA)
+    
+    if (verbose==TRUE)
+        {pb <- txtProgressBar(min = 1, max = nrow(res), style=3)}
+
+    if (multicore==TRUE)
+        {
+            require(utils)
+            require(foreach)
+            require(doParallel)
+            registerDoParallel(cores=cores)
+            tmp=foreach(i=1:nrow(res), .combine=rbind) %dopar%
+            {
+                if (verbose==TRUE) {setTxtProgressBar(pb, i)}
+                allee(ini.n=res$ini.n[i],ini.m=res$ini.m[i],
+                      r.n=r.n,r.m=r.m,Am=Am,Km=Km,An=An,Kn=Kn,
+                      z=z,timeSteps=timeSteps,b=b,
+                      cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,
+                      storeFinalOnly=TRUE)}
+            res[,3:4]=tmp
+        }
+
+    if (multicore==FALSE)
+        {
+            for (i in 1:nrow(res))
+                {
+                    if (verbose==TRUE) {setTxtProgressBar(pb, i)}
+                    res[i,3:4]= allee(ini.n=res$ini.n[i],ini.m=res$ini.m[i],
+                           r.n=r.n,r.m=r.m,Am=Am,Km=Km,An=An,Kn=Kn,
+                           z=z,timeSteps=timeSteps,b=b,
+                           cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,
+                           storeFinalOnly=TRUE)
+                }
+        }
+    if (verbose==TRUE) {close(pb)}
+    
+    return(res)
+}
+
+basinPlot<-function(res,Am,Km,An,Kn,cKn,cKm,...)
+    {
+        require(raster)
+
+        #Identify Stable Attractor Node
[…]

### 2015-05-18 experimentMN3/run.R

@@ -1,0 +1,36 @@
+
+source("alleeIGC.R")
+source("basin.R")
+source("utility.R")
+
+args <- commandArgs(trailingOnly = TRUE)
+
+# params run competition z lambda
+run = as.numeric(args[1])
+competition = as.numeric(args[2])
+z = as.numeric(args[3])
+lambda = as.numeric(args[4])
+
+# fixed values
+Am=0.2
+Km=0.8
+r.n=0.05
+r.m=r.n
+x.lim=c(0,1.2)
+y.lim=c(0,1.2)
+
+# exploration
+cAn=competition
+cAm=-competition
+cKn=competition
+cKm=-competition
+diff = Km-Am
+An=Am + (1-lambda)*diff
+Kn=Km + (1-lambda)*diff
+
+result <- basin(x.lim=x.lim,y.lim=x.lim,points=150,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=F,multicore=TRUE,cores=16)
+
+resultsFile <- paste('result_',paste(sprintf('%03d',run, sep=""),'.Rdata', sep=""), sep="")
+print(resultsFile)
+save.image(resultsFile)
+

### 2015-05-18 experimentMN3/utility.R

@@ -1,0 +1,32 @@
+#Utility Functions#
+
+#Calculate Expected m and n stable attractor node for m>0 and n>0 (close to carrying capacity):
+
+ExpStableNode=function(Kn,Km,cKn,cKm)
+    {
+         m=(Km-cKn*Kn)/(1-cKn*cKm)
+         n=(cKm*Km-Kn)/(cKm*cKn-1)
+         return(list(m=m,n=n))
+    }
+
+#Calculate Expected m and n unstable attractor node for m>0 and n>0 (close to Allee Effect):
+
+ExpUnStableNode=function(An,Am,cAn,cAm)
+    {
+         m=(cAn*An+Am)/(1-cAn*cAm)
+         n=(cAm*Am+An)/(1-cAn*cAm)
+         return(list(m=m,n=n))
+    }
+
+
+
+
+#Calculate parameter g, that equalises the max fitness of the two strategies
+EqualiseMaxGrowthRate=function(An,Am,Kn,Km)
+{
+    g=(((Am/Km+1)/2/(Am/Km)-1)*(1-((Am/Km+1)/2)))/
+        (((An/Kn+1)/2/(An/Kn)-1)*(1-((An/Kn+1)/2)))
+    return(g)
+    # Rm=Rn*g to make the two max intrisic growth rates equal
+}
+

### 2015-05-26 manuscript/draft.tex
minor updates

@@
 \begin{abstract}
-Several forms of social learning rules are guided by the amount of payoff received by social teachers adopting specific cultural variants.
-The rationale, is that such payoff will be correlated to the beneficial or detrimental properties of a given variant.
-Here we assume that the payoff directly affects the reproductive fitness of individuals, and that it is determined by the density of individuals possessing the same cultural variant.
-More specifically we consider a  form of density-dependence where a too small or too large population of individuals possessing the same variant will be subject to a decline in fitness, while an intermediate size is optimal.
-This non-linear relationship, known as Allee effect in ecology, portrays a large number of cultural variants, in particular subsistence strategies that are enhanced  by the effect of cooperation but are constrained by interference and limited resources.
-This paper explores the evolutionary dynamics of a two trait model where each variant exhibit Allee effect.
-We used a modified Lotka-Volterra model of payoff-biased social learning, and show that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion to suboptimal variants, long-term co-existence of multiple variants, and fixation to a single optimal or suboptimal strategy.
+Social learning rules are often guided by the amount of payoff received by potential social teachers while expressing specific cultural traits.
+The assumption of the social learner is that such payoff is a proxy for inferring the qualitative features of the target trait.
+Here we explore the theoretical implications when payoff directly affects  reproductive fitness and its expected amount is a function of the number of individuals possessing the given trait.
+More specifically we consider a particular form of density-dependence where a too small or too large population of individuals possessing the same variant will be subject to a decline in fitness, while an intermediate size is optimal.
+This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits, in particular subsistence strategies that are enhanced  by the effect of cooperation but are constrained by interference and limited resources.
+We  explores the evolutionary dynamics of a single trait model  with two variants, each exhibiting a different Allee effect.
+We used a modified Lotka-Volterra model of payoff-biased social learning, and show that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion, stable co-existence, and fixation to one variant.
 
 % XAVI: I'm confused by the use of trait and variant: shouldn't it be 'a single trait model with two variants'?
+% I'm using them as synonym, but I guess you have a point.
+
 \end{abstract}
@@
 
-A key element in all social learning strategies is the availability of some heuristics and learning bias that guides the selection of a given variant over another.
-This information can be based on some properties of the variant itself (\emph{content-bias}) or derived from contextual cues (\emph{context-bias}), such as the rarity or the commonness or the rareness of a trait, or some features associated with the model (i.e.
-the social teacher) such as prestige, success, or the similarity to the learner~\citep{henrich_mcelreath2003}.
+A central element of all social learning strategies is the presence of some heuristics that offers some guidance in the selection of a given variant over another.
+This  can be based on some properties of the variant itself (\emph{content-bias}) or derived from contextual cues (\emph{context-bias}), such as the commonness or the rareness of a trait, or some features associated with the transmitter such as prestige, success, or his/her similarity to the learner~\citep{henrich_mcelreath2003}.
 An expected consequence of \emph{context-bias} is the possibility of a feedback process, whereby the attractiveness of a given variant might change as a consequence of the learning strategy~\citep{kendal_etal_2009}.
 For example a strong anti-conformist bias might initially favour rare traits, but as they spread into a population their selective advantage will decline as the result of a negative feedback.
-In a similar, but in a opposite direction, both conformist-bias and homophily-guided learning strategies would exercise a positive feedback loop.
-%citation needed?
+On the other hand conformist-bias and homophily-guided learning strategies are expected to exercise a positive feedback to some form of fixation.
 
 This paper examines a group of learning strategies generally referred to as \emph{payoff-based transmission}~\citep{schlag1998,kendal_etal_2009,lake_and_crema_2012,baldini2013,kandler_and_laland_2013,crema_lake_inpress}.
-The core assumption shared by these models is that the probability of copying a given cultural variant is positively correlated with some payoff signal expressed by individuals possessing such a variant.
+The core assumption shared by these models is that the probability of copying a given cultural trait $i$ is positively correlated with a payoff signal $s$, determined by some function $f(i)$ expressed by individuals possessing the trait itself.
 Several variants of this learning strategies have been proposed and explored, including: \emph{imitate the best}, where the social learner identifies and copies the trait possessed by the individual with the highest payoff; \emph{compare means}, where the social learner compare the average payoff of different strategies and chooses the highest variant; and the \emph{proportional imitation rule}, where the probability of adopting a variant is proportional to the difference in payoff between the demonstrator and the learner~\citep{schlag1998,baldini2013}.
@@
 Given the same cultural trait, the variability in the payoff signal is partly derived by the presence of interacting cultural and physical traits, as well as the environment where the trait is manifest.
+Thus the payoff is better described by the function $f(i,\epsilon)$, where $i$ is the trait, and $\epsilon$ is a series of additional hidden or latent variables that might or might interact with $i$.
 For example, a fisherman might evaluate the performance of a hook (the target cultural trait) using as a cue the number of fish captured (the payoff signal).
@@
 In many cases we should expect that the social teacher and learner share many of these interacting traits as well as the environment where these are displayed.
-In other words, cultural and ecological inheritance can generally reduce the uncertainty and the variability in the payoff-signal.
+In other words, cultural and ecological inheritance can generally reduce the impact of $\epsilon$, and consequently the uncertainty and the variability in the payoff-signal.
 There are however exceptions.
-One of these are contexts where the success of a variant is strongly correlated with the absolute number of individuals possessing the same trait within a population.
-This density dependence implies the adoption (or abandonment) of a trait can alter the expected payoff-signal, possibly generating a variety of feedback mechanism.
+One of these are contexts where the success of a variant is strongly correlated with the absolute number of individuals $n_i$ possessing the same trait within a population.
+This density dependence, which can be described by the payoff function $f(i,n_i)$, implies that the adoption (or abandonment) of a trait can alter the expected payoff-signal, possibly generating a variety of feedback mechanism.
 
@@
 In particular, behaviour linked to the exploitation of limited resources are expected to exhibit a negative density dependence.
-Novel optimal variants are likely to spread rapidly when the trait is rare, but once a threshold frequency is exceeded,  beneficial effect will start to decline, the payoff will become smaller, and eventually rarer suboptimal variants can be preferred (see Lake and Crema 2012 for an extensive analysis).
-Density dependence can also have a positive direction, whereby an increase in the number of individuals possessing a trait might have a selective advantage additional to what is already expected from its frequency.
+Novel optimal variants are likely to spread rapidly when the trait is rare, but once a threshold frequency is exceeded (i.e.
+the carrying capacity),  beneficial effect will start to decline, the payoff will become smaller, and eventually rarer suboptimal variants can be preferred (see Lake and Crema 2012 for an extensive analysis).
+
+However, density dependence does not necessarily imply only a negative correlation between population size and payoff/fitness.
+In many ecological contexts an inverse density dependence, whereby per capita growth rate is positively correlated with population density, has been noted.
+This phenomena, known as Allee effect \citep{allee1958,courchamp_etal_1999}, is generally explained by a range of processes from cooperation to unconscious mutual facilitation that enhances individual fitness at low population density.
+This inverse density dependence can be further distinguished between \emph{weak} and \emph{strong} Allee effect.
+In the former case, the reproductive fitness is always positive at low population density, whilst in the latter case we expect the presence of a critical density below which fitness becomes negative \citep{Wang_and_Kot_2001}.
+
 Variants that are enhanced by some form of cooperation, or more in general traits exhibiting a positive niche construction are expected to show such a pattern, known in ecology as \emph{Allee Effect}~\citep{allee1958}.
+
+While a systematic review of the empirical evidence of Allee effect has been conducted in population ecology \citep{kramer_etal_2009}, no
+
 For example foraging groups are known to show higher performance when the number of individuals is not too small~\citep{hill_and_hawkes_1983,janssen_and_hill_2014}, while certain activities such as anthropogenic fire~\citep{bird2013} or selective hunting~\citep{dods_2002} might require a population density beyond a certain threshold in order to promote sufficient niche enhancement and a consequent increase in payoff~\citep{rowley-conwy_and_layton_2011}.
@@
 
+%We definitely need to add some literature review on the theorethical models of Allee effect here as an additional paragraph.
+
 Here we explore the combined effect of positive and negative density-dependent selection, assuming that: 1) the social learning strategy is payoff-based transmission; and 2)  payoff is equivalent to the reproductive fitness of each individual.
-We will demonstrate, by exploring a two-trait difference equation, that the degree by which a population relies on social learning can strongly affect the evolutionary trajectory of the system, leading to the emergence of stable and unstable equilibria, as well as the occurrence of temporary or permanent episodes of reversion to suboptimal variants.
-% XAVI not sure if social learning or allee is more important...what about: "we will demonstrate...that the impact of Allee effect on social learning can strongly affect the evolutionary trajectory of the system
+We will demonstrate, by exploring a two-trait difference equation, that the impact of Allee effect on social learning can strongly affect the evolutionary trajectory of the system leading to the emergence of stable and unstable equilibria, as well as the occurrence of temporary or permanent episodes of reversion to suboptimal variants.
 
@@
 
+%Mention that we need more empirical evidence of Allee effect
+
 %General Intro on Human History%

### 2015-05-28 manuscript/draft.tex

@@
 Several cultural traits are expected to show such density dependence.
-In particular, behaviour linked to the exploitation of limited resources are expected to exhibit a negative density dependence.
+In particular, behaviour linked to the exploitation of limited resources are expected to exhibit a inverse density dependence.
 Novel optimal variants are likely to spread rapidly when the trait is rare, but once a threshold frequency is exceeded (i.e.
-the carrying capacity),  beneficial effect will start to decline, the payoff will become smaller, and eventually rarer suboptimal variants can be preferred (see Lake and Crema 2012 for an extensive analysis).
+the carrying capacity),  beneficial effect will start to decline, the payoff will become smaller, and eventually rarer suboptimal variants can be preferred (see Lake and Crema \citeyear{lake_and_crema_2012} for an extensive analysis).
 
@@
 In the former case, the reproductive fitness is always positive at low population density, whilst in the latter case we expect the presence of a critical density below which fitness becomes negative \citep{Wang_and_Kot_2001}.
+The implications of direct and inverse density dependence has been extensively studied both from the empirical and theoretical perspectives, including its effect of habitat selection \citep{greene_and_stamps_2001}, resource competition \citep{jang2013}, and dispersal \citep{steele_2009}, as well as specific events such as shifts in settlement pattern \citep{crema_2014}, colonisation events,  \citep{kennet_etal_2006}, genetic consequences on expanding population \citep{roques_etal_2012} and extinction of languages \citep{sutherland_2003}
 
-Variants that are enhanced by some form of cooperation, or more in general traits exhibiting a positive niche construction are expected to show such a pattern, known in ecology as \emph{Allee Effect}~\citep{allee1958}.
+Although a systematic review and a shared theoretical framework akin to those proposed in population ecology \citep{kramer_etal_2009} is not present in the social sciences, these studies suggest that there are several lines of evidence supporting the existence of both direct and inverse density dependence in cultural and behavioural traits.
+For example several ethnographic studies on small scale societies show how the per-capita return of hunting groups show higher yield when the group size is not too small nor too big~\citep{hill_and_hawkes_1983,janssen_and_hill_2014}.
+Other instances of inverse density dependence include population level activities such as the use of anthropogenic fire~\citep{bird2013} or selective hunting~\citep{dods_2002} which often require a population size above a certain critical density in order to promote sufficient niche enhancement and a consequent increase in payoff~\citep{rowley-conwy_and_layton_2011}.
+In a broader sense any cultural trait that benefits from direct or indirect cooperation \citep{}, promote positive niche construction \citep{vandermeer_2008}, and requires a critical mass \citep{rogers_2003} are likely to exhibit some degree of Allee effect.
+This is thus not limited necessarily to the domain of traits directly promoting reproductive fitness, and hence can include language~\citep{kandler2009}, organisational populations~\citep{caroll_and_hannan_1989} and communication technologies \citep{van_slyke_perceived_2007}.
 
-While a systematic review of the empirical evidence of Allee effect has been conducted in population ecology \citep{kramer_etal_2009}, no
-
-For example foraging groups are known to show higher performance when the number of individuals is not too small~\citep{hill_and_hawkes_1983,janssen_and_hill_2014}, while certain activities such as anthropogenic fire~\citep{bird2013} or selective hunting~\citep{dods_2002} might require a population density beyond a certain threshold in order to promote sufficient niche enhancement and a consequent increase in payoff~\citep{rowley-conwy_and_layton_2011}.
-Other traits that might be promoted by increased density includes language~\citep{kandler2009}, organisational populations~\citep{caroll_and_hannan_1989} and communication technologies~\citep{van_slyke_perceived_2007}.
-
-%We definitely need to add some literature review on the theorethical models of Allee effect here as an additional paragraph.
-
-Here we explore the combined effect of positive and negative density-dependent selection, assuming that: 1) the social learning strategy is payoff-based transmission; and 2)  payoff is equivalent to the reproductive fitness of each individual.
-We will demonstrate, by exploring a two-trait difference equation, that the impact of Allee effect on social learning can strongly affect the evolutionary trajectory of the system leading to the emergence of stable and unstable equilibria, as well as the occurrence of temporary or permanent episodes of reversion to suboptimal variants.
+Here we explore the consequences of Allee effect in the evolutionary dynamics of cultural transmission, assuming that: 1) the social learning strategy is payoff-based; and 2)  the payoff is equivalent to the reproductive fitness of each individual.
+We will demonstrate, by means of a difference equation, that direct and inverse density dependence can strongly affect the evolutionary trajectory of the system leading to the emergence of stable and unstable equilibria, as well as the occurrence of temporary or permanent episodes of reversion to suboptimal cultural variants.
 
 \section{The Model}
-We consider the evolutionary dynamics of two mutually exclusive subsistence strategies whose fitness has both positive and negative density dependence.
-In particular we examine how two population of individuals, $a$ and $b$, change over time if we allow for a social learning strategy guided by a \emph{proportional imitation rule} (i.e.
-the probability of an individual with lower payoff to adopt the trait of an individual with higher payoff is proportional to the difference in the payoff).
-The model can theoretically be extended to $n$ strategies, but here we decide to explore a two trait version for simplicity.
-The core model can be depicted with the coupled difference equation \eqref{eq1}:
+We consider the evolutionary dynamics of two mutually exclusive cultural variants whose payoff function is characterised by both direct and inverse density dependence.
+We assume a non-fixed population of size $N$, subdivided into two subpopulations with sizes $a$ and $b$, each associated with the two cultural variants.
+We further assume that $a$ and $b$ can change as a result of intrinsic population growth/decline as well as shifts from one to the other as a result of cultural transmission based.
+The core model can be depicted with the following coupled difference equation \eqref{eq1}:
 
@@
 % XAVI do you want to use cultural trasmission here or better stick with social learning?
+
 The reproductive growth rate $R$ of each population is defined by the following paired equation:
@@
 
-where $r$ is the basic growth rate, $A$ is the \emph{Allee} threshold and $K$ is the carrying capacity, following the condition $A < K$.
-Equation \eqref{eq2} implies that the reproductive rate is negative when density is below $A$ or above $K$, and will reach its maxima $R_{max}=\frac{r(A-K)2}{4AK}$ when the population density is equal to $\frac{(K-A)}{2}$.
+where $r$ is the basic growth rate, $A$ is the critical density and $K$ is the carrying capacity, following the condition $0<A < K$.
+Equation \eqref{eq2} depicts a \emph{strong} Allee effect, hence the reproductive rate is negative when density is below $A$ or above $K$, and will reach its maxima $R_{max}=\frac{r(A-K)2}{4AK}$ when the population density is equal to $\frac{(K-A)}{2}$.
 
-Cultural transmission is portrayed in \eqref{eq3}:
+Cultural transmission is based on the \emph{proportional imitation rule} \citep{schlag1998}, and hence in this case the probability of an individual with lower growth rate to adopt the trait of an individual with higher growth rate is proportional to the difference in the growth rates.
+This is portrayed in \eqref{eq3}:
 
@@
 
-% XAVI b is not explained?
+% XAVI b is not explained? %It's no longer b but m
 where $z$ is a measure of reliance in social learning (effectively equivalent to highest possible transmission rate), $\Delta$ is the maximum between $R_{max(a,t)}$ and $R_{max(b,t)}$, and $m$ is a calibration parameter that measure the perception of the difference in the payoffs (i.e.
 small values of $m$ determines a higher rate by which the theoretical maximum transmission rate is reached).
-Thus \eqref{eq3} and \eqref{eq4}  conforms with the \emph{proportional imitation rule}~\citep{schlag1998}, although here the payoff is equivalent to the reproduction rate.
 
@@
 
+The results
+
 %Mention that we need more empirical evidence of Allee effect

### 2015-05-28 manuscript/figures/figure6.R
final figure

@@ -45,52 +45,23 @@ getPercentages<- function(raw)
     return(percentages)
 }
 
-getRelativeValue <- function(row, z0)
-{
-    z0Value <- subset(z0, lambda==row$lambda & eq==row$eq)$value
-    return(1+(row$value-z0Value)/z0Value)
-}
-
-
 raw <- data.frame()
-for(run in 0:109)
+for(run in 0:65)
 {
     raw <- rbind(raw, getRawValues(run))
 }
 
 percentages <- getPercentages(raw)
 
+plotted <- subset(percentages, lambda>=0.9 | lambda<=0.6)
+plotted$lambda <- factor(plotted$lambda, levels=c(1,0.95,0.9,0.6,0.55,0.5))    
+
 # coexistence: #006400
-# unstable: #D3D3D3
+# extinct: #D3D3D3
 # nExtinct: #4169E1
 # mExtinct: #CD5C5C
 
-pdf('figure6a.pdf', width=12, height=7)
-ggplot(subset(percentages, z<0.6), aes(x=lambda, y=value, fill=eq)) + geom_area(stat='identity') + scale_fill_manual(name='Equilibria', values=c('#006400','gold2','#4169E1','#CD5C5C','#D3D3D3'))+ facet_wrap(~z, nrow=2) + scale_x_reverse()
+pdf('figure6.pdf', width=10, height=6)
+ggplot(plotted, aes(x=z, y=value, fill=eq)) + geom_area(stat='identity') + scale_fill_manual(name='Equilibria', values=c('#006400','#D3D3D3','#4169E1','#CD5C5C','gold2'))+ facet_wrap(~lambda, nrow=2)
 dev.off()
-
-
-pdf('figure6b.pdf', width=15, height=7)
-ggplot(subset(percentages, z<0.6), aes(x=z, y=value, fill=eq)) + geom_area(stat='identity') + scale_fill_manual(name='Equilibria', values=c('#006400','gold2','#4169E1','#CD5C5C','#D3D3D3'))+ facet_wrap(~lambda, nrow=2)
-dev.off()
-
-pdf('figure6c.pdf', width=18, height=7)    
[…]

### 2015-05-28 experimentMN3/alleeTasks.py
updated overlap figure

@@ -3,23 +3,30 @@
 from scipy import arange
 
 def main():
-    # 1 - exploration lambda/z
     """
+    # 1 - exploration lambda/z
     competitionValues = [0.0]
     zValues = [0.05, 0.2]
     lambdaValues = [0.5, 0.75, 1]
     """
-    """
     # 2 - exploration lambda/z (relative figure)
     competitionValues = [0.0]
     zValues = arange(0,1.1,0.1)
-    lambdaValues = arange(0.5,1,0.05)
-    """
+    lambdaValues = arange(0.5,1.1,0.05)
 
+    """
     # 3 - exploration competition/z 
     competitionValues = [-0.1,0.1]
     zValues = [0.05, 0.2]
     lambdaValues = [0.75]
+    """
+
+    """
+    # 4 - z (figure 3)
+    competitionValues = [0.0]
+    zValues = [0,0.05]
+    lambdaValues = [1/3]
+    """
 
     tasksFile = open('allee.tasks', 'w')
     numRun = 0

### 2015-05-28 experimentMN3/run.R

@@ -28,7 +28,7 @@ diff = Km-Am
 An=Am + (1-lambda)*diff
 Kn=Km + (1-lambda)*diff
 
-result <- basin(x.lim=x.lim,y.lim=x.lim,points=150,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=F,multicore=TRUE,cores=16)
+result <- basin(x.lim=x.lim,y.lim=x.lim,points=300,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=F,multicore=TRUE,cores=16)
 
 resultsFile <- paste('result_',paste(sprintf('%03d',run, sep=""),'.Rdata', sep=""), sep="")
 print(resultsFile)

### 2015-05-29 manuscript/draft.tex
updated tex

@@
 
+% XAVI put high res figure
 \begin{figure}[h!]
@@
 The dynamics of two strategies are similar for low values of $\lambda$.
-High values indicate that the minimum number of individuals needed to achive positive fitness in $b$ gets closer to $K_a$, while the population size that $b$ can sustain also grows.
+High values indicate that the minimum number of individuals needed to achieve positive fitness in $b$ gets closer to $K_a$, while the population size that $b$ can sustain also grows.
 As a consequence, the number of individuals switching strategies when $b$ has lower payoff is also higher and could potentially overload $a$ due to its lower $K_a$.
 
-Figure XX shows how the increase in $\lambda$ affects the equilibria of our model.
+Figure \ref{fig:overlap} shows how the degree of overlap affects the equilibria of our model with and without social learning.
+The relation between the two processes strongly affects the shape of the different basins in a non-dynamic fashion.
 
-% XAVI add results and figure
+% XAVI put high res figure, and update left, middle, right order
+
+\begin{figure}[h!]
+  \centering
+      \includegraphics[width=\textwidth]{./figures/figure5}
+  \caption{Basins of attraction for $z=0.05$ (top) and $z=0.2$ (bottom); $\lambda=1$ (left), $\lambda=0.75$ (middle) and $\lambda=0.5$}
+    \label{fig:overlap}
+\end{figure}
+
+Figure \ref{fig:percentages} explores in deeper detail the effects of both parameters ($z$ and $lambda$) on the basin of each equilibria.
+An increase in $z$ decreases the possibility of having coexistence for high similarity between strategies $a$ and $b$ (i.e.
+$\lambda \geq 0.9$.
+From $0.9 < \lambda < 0.55$ the dynamics remain remarkably similar, while for $\lambda = 0.5$ the effects of social learning decreases because of the large gap between $A$ and $K$ of both strategies.
+
+\begin{figure}[h!]
+  \centering
+      \includegraphics[width=\textwidth]{./figures/figure6}
+  \caption{Percentages of the entire space falling in each basin of attraction for different values of $\lambda$. X axis controls the parameter $z$ defining social learning.}
+    \label{fig:percentages}
+\end{figure}
 
@@
 
-Figure XX shows a scenario where strategy $a$ is dominant over strategy $b$.
-% XAVI add results and figure
+Figure \ref{fig:competition} shows a scenario where strategy $a$ is dominant over strategy $b$.
+
+\begin{figure}[h!]
+  \centering
+      \includegraphics[width=\textwidth]{./figures/figure7}
+  \caption{The impact of competition on the basins of attraction for the different equilibria.
+The Parameters used are $z=0.05$ (left) and $z=0.2$ (right) for social learning, and $c=-0.1$ (top) and $c=0.1$ (bottom) for competition.}
+    \label{fig:competition}
+\end{figure}
+
+% XAVI some brief discussion on results here, and update the dashed lines
 

### 2015-05-29 manuscript/draft.tex
Commit for working from home

--- 2015-05-28 manuscript/draft.tex
@@
-\documentclass[preprint,authoryear]{elsarticle}
+%\documentclass[review,authoryear]{elsarticle}
+ \documentclass[3p,authoryear,twocolumn]{elsarticle} %Just to give a feel of how it would look like
 
@@
 Other instances of inverse density dependence include population level activities such as the use of anthropogenic fire~\citep{bird2013} or selective hunting~\citep{dods_2002} which often require a population size above a certain critical density in order to promote sufficient niche enhancement and a consequent increase in payoff~\citep{rowley-conwy_and_layton_2011}.
-In a broader sense any cultural trait that benefits from direct or indirect cooperation \citep{}, promote positive niche construction \citep{vandermeer_2008}, and requires a critical mass \citep{rogers_2003} are likely to exhibit some degree of Allee effect.
+In a broader sense any cultural trait that benefits from direct or indirect cooperation %Add reference \citep{},
+promote positive niche construction \citep{vandermeer_2008}, and requires a critical mass \citep{rogers_2003} are likely to exhibit some degree of Allee effect.
 This is thus not limited necessarily to the domain of traits directly promoting reproductive fitness, and hence can include language~\citep{kandler2009}, organisational populations~\citep{caroll_and_hannan_1989} and communication technologies \citep{van_slyke_perceived_2007}.
@@
   \centering
-      \includegraphics[width=0.7\textwidth]{./figures/figure1.jpg}
+      \includegraphics[width=0.5\textwidth]{./figures/figure1.jpg}
   \caption{Bifurcation diagram of a single population model with $A_x=0.2$ and $K_x=0.8$. For $0<r_x<0.6649747$, any initial value of $x$ greater than $A_x$ converge to $K_x$.}
@@
   \centering
-      \includegraphics[width=0.7\textwidth]{./figures/figure2.pdf}
+      \includegraphics[width=0.5\textwidth]{./figures/figure2.pdf}
   \caption{Basins of attraction for the four equilibria (mutual extinction: grey; extinction of $a$: blue; extinction of $b$: red; stable coexistence: green).
@@
   \centering
-      \includegraphics[width=\textwidth]{./figures/figure3.pdf}
+      \includegraphics[width=0.5\textwidth]{./figures/figure3.pdf}
   \caption{Basins of attraction for the four equilibria (mutual extinction: grey; extinction of $a$: blue; extinction of $b$: red; stable coexistence: green) with (right) and without (left) social learning.
@@
   \centering
-      \includegraphics[width=\textwidth]{./figures/figure4.jpg}
+      \includegraphics[width=0.5\textwidth]{./figures/figure4.jpg}
   \caption{Bifurcation diagram for $a$ (red) and $b$ (blue) as a function of social learning ($z$): left: $a=0.11$, $b=0.4$; Middle: $a=0.5$, $b=0.21$, $a=0.4$, $b=0.4$.
@@
 \end{figure}
+
+%ENRICO: Add here basin plot with chaotic system and show tiny difference in initial condition can lead to stable coexistence or chaotic system
 
@@
 
-The results
+% Cooperation / Facilitation / and Niche Construction review ?
 
-%Mention that we need more empirical evidence of Allee effect
+Our analysis examined the impact of payoff-based social learning strategies in the evolutionary dynamics of cultural traits with direct and inverse density dependence.
+The results showed that social learning can alter the basins of attraction of stable nodes, leading to different dynamics and equilibria from the very same initial conditions.
+When the reliance on social learning is comparatively high, and  the two variants have a substantial overlap in their range of positive payoff (i.e.
+they share similar critical densities and carrying capacities), we also observe the emergence of a chaotic behaviour where the system fluctuates without reaching a stable equilibria.
+Our results imply that evolutionary trajectories that are generally expected from a standard two population Allee models can exhibit unexpected outcomes.
+Variants below critical density can survive and a stable coexistence of strategies above this threshold is not necessarily guaranteed.
+The effects of social learning can also be observed in the trajectories of these evolutionary dynamics .
+Reversions to lower densities do not exclusively occur only when the variant is above its carrying capacity, and when the reliance on social learning is high these episodes becomes more frequent and extreme in its magnitude.
 
-%General Intro on Human History%
-Major transitions in human history have been marked by the fast spread of key innovations, which laid the foundations of further cumulative development in economy, society, and culture.
-Shifts in subsistence economy has often been regarded as a prime example of these transitions, triggered by the spread of a suite of behavioural traits such as the know-how of the domestication process~\citep{barker2006} the use of new subsistence tools~\citep{petraglia_population_2009} and processing techniques~\citep{molleson1993}, and the inclusion/exclusion and extensification/intensification of different prey species %citation needed.
-Archaeological evidence does not support a simple progressive adoption of these key cultural variants, and instead a showcase a variety of trajectories, including the long-term coexistence of multiple subsistence strategies, or even reversions to older traits~\citep{rowley2001}.
-  %add long list of citations. or perhaps one or two concrete examples (probably better)
-\enrico{General Intro on HBE? Not sure if needed for TPB? Maybe move to discussion?}
-Formal models portraying changes in human subsistence have been mostly offered in human behavioral ecology~\citep{smith1992,bird2006,kennett2006}, where several mathematical models inspired from evolutionary ecology %we need a better one-line description of HBE.
- used to predict  based on expected patterns under the assumption of optimality~\citep{belovsky1988}. 
-While earliest models relied almost exclusively on this assumption, most recent works are increasing discussing how the knowledge of alternative subsistence strategies are acquired %citation needed?
-Are there such works?
-Need to check
-and whether this process of learning can drive the evolutionary trajectory of a given population~\citep{henrich1998}, partly as a synergy with the dual inheritance theory~\citep{boyd1985}.
+%Not sure where to put competition, maybe the whole thing should be in the appendix?
 
-The results of these works often exhibits counter-intuitive outcomes that cannot be predicted by standard assumptions of optimality.
-For instance, extreme reliance on social learning can promote information parasitism~\citep{giraldeau2002}, leading to a decline in the proportion of asocial learners that in certain environmental context can have strongly detrimental effects at the population level~\citep{whitehead2009}.
-Other examples include instances where a reduced effective population size promotes an increase in random drift, and the ultimate loss of adaptive traits~\citep{henrich2004}, or instances where neutral traits are jointly transmitted together with adaptive ones~\citep{ackland2007}.
-These examples illustrate how theoretical models based on the dual inheritance theory can extend predictions of the standard optimal foraging theory, integrating dynamics that, at least at a first glance, are regarded as counter-intuitive.
-
-These include episodes of reversions (e.g.
-\footnote{citation needed}), where a more recent optimal strategies are abandoned for older suboptimal variants; stable coexistence (e.g.
-\footnote{citation needed}), where optimal and suboptimal strategies co-exist within a population; and even the complete loss of optimal variants~\citep[e.g.][]{henrich2004}.
-The evolutionary processes leading to these different outcomes are undoubtedly contingent to local historical dynamics, nonetheless we argue that a simple model of social learning in combination with a positive and negative frequency dependent selection can lead all these equilibria.
-
-\enrico{We need to mention that some behaviour might actually enhance the other community (or be detrimental).
-This can be modelled using a competition coefficient as in~\citep{jang2013}.
-I don't think we need to explore this but it is definitely worth mentioning!}
-
-\enrico{Mutually exclusive trait, but does not necessarily imply HG vs FA mixture of strategy (say 50\%HG vs 50\%F) can be assumed to be a trait.
-Perhaps this needs to be mentioned earlier when the model is introduced.}
+The implications of our model can provide some insights in the evolutionary history of subsistence traits in human history.
+Many prime examples of keystone innovations are closely related to the introduction of new subsistence strategies that enabled a larger carrying capacity, often coupled with the cost of higher critical densities.
+The spread of behavioural traits such as the know-how of the domestication process~\citep{barker2006} the use of new subsistence tools~\citep{petraglia_population_2009} and processing techniques~\citep{molleson1993}, as well as the inclusion/exclusion and extensification/intensification of different prey species are just some of the most notable examples.
+Both ethnographic and archaeological evidence show many instances where these critical transitions where not smooth, often leading to the coexistence of alternative strategies as well as episodes of a return to a previously abandoned strategies.
+%PUT HERE LIST of EXAMPLES.
+While in most cases these are more anecdotal evidence than testable empirical cases, the insights of our models suggests that different levels of reliance on social learning could in fact profoundly drive the system into different trajectories.
+These insights can possibly contribute to existing models of subsistence shifts proposed in human behavioural ecology\citep{smith1992,bird2006,kennett2006}, combining its core theoretical framework to the one developed by cultural evolution studies.
 

### 2015-05-29 manuscript/figures/figure3.R
updated figures and data

@@ -2,49 +2,16 @@ source("../../src/alleeIGC.R")
 source("../../src/basin.R")
 source("../../src/utility.R")
 
-                                        #Figure 3
-lambda=0.15
-z=0.05
-Am=0.2
-An=0.4
-Km=0.8
-Kn=1
-cAn=0
-cAm=0
-cKn=0
-cKm=0
-r.n=0.05
-r.m=r.n
-x.lim=c(0,1.2)
-y.lim=c(0,1.2)
-
-
-z=0
-
-figure3a=basin(x.lim=x.lim,y.lim=x.lim,points=41,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=TRUE,multicore=TRUE,cores=5)
-
-z=0.05
-
-figure3b=basin(x.lim=x.lim,y.lim=x.lim,points=150,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=TRUE,multicore=TRUE,cores=5)
-
-save.image("figure3.RData")
-
-
-load("figure3.RData")
-source("../../src/alleeIGC.R")
-source("../../src/basin.R")
-source("../../src/utility.R")
-
-
-pdf('figure3.pdf', width=14, height=8)
-
+pdf('figure3.pdf', width=16, height=8)
 par(mfrow=c(1,2))
-basinPlot(res=figure3a,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main="z=0",xlab="a",ylab="b")
+
+load('../../experimentMN3/00_fig3/result_000.Rdata')
+title=substitute(paste(z == i), list(i=z))
+basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b")
 abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
[…]

### 2015-05-29 manuscript/figures/figure5.R

@@ -5,13 +5,13 @@ source("../../src/utility.R")
 
 pdf('figure5.pdf', width=15, height=10)
 par(mfrow=c(2,3))
-load('../../experimentMN3/01_lambda/result_000.Rdata')
+load('../../experimentMN3/01_lambda/result_002.Rdata')
 title=substitute(paste(z == i, "  ", lambda == j), list(i=z,j=lambda))
 basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b")
 abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
 axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
-axis(side=4,at=c(An,Kn),labels=c(expression(Ab),expression(Kb)),hadj=0.5,las=2)
+axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
 
 load('../../experimentMN3/01_lambda/result_001.Rdata')
 title=substitute(paste(z == i, "  ", lambda == j), list(i=z,j=lambda))
@@ -19,31 +19,15 @@ basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a"
 abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
 axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
-axis(side=4,at=c(An,Kn),labels=c(expression(Ab),expression(Kb)),hadj=0.5,las=2)
+axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
 
-load('../../experimentMN3/01_lambda/result_002.Rdata')
+load('../../experimentMN3/01_lambda/result_000.Rdata')
 title=substitute(paste(z == i, "  ", lambda == j), list(i=z,j=lambda))
 basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b")
 abline(h=c(An,Kn),lty=2,col="white")
[…]

### 2015-05-29 manuscript/figures/figure7.R

@@ -11,7 +11,7 @@ basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a"
 abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
 axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
-axis(side=4,at=c(An,Kn),labels=c(expression(Ab),expression(Kb)),hadj=0.5,las=2)
+axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
 
 load('../../experimentMN3/03_competition/result_001.Rdata')
 title=substitute(paste(z == i, " ", c== j), list(i=z,j=competition))
@@ -19,7 +19,7 @@ basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a"
 abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
 axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
-axis(side=4,at=c(An,Kn),labels=c(expression(Ab),expression(Kb)),hadj=0.5,las=2)
+axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
 
 load('../../experimentMN3/03_competition/result_002.Rdata')
 title=substitute(paste(z == i, " ", c== j), list(i=z,j=competition))
@@ -27,7 +27,7 @@ basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a"
 abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
 axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
-axis(side=4,at=c(An,Kn),labels=c(expression(Ab),expression(Kb)),hadj=0.5,las=2)
+axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
 
[…]

### 2015-05-29 manuscript/figures/figureUnstable.R
new figue

@@ -1,0 +1,45 @@
+
+library(ggplot2)
+
+source("../../src/alleeIGC.R")
+source("../../src/basin.R")
+source("../../src/utility.R")
+
+
+# load lambda=1, z=0.5
+#load('../../experimentMN3/02_relative_lambda/result_065.Rdata')
+# or load lambda=1, z=0.4
+load('../../experimentMN3/02_relative_lambda/result_054.Rdata')
+
+pdf('figureUnstable1.pdf', width=8, height=8)
+
+basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main="",xlab="a",ylab="b")
+abline(h=c(An,Kn),lty=2,col="white")
+abline(v=c(Am,Km),lty=2,col="white")
+axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
+axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
+points(x=c(0.54,0.56), y=c(0.8, 0.8), pch=20)
+
+# text with buffer    
+text(0.52,0.8,labels="j",col='black',font=2, cex=1)
+text(0.58,0.8,labels="k",col='black',font=2, cex=1)
+
+dev.off()
+
+stable <- allee(ini.m=0.55,ini.n=0.8,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
+unstable <- allee(ini.m=0.56,ini.n=0.8,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
+
+stableA <- data.frame(step=1:length(stable$m), pop=stable$m, strategy='a', equilibrium='j')
+stableB <- data.frame(step=1:length(stable$n), pop=stable$n, strategy='b', equilibrium='j')
+unstableA <- data.frame(step=1:length(unstable$m), pop=unstable$m, strategy='a', equilibrium='k')
+unstableB <- data.frame(step=1:length(unstable$n), pop=unstable$n, strategy='b', equilibrium='k')
+
+trajectories <- rbind(stableA, stableB, unstableA, unstableB)
+
+pdf('figureUnstable2.pdf', width=10, height=4)
+
[…]

### 2015-06-01 manuscript/figures/figureUnstable.R
unstable figure

@@ -5,11 +5,8 @@ source("../../src/alleeIGC.R")
 source("../../src/basin.R")
 source("../../src/utility.R")
 
-
-# load lambda=1, z=0.5
-#load('../../experimentMN3/02_relative_lambda/result_065.Rdata')
-# or load lambda=1, z=0.4
-load('../../experimentMN3/02_relative_lambda/result_054.Rdata')
+# lambda=0.9, z=0.4
+load('../../experimentMN3/05_unstability/result_003.Rdata')
 
 pdf('figureUnstable1.pdf', width=8, height=8)
 
@@ -18,16 +15,16 @@ abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
 axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
 axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
-points(x=c(0.54,0.56), y=c(0.8, 0.8), pch=20)
+points(x=c(0.4,0.39), y=c(1.0, 0.99), pch=20)
 
 # text with buffer    
-text(0.52,0.8,labels="j",col='black',font=2, cex=1)
-text(0.58,0.8,labels="k",col='black',font=2, cex=1)
+text(0.41, 1.01,labels="k",col='black',font=2, cex=1)
+text(0.38,0.98,labels="j",col='black',font=2, cex=1)
 
 dev.off()
 
-stable <- allee(ini.m=0.55,ini.n=0.8,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
-unstable <- allee(ini.m=0.56,ini.n=0.8,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
+stable <- allee(ini.m=0.39,ini.n=0.99,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
[…]

### 2015-06-02 manuscript/draft.tex
unstable figures added

--- 2015-05-29 manuscript/draft.tex
@@
 \usepackage{amsmath}
+\usepackage{subcaption}
 
@@
 
+Small differences in initial conditions can have major consequences in the boundaries between the unstable region and the rest of equilibria.
+Figure \ref{fig:unstable} shows an example with high value of social learning $z=0.4$, where two scenarios ($j$ and $k$) generate different outcomes from very close initial conditions.
+
+\begin{figure}[h!]
+  \centering
+      \includegraphics[width=\textwidth]{./figures/figureUnstable}   
+      \caption{Basin plot (left) and time series (right) for different initial conditions.
+Scenario $j$ reaches \emph{coexistence} equilibrium with initial conditions $a_0=0.39; b_0=0.99$.
+Scenario $k$ follows a limit cycle after 1000 time-steps with initial conditions $a_0=0.4,b_0=1.0$.
+This pattern is reproduced inside the orange zone of the basin plot.}
+    \label{fig:unstable}
+\end{figure}
+
 \subsection{Effects of strategy overlap}

### 2015-06-02 manuscript/figures/figureUnstable.R

@@ -33,9 +33,9 @@ unstableB <- data.frame(step=1:length(unstable$n), pop=unstable$n, strategy='b',
 
 trajectories <- rbind(stableA, stableB, unstableA, unstableB)
 
-pdf('figureUnstable2.pdf', width=10, height=4)
+pdf('figureUnstable2.pdf', width=6, height=6)
 
-ggplot(subset(trajectories, step<1501), aes(x=step, y=pop, col=strategy)) + geom_line() + scale_color_manual(values=c('red','blue')) + facet_wrap(~equilibrium) + theme(legend.position='none') + xlab('time (steps)') + ylab('')
+ggplot(subset(trajectories, step<1501), aes(x=step, y=pop, col=strategy)) + geom_line() + scale_color_manual(values=c('red','blue')) + facet_wrap(~equilibrium, nrow=2) + theme(legend.position='none') + xlab('time (steps)') + ylab('')
 
 dev.off()
 

### 2015-06-02 manuscript/draft.tex
Updated Manuscript

--- 2015-06-01 manuscript/draft.tex
@@
 \begin{abstract}
-Social learning rules are often guided by the amount of payoff received by potential social teachers while expressing specific cultural traits.
-The assumption of the social learner is that such payoff is a proxy for inferring the qualitative features of the target trait.
-Here we explore the theoretical implications when payoff directly affects  reproductive fitness and its expected amount is a function of the number of individuals possessing the given trait.
-More specifically we consider a particular form of density-dependence where a too small or too large population of individuals possessing the same variant will be subject to a decline in fitness, while an intermediate size is optimal.
-This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits, in particular subsistence strategies that are enhanced  by the effect of cooperation but are constrained by interference and limited resources.
+Several social learning strategies are guided by an evaluation of alternative cultural variants by means of the expected payoffs associated by these.
+The assumption of the social learner is that the payoff received by the transmitter while expressing a given cultural trait, is a proxy for for inferring the qualitative features of the trait itself.
+Here we explore the theoretical implications when payoff directly affects reproductive fitness and it is density dependent, that is its expected amount is a function of the number of individuals possessing the same trait.
+More specifically we consider situations where the density dependence is simultaneously direct and inverse, whereby the highest payoff is achieved at intermediate densities.This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits that are enhanced by cooperation or mutual facilitation, but are somehow constrained by some carrying capacity.
 We  explores the evolutionary dynamics of a single trait model with two variants, each exhibiting a different Allee effect.
-We used a modified Lotka-Volterra model of payoff-biased social learning, and show that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion, stable co-existence, and fixation to one variant.
-
-% XAVI: I'm confused by the use of trait and variant: shouldn't it be 'a single trait model with two variants'?
-% I'm using them as synonym, but I guess you have a point.
+We used a modified Lotka-Volterra model of payoff-biased social learning, and show that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion, stable co-existence, and fixation to single variant.
 
@@
 
-A central element of all social learning strategies is the presence of some heuristics that offers some guidance in the selection of a given variant over another.
-This  can be based on some properties of the variant itself (\emph{content-bias}) or derived from contextual cues (\emph{context-bias}), such as the commonness or the rareness of a trait, or some features associated with the transmitter such as prestige, success, or his/her similarity to the learner~\citep{henrich_mcelreath2003}.
-An expected consequence of \emph{context-bias} is the possibility of a feedback process, whereby the attractiveness of a given variant might change as a consequence of the learning strategy~\citep{kendal_etal_2009}.
-For example a strong anti-conformist bias might initially favour rare traits, but as they spread into a population their selective advantage will decline as the result of a negative feedback.
-On the other hand conformist-bias and homophily-guided learning strategies are expected to exercise a positive feedback to some form of fixation.
+The last three decades have witnessed a steady growth of formal models of cultural evolution.
+Early works inspired by population biology \citep{cavallisforza_feldman_1981,boyd1985} have laid the foundation of a now mature trans-disciplinary field which unites the social and biological sciences \citep{mesoudi_etal_2006}.
+The rich cross-fertilisation between different fields has fomented the designing of new analytical techniques as well as new usage for existing methods, but more crucially have prompted the development of theoretical models of how culture is acquired and transmitted between individuals.
+%Add reference?
 
-This paper examines a group of learning strategies generally referred to as \emph{payoff-based transmission}~\citep{schlag1998,kendal_etal_2009,lake_and_crema_2012,baldini2013,kandler_and_laland_2013,crema_lake_inpress}.
-The core assumption shared by these models is that the probability of copying a given cultural trait $i$ is positively correlated with a payoff signal $s$, determined by some function $f(i)$ expressed by individuals possessing the trait itself.
+A central element, shared by most models of social learning, is the existence of some heuristics that promotes the selection of a given cultural variant over another \citep{laland2004}.
+This can be based on some properties of the variant itself (\emph{content bias}) or derived from contextual cues (\emph{context bias}), such as the commonness or the rareness of a trait (\emph{conformist bias} and \emph{anti-conformist bias}), or some features associated with the transmitter such as prestige, success, or similarity to the learner (\emph{context bias})\citep{henrich_mcelreath2003}.
+
+This paper examines a group of social learning strategies generally referred to as \emph{payoff-based transmission} \citep{schlag1998,kendal_etal_2009,lake_and_crema_2012,baldini2013,kandler_and_laland_2013,crema_lake_inpress}.
+Here the probability of copying a cultural variant $i$ is directly proportional to a payoff $s$ observed amongst individuals possessing $i$.
+In other words the social learner assumes that the success of an individual (measured in $s$) is to some extent function of the observed behavioural trait, that is $s=f(i)$.
 Several variants of this learning strategies have been proposed and explored, including: \emph{imitate the best}, where the social learner identifies and copies the trait possessed by the individual with the highest payoff; \emph{compare means}, where the social learner compare the average payoff of different strategies and chooses the highest variant; and the \emph{proportional imitation rule}, where the probability of adopting a variant is proportional to the difference in payoff between the demonstrator and the learner~\citep{schlag1998,baldini2013}.
 
-Whilst most studies focused on how the difference in these imitation rules can drive the dynamics of the system, less attention has been paid on the payoff signal.
-The exact nature of the signal is most likely some indirect proxy that is assumed, by the social learner, to be linked with the target cultural trait.
-Thus one might be guided by a variety of cues (e.g.
-number of offpsring, income, presence/absence of other traits associated with success or prestige, etc.) that are differently correlated with the target trait resulting into different degrees of uncertainty in the signal.
-The magnitude of this uncertainty can have strong impact to efficiency of specific learning rules.
-For example,~\citet{crema_lake_inpress} have shown that \emph{imitate the best} is negatively correlated with payoff uncertainty.
-In the case of high uncertainty, a novel trait with payoff higher than the average has a decreased probability of being selected, due to being initially rare in a population.
-This can generate scenarios dominated by extant suboptimal variants.
+Whilst most studies focused on how difference in these imitation rules can drive the dynamics of the system, less attention has been given to the relationship between target cultural traits and their payoff signals.
+Typical signal are most likely some indirect proxy of success (e.g.
+number of offspring, income, presence/absence of other traits associated with success or prestige, etc.)  that are assumed, by the social learner, to be more or less linked with a cultural trait.
+In most cases the causal link between the payoff and a cultural trait is just inferred, and the exact processes and variables leading the latter to the former is most likely unknown.
+This implies that the strength of correlation between payoff and cultural trait can vary, introducing different degrees of uncertainty that can have a strong impact to the efficiency of specific learning rules.
+For example,~\citet{crema_lake_inpress} have shown that \emph{imitate the best} is becomes increasingly detrimental with high payoff uncertainty and a large pool of potential social teachers.
+This is because innovations are by definition rare in a population, and hence even if their average payoff are higher than extant suboptimal alternatives, their probability of signalling the highest payoff, and hence be selected, is low.
 
-Given the same cultural trait, the variability in the payoff signal is partly derived by the presence of interacting cultural and physical traits, as well as the environment where the trait is manifest.
-Thus the payoff is better described by the function $f(i,\epsilon)$, where $i$ is the trait, and $\epsilon$ is a series of additional hidden or latent variables that might or might interact with $i$.
-For example, a fisherman might evaluate the performance of a hook (the target cultural trait) using as a cue the number of fish captured (the payoff signal).
-The correlation between the two is partly due to the efficiency of the hook, but also by the rod, the choice of the bait, the availability of the prey species, the choice of the fishing spot, the physical properties of the fisherman, and chance.
+This uncertainty in the payoff signal is derived by the presence of interacting cultural and physical traits, as well as the environment where the trait is manifest.
+Thus the function $s=f(i)$ should be better described as $s=f(i,\epsilon)$, where $\epsilon$ represent all additional hidden and latent variables that contributes, along with the cultural variant of interest $i$, to the payoff signal $s$.
+A trivial example is a fisherman who might evaluate the performance of a new hook (the target cultural trait) using as a cue the number of fish captured (the payoff signal).
+The correlation between the two is partly derived by the efficiency of the hook, but also by the rod, the choice of the bait, the availability of the prey species, the choice of the fishing spot, the physical strength of the fisherman, and chance.
 In many cases we should expect that the social teacher and learner share many of these interacting traits as well as the environment where these are displayed.
-In other words, cultural and ecological inheritance can generally reduce the impact of $\epsilon$, and consequently the uncertainty and the variability in the payoff-signal.
+In other words, cultural and ecological inheritance should in part reduce the impact played by $\epsilon$, and consequently the uncertainty in the payoff-signal.
 There are however exceptions.
-One of these are contexts where the success of a variant is strongly correlated with the absolute number of individuals $n_i$ possessing the same trait within a population.
-This density dependence, which can be described by the payoff function $f(i,n_i)$, implies that the adoption (or abandonment) of a trait can alter the expected payoff-signal, possibly generating a variety of feedback mechanism.
+One of these are contexts where the success of a variant is strongly correlated with the absolute number of individuals $n_i$ possessing the same trait ($i$) within a population.
+This density dependence, which can be described by the payoff function $f(i,n_i)$, implies that the adoption (or abandonment) of a trait can alter its expected payoff-signal, possibly generating a variety of feedback mechanism.
 
-Several cultural traits are expected to show such density dependence.
-In particular, behaviour linked to the exploitation of limited resources are expected to exhibit a inverse density dependence.
+Several behavioural traits are expected to show such density dependence.
+In particular, behaviour linked to the exploitation of limited resources are expected to exhibit density dependence.
 Novel optimal variants are likely to spread rapidly when the trait is rare, but once a threshold frequency is exceeded (i.e.
@@
 
-However, density dependence does not necessarily imply only a negative correlation between population size and payoff/fitness.
-In many ecological contexts an inverse density dependence, whereby per capita growth rate is positively correlated with population density, has been noted.
-This phenomena, known as Allee effect \citep{allee1958,courchamp_etal_1999}, is generally explained by a range of processes from cooperation to unconscious mutual facilitation that enhances individual fitness at low population density.
-This inverse density dependence can be further distinguished between \emph{weak} and \emph{strong} Allee effect.
-In the former case, the reproductive fitness is always positive at low population density, whilst in the latter case we expect the presence of a critical density below which fitness becomes negative \citep{Wang_and_Kot_2001}.
-The implications of direct and inverse density dependence has been extensively studied both from the empirical and theoretical perspectives, including its effect of habitat selection \citep{greene_and_stamps_2001}, resource competition \citep{jang2013}, and dispersal \citep{steele_2009}, as well as specific events such as shifts in settlement pattern \citep{crema_2014}, colonisation events,  \citep{kennet_etal_2006}, genetic consequences on expanding population \citep{roques_etal_2012} and extinction of languages \citep{sutherland_2003}
+However, density dependence does not necessarily imply only a negative correlation between population size and payoff.
+In many ecological contexts per capita growth rate is positively correlated with population density, thus resulting into an inverse density dependence.
+This phenomena, known as \emph{Allee} effect \citep{allee1958,courchamp_etal_1999}, is generally explained by a range of processes including cooperation, mutual facilitation, and environmental conditioning; all processes that we expect to observe  in many cultural traits.
+The Allee effect has been extensively studied  from  empirical and theoretical standpoints, and its implications assessed for a variety of phenomena, including habitat selection \citep{greene_and_stamps_2001}, resource competition \citep{jang2013},  dispersal \citep{steele_2009}, shifts in settlement pattern \citep{crema_2014}, colonisation events,  \citep{kennet_etal_2006}, genetic diversity \citep{roques_etal_2012} and extinction of languages \citep{sutherland_2003} .
+Yet, despite this wide array of applications, we are unaware of any theoretical studies exploring how the simultaneous presence of direct and inverse density dependence in the payoff signal can affect the evolutionary dynamics of cultural transmission.
 
-Although a systematic review and a shared theoretical framework akin to those proposed in population ecology \citep{kramer_etal_2009} is not present in the social sciences, these studies suggest that there are several lines of evidence supporting the existence of both direct and inverse density dependence in cultural and behavioural traits.
+Although a systematic review and a shared theoretical framework akin to those proposed in population ecology \citep{kramer_etal_2009} is not present in the social sciences, there are several lines of evidence supporting the existence of both direct and inverse density dependence in cultural and behavioural traits.
 For example several ethnographic studies on small scale societies show how the per-capita return of hunting groups show higher yield when the group size is not too small nor too big~\citep{hill_and_hawkes_1983,janssen_and_hill_2014}.
 Other instances of inverse density dependence include population level activities such as the use of anthropogenic fire~\citep{bird2013} or selective hunting~\citep{dods_2002} which often require a population size above a certain critical density in order to promote sufficient niche enhancement and a consequent increase in payoff~\citep{rowley-conwy_and_layton_2011}.
-In a broader sense any cultural trait that benefits from direct or indirect cooperation %Add reference \citep{},
+In a broader sense any cultural trait that benefits from direct or indirect cooperation,
+%Add reference \citep{}
 promote positive niche construction \citep{vandermeer_2008}, and requires a critical mass \citep{rogers_2003} are likely to exhibit some degree of Allee effect.
@@
 Here we explore the consequences of Allee effect in the evolutionary dynamics of cultural transmission, assuming that: 1) the social learning strategy is payoff-based; and 2)  the payoff is equivalent to the reproductive fitness of each individual.
-We will demonstrate, by means of a difference equation, that direct and inverse density dependence can strongly affect the evolutionary trajectory of the system leading to the emergence of stable and unstable equilibria, as well as the occurrence of temporary or permanent episodes of reversion to suboptimal cultural variants.
+We will demonstrate, by means of a difference equation, that direct and inverse density dependence can strongly affect the evolutionary trajectory of the system, leading to changes in the conditions required to reach specific equilibria, as well as to the emergence of unstable patterns characterised by temporary or permanent episodes of shifts between alternative cultural variants.
 
 \section{The Model}
-We consider the evolutionary dynamics of two mutually exclusive cultural variants whose payoff function is characterised by both direct and inverse density dependence.
+We consider the evolutionary dynamics of two mutually exclusive cultural variants which payoff functions are characterised by an Allee effect.
 We assume a non-fixed population of size $N$, subdivided into two subpopulations with sizes $a$ and $b$, each associated with the two cultural variants.
-We further assume that $a$ and $b$ can change as a result of intrinsic population growth/decline as well as shifts from one to the other as a result of cultural transmission based.
+We further assume that $a$ and $b$ can change as a result of intrinsic population growth/decline as well as shifts from one to the other as a result of cultural transmission.
 The core model can be depicted with the following coupled difference equation \eqref{eq1}:
 
-\begin{align}
-\begin{cases}
+\begin{equation}
+\begin{aligned}
 a_{t+1}& = a_t + a_t R_{a,t} + C_{a,b,t} \\
 b_{t+1}& = b_t + b_t R_{b,t} + C_{a,b,t}
-\end{cases}
 \label{eq1}
-\end{align}
+\end{aligned}
+\end{equation}
 
-where $a_{t+1}$ and $b_{t+1}$ are the population using each trait at time $t+1$.
-This is given by the number of individuals during the previous time-step ($a_t$ and $b_t$), the reproductive rate of each population ($R_{a,t}$ and $R_{b,t}$), and the shift from one population to another dictated by cultural transmission ($C_{a,b,t}$).
-% XAVI do you want to use cultural trasmission here or better stick with social learning?
+where $a_{t+1}$ and $b_{t+1}$ are the population sizes of each variant at time $t+1$.
+This is given by the number of individuals during the previous time-step ($a_t$ and $b_t$), the reproductive rate of each population ($R_{a,t}$ and $R_{b,t}$), and the shift from one population to another as a result of cultural transmission ($C_{a,b,t}$).
 
-The reproductive growth rate $R$ of each population is defined by the following paired equation:
+The reproductive growth rate $R$ of each population is defined by the following set of equations:
 
-\begin{align}
-\begin{cases}
+\begin{equation}
+\begin{aligned}
 R_{a,t}& = r_a (\frac{a_t}{A_a}-1)(1-\frac{a_t}{K_a})\\
 R_{b,t}& = r_b (\frac{b_t}{A_b}-1)(1-\frac{b_t}{K_b}) 
-\end{cases}
 \label{eq2}
-\end{align}
+\end{aligned}
+\end{equation}
 
 where $r$ is the basic growth rate, $A$ is the critical density and $K$ is the carrying capacity, following the condition $0<A < K$.
-Equation \eqref{eq2} depicts a \emph{strong} Allee effect, hence the reproductive rate is negative when density is below $A$ or above $K$, and will reach its maxima $R_{max}=\frac{r(A-K)2}{4AK}$ when the population density is equal to $\frac{(K-A)}{2}$.
 
-Cultural transmission is based on the \emph{proportional imitation rule} \citep{schlag1998}, and hence in this case the probability of an individual with lower growth rate to adopt the trait of an individual with higher growth rate is proportional to the difference in the growth rates.
-This is portrayed in \eqref{eq3}:
+Equation \eqref{eq2} depicts the so called \emph{strong} Allee effect \citep{Wang_and_Kot_2001}, where the reproductive rate is negative when population density is either below $A$ or above $K$, and  in this case reach its maxima $R_{max}=\frac{r(A-K)2}{4AK}$ when density is equal to $\frac{(K-A)}{2}$.
 
-\begin{align}
+The cultural transmission component of equation \eqref{eq1} is based on the \emph{proportional imitation rule} \citep{schlag1998}.
+In this case, the probability of an individual with lower growth rate to adopt the trait of an individual with higher growth rate is proportional to the difference in the growth rates.
+This is formally portrayed as follows:
+
+\begin{equation}
+\begin{aligned}
 \label{eq3}
@@
 \end{cases}
-\end{align}
+\end{aligned}
+\end{equation}
 
@@
 
-\begin{align}
+\begin{equation}
+\begin{aligned}
 \label{eq4}
@@
 \begin{cases}
-z& \text{if }|R_{a,t}-R_{b,t}| > b\Delta\\
-z\frac{|R_{a,t}-R{b,t}|}{m\Delta}& \text{if }|R_{a,t}-R_{b,t}| \leq b\Delta\\
+z& \text{if }|R_{a,t}-R_{b,t}| > m\Delta\\
+z\frac{|R_{a,t}-R{b,t}|}{m\Delta}& \text{if }|R_{a,t}-R_{b,t}| \leq m\Delta\\
 \end{cases}
-\end{align}
+\end{aligned}
+\end{equation}
 
-% XAVI b is not explained? %It's no longer b but m
-where $z$ is a measure of reliance in social learning (effectively equivalent to highest possible transmission rate), $\Delta$ is the maximum between $R_{max(a,t)}$ and $R_{max(b,t)}$, and $m$ is a calibration parameter that measure the perception of the difference in the payoffs (i.e.
+where $z$ is a measure of reliance in social learning (effectively equivalent to highest possible transmission rate), $\Delta$ is the maximum between $R_{max(a,t)}$ and $R_{max(b,t)}$, and $m$ is a calibration parameter that measures the perception of the difference in the payoffs (i.e.
 small values of $m$ determines a higher rate by which the theoretical maximum transmission rate is reached).
@@
 
-When $z=0$, equations \ref{eq1}-\ref{eq4} \emph{de facto} portray a standard cubic growth model with Allee effect, where the success (or failure) of a given focal trait is simply given by its own density, and the density of alternative cultural traits has no influence.
-For a generic cultural trait $x$, the model predicts three point attractors at $x=0$ (extinction), $x=A_x$, and $x=K_x$.
+When $z=0$, equations \ref{eq1}-\ref{eq4} \emph{de facto} portray a standard cubic growth model with Allee effect, where the success (or failure) of a given focal trait is exclusively given by its own density.
+For a generic cultural trait $x$, the model predicts three equilibrium points at $x=0$ (extinction), $x=A_x$, and $x=K_x$.
 The equilibrium $x=A_x$ is unstable in all cases, whereas the stability of the node at $x=K_x$ is a function of the per capita growth rate $r_x$.
@@
 \ref{fig:bifurcationDiagram}).
-% Not sure if there are strange attractors.
-Reference  any  bifurcation analysis on the equation we used.
-Scheuring 1999 on JTB?
 
@@
       \includegraphics[width=0.5\textwidth]{./figures/figure1.jpg}
-  \caption{Bifurcation diagram of a single population model with $A_x=0.2$ and $K_x=0.8$. For $0<r_x<0.6649747$, any initial value of $x$ greater than $A_x$ converge to $K_x$.}
+  \caption{Bifurcation diagram of a single population model with $A_x=0.2$ and $K_x=0.8$. For $0<r_x<0.6649747$, any initial value of $x$ greater than $A_x$ converges to $K_x$.}
     \label{fig:bifurcationDiagram}
@@
 Here we consider only scenarios where the node $x=K_x$ is stable.
-In this case the initial value of $x(t=0)$ will determine the ultimate equilibria of the system: extinction, if $x(t=0)<A_x$; unstable equilibria, if $x(t=0)=A_x$; and stable equilibrium, if $x(t=0)>A_x$.
-Thus in a 2-trait model, the coexistence of both variants is guaranteed only when the initial density of both traits are above their respective Allee threshold, otherwise we should expect that either (i.e.
+In this case the initial value of $x(t=0)$ will determine the ultimate equilibria of the system: extinction, if $x(t=0)<A_x$; unstable equilibrium, if $x(t=0)=A_x$; and stable equilibrium, if $x(t=0)>A_x$.
+Thus in a 2-trait model, the coexistence of both variants is guaranteed only when the initial density of both traits are above their respective critical densities, otherwise we should expect that either (i.e.
 $a(t=0) \geq A_a$ and $b(t=0)<A_b$ or $a(t=0)<A_a$ and $b(t=0) \geq A_b$) or both (i.e.
-when $a(t=0)<A_a$ and $b(t=0)<A_b$) variants will be ultimately extinct (see fig.
+when $a(t=0)<A_a$ and $b(t=0)<A_b$) variants are ultimately extinct (see fig.
 \ref{fig:NoTransmissionBasin}).
@@
 when $z>0$), the stability of the nodes are perturbed by individuals switching from one strategy to another.
-The minimum implication of these perturbations is a change in the shape of the basins of attractions.
-Initial conditions closer to the boundaries between one basin and another will most profoundly be affected, potentially leading to a complete different equilibria with slightest change in the degree of reliance on social learning.
+We identify two major changes in the equilibrium properties of the system as a result of cultural transmission.
+Firstly the shape of the basins of attraction are altered, and hence initial conditions closer to the boundaries between one basin and another will be strongly affected,  leading occasionally to very different evolutionary equilibria.
 Figure \ref{fig:TransmissionBasin} shows an example of this.
 When social learning is not present (fig.~\ref{fig:TransmissionBasin}a), the initial conditions $i$, $j$, and $k$ result into a stable equilibrium at $a=K_a$ and $b=K_b$, whilst the starting point $l$ leads to the extinction of $b$.
-When $z=0.05$ (fig.~\ref{fig:TransmissionBasin}b), the basins of attractions are modified, resulting into different trajectories and possibly equilibria for the same initial conditions without social learning.
-Thus $i$ and $k$ are no longer leading to the coexistence of both traits but to the demise of one or the other, whilst $l$ now ensure the coexistence of both variants.
+When $z=0.05$ (fig.~\ref{fig:TransmissionBasin}b), the basins of attractions are modified, resulting into different trajectories and in some cases different equilibria for the same initial conditions.
+Thus $i$ and $k$ are no longer leading to the coexistence of both traits but to the demise of one or the other, whilst $l$ now ensures the coexistence of both variants.
 
 % XAVI put high res figure
-\begin{figure}[h!]
+\begin{figure*}
   \centering
-      \includegraphics[width=0.5\textwidth]{./figures/figure3.pdf}
+      \includegraphics[width=2\columnwidth]{./figures/figure3.pdf}
   \caption{Basins of attraction for the four equilibria (mutual extinction: grey; extinction of $a$: blue; extinction of $b$: red; stable coexistence: green) with (right) and without (left) social learning.
@@
     \label{fig:TransmissionBasin}
-\end{figure}
+\end{figure*}
 
@@
 Figure \ref{fig:bifurcationWithTransmission} shows the bifurcation diagram for three different initial settings of $a$ and $b$.
-In the first two cases, increase in $z$ leads to a shift of the equilibria, that remains however constant for larger reliance on social learning.
+In the first two cases, increase in $z$ leads to a shift of the equilibria, which remains stable even when $z$ is increased.
 The third case shows, around $z=0.5$ a transition from a point attractor to a system characterised by repeated shifts between the two transmission strategies.
-% XAVI shall we generate the pdf here?
-\begin{figure}[h!]
+
+\begin{figure*}
   \centering
-      \includegraphics[width=0.5\textwidth]{./figures/figure4.jpg}
+      \includegraphics[width=1\textwidth]{./figures/figure4.jpg}
   \caption{Bifurcation diagram for $a$ (red) and $b$ (blue) as a function of social learning ($z$): left: $a=0.11$, $b=0.4$; Middle: $a=0.5$, $b=0.21$, $a=0.4$, $b=0.4$.
@@
     \label{fig:bifurcationWithTransmission}
-\end{figure}
+\end{figure*}
 
-%ENRICO: Add here basin plot with chaotic system and show tiny difference in initial condition can lead to stable coexistence or chaotic system
+High reliance on social learning thus generates a new basin of attraction ...
+%Not sure how to link this to the unstability figures.
+I think we already made the point with the analysis above, and I'm no longer sure what this new plot is ultimately adding (I know, I asked you to make it, sorry!)
 
-\subsection{Effects of strategy overlap}
+%\begin{figure}
+%  \centering
+%      \includegraphics[width=0.5\textwidth]{./figures/figureUnstable1.pdf}
+%  \caption{}
+%    \label{fig:bifurcationWithTransmission}
+%\end{figure}
+
+\subsection{Differences in critical density and carrying capacities}
 % XAVI better title?
 
-Previous results suggest that the percentage of population that can switch strategies (i.e.
-the $z$ parameter) has major effects on the final equilibria.
-However, this will also depend on the interaction between the intervals where frequency dependence is $\geq 0$ (i.e.
-$[A_a,K_a]$ and $[A_b,K_b]$) .
-We have explored this relation introducing a new parameter $\lambda \in (0,1)$ and defining the positive interval for strategy $b$ as follows:
-
-\begin{align}
+The analyses of the previous section suggest that maximum proportion of population that can switch strategies (i.e.
+the $z$ parameter) has major effects on the final equilibria (c.f.
+fig.\ref{fig:bifurcationWithTransmission}).
+However, this will also depend on the key parameters defining the range of positive payoff for the two strategies (i.e.
+$[A_a,K_a]$ and $[A_b,K_b]$).
+While a comprehensive analysis is not possible, here we illustrate the relationship between these parameters and $z$ by introducing a new parameter $\lambda \in (0,1)$ adding the following constraints to equation \eqref{eq1}:
+%I slightly revised your intro here.
+Mainly because this reads as if changes depends on the intervals of positive payoff alone, while I think it actually depends on any possibly combination of values.
+\begin{equation}
+\begin{aligned}
 \label{eqOverlap}
-\begin{cases}
 A_b = A_a + \lambda A_a\\
 K_b = K_a + \lambda K_a
-\end{cases}
-\end{align}
+\end{aligned}
+\end{equation}
+
+%ENRICO:
+%I'm a bit confused by \eqref{eqOverlap}.
+Clearly lambda is no longer a measure of overlap percentage, as when lambda=0 we have identical strategy, but when lambda=1 we can have a variety of patterns depending on what you have at A_a and K_a.
+For example if you have A_a=3 and K_a=5, you will get A_b=6 and K_b=10, thus no overlap between the two strategies.
+But if you have A_a=1 and K_a=5, then you will have A_b=2, and K_b=10 which implies an almost complete overlap.
+Besides this approach does not hold the differences constant.
+I thought you did:
+
+% \begin{equation}
+%\begin{aligned}
+%\label{eqOverlap}
+%A_b = A_a + (1-\lambda A_a)(K_a-A_a)\\
+%K_b = K_a + (1-\lambda A_a)(K_a-A_a)
+%\end{aligned}
+% \end{equation}
+
+%which would instead give, with lambda=0 this time, A_a=3 K_a=5 A_b=5, K_b=7 for the first example, and  A_a=1 K_a=5 A_b=5, K_b=9 for the second.
 
@@
 High values indicate that the minimum number of individuals needed to achieve positive fitness in $b$ gets closer to $K_a$, while the population size that $b$ can sustain also grows.
+
 As a consequence, the number of individuals switching strategies when $b$ has lower payoff is also higher and could potentially overload $a$ due to its lower $K_a$.
@@
 Figure \ref{fig:overlap} shows how the degree of overlap affects the equilibria of our model with and without social learning.
-The relation between the two processes strongly affects the shape of the different basins in a non-dynamic fashion.
+The relation between the two processes strongly affects the extent of the different basins in a non-dynamic fashion.
+%I've put "extent" instead of "shape" as we are not directly assessing the latter. Also I would omit "strongly".
 
@@
 
-\begin{figure}[h!]
+\begin{figure*}
   \centering
@@
     \label{fig:overlap}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[h!]
+%Ok this reads more like my interpretation of lambda..
+
+\begin{figure*}
   \centering
       \includegraphics[width=\textwidth]{./figures/figure6}
-  \caption{Percentages of the entire space falling in each basin of attraction for different values of $\lambda$. X axis controls the parameter $z$ defining social learning.}
+  \caption{Proportion of initial conditions falling in each basin of attraction for different values of $\lambda$.
+X axis controls the parameter $z$ defining social learning.} %We need to add what is the range of initial conditions being explored and what are the initial values of A_a and K_a
     \label{fig:percentages}
-\end{figure}
+\end{figure*}
 
 \subsection{Effect of competition}
-%what if competition
-%show new stable points
-%show dynamics?
 
-The model described in equations \ref{eq1}-\ref{eq4} assumes that the resource pool of the two strategy are completely independent to each other.
-In reality many strategies have some degree of competition, whereby the frequency dependence is not exclusively internal.
-In other words, the frequency of one trait might increase or decrease the fitness of the other.
-The following equation can substitute equation~\eqref{eqCompetition}:
+Figure \ref{fig:percentages} also suggest that with decreasing values of $\lambda$, the variant with lower carrying capacity/critical density expands its basin of attraction.
+This implies that, with other things being equal, traits with higher $A$ and $K$ have a smaller number of favourable conditions that can promote its existence.
+Indeed in the case of the two-variants model examined here, the strategy with higher $A$ and $K$ can survive primarily when its initial density is comparatively high (e.g.
+above or equal its maximum payoff) and when the alternative strategy is close to its carrying capacity.
+The latter condition implies that novel strategies with higher carrying capacity but also higher critical densities are fully adopted only when the extant strategy with lower $A$ and $K$ has a density closer to its upper threshold.
+Thus innovations capable to sustain larger population density requires either a comparatively lower critical density that ensures a higher overlap with extant strategies, or an initial number of adoptees that are not guided by a payoff-based transmission.
+A third option, which we briefly explore here, is to relax the assumption that the payoff of each variant is exclusively a function of their intrinsic density.
+For instance several subsistence strategies do not have fully independent resource pools, and hence the success of a given variant might be detrimental to the payoff of the other irregardless of the effects provoked by cultural transmission.
+We can revise our core model by including new terms portraying the competition effect between the two resources:
 
-\begin{align}
-\begin{cases}
+\begin{equation}
+\begin{aligned}
 R_{a,t}& = r_a (\frac{a_t-c_{A_a}b_t}{A_a}-1)(1-\frac{a_t-c_{K_a}b_t}{K_a})\\
 R_{b,t}& = r_b (\frac{b_t-c_{A_b}a_t}{A_b}-1)(1-\frac{b_t-c_{K_b}a_t}{K_b}) 
-\end{cases}
 \label{eqCompetition}
-\end{align}
+\end{aligned}
+\end{equation}
 
@@
 
+%what if competition
+%show new stable points
+%show dynamics?
+
 Figure \ref{fig:competition} shows a scenario where strategy $a$ is dominant over strategy $b$.
 
-\begin{figure}[h!]
+\begin{figure*}[h!]
   \centering
@@
     \label{fig:competition}
-\end{figure}
+\end{figure*}
 
 % XAVI some brief discussion on results here, and update the dashed lines
+
+% Do we need more? Perhaps another iteration of the overlap analysis with the competition promoting the variant with higher critical density and carrying capacity?
 

### 2015-06-03 manuscript/draft.tex
updated lambda equations
Update

@@
 
-Equation \eqref{eq2} depicts the so called \emph{strong} Allee effect \citep{Wang_and_Kot_2001}, where the reproductive rate is negative when population density is either below $A$ or above $K$, and  in this case reach its maxima $R_{max}=\frac{r(A-K)2}{4AK}$ when density is equal to $\frac{(K-A)}{2}$.
+Equation \eqref{eq2} depicts the so called \emph{strong} Allee effect \citep{Wang_and_Kot_2001}, where the reproductive rate is negative when population density is either below $A$ or above $K$, and  in this case reach its maxima $R_{max}=\frac{r(A-K)^2}{4AK}$ when density is equal to $\frac{(K-A)}{2}$.
+
+%Check if $R_{max}=\frac{r(A-K)^2}{4AK}$ this actually true, or is rather $R_{max}=\frac{r(K-A)2}{4AK}$
 
@@
 
+%This should be changed with the basin plot and the time-series; omit discussion on bifurcation
+
 \begin{figure*}
@@
 I think we already made the point with the analysis above, and I'm no longer sure what this new plot is ultimately adding (I know, I asked you to make it, sorry!)
+
+Small differences in initial conditions can have major consequences in the boundaries between the unstable region and the rest of equilibria.
+Figure \ref{fig:unstable} shows an example with high value of social learning $z=0.4$, where two scenarios ($j$ and $k$) generate different outcomes from very close initial conditions.
 
@@
 While a comprehensive analysis is not possible, here we illustrate the relationship between these parameters and $z$ by introducing a new parameter $\lambda \in (0,1)$ adding the following constraints to equation \eqref{eq1}:
-%I slightly revised your intro here.
-Mainly because this reads as if changes depends on the intervals of positive payoff alone, while I think it actually depends on any possibly combination of values.
+%I slightly revised your intro here. Mainly because this reads as if changes depends on the intervals of positive payoff alone, while I think it actually depends on any possibly combination of values.
+
 \begin{equation}
@@
 \label{eqOverlap}
-A_b = A_a + \lambda A_a\\
-K_b = K_a + \lambda K_a
+A_b = A_a + (1-\lambda)(K_a-A_a)\\
+K_b = K_a + (1-\lambda)(K_a-A_a)
 \end{aligned}
 \end{equation}
-
-%ENRICO:
-%I'm a bit confused by \eqref{eqOverlap}.
-Clearly lambda is no longer a measure of overlap percentage, as when lambda=0 we have identical strategy, but when lambda=1 we can have a variety of patterns depending on what you have at A_a and K_a.
-For example if you have A_a=3 and K_a=5, you will get A_b=6 and K_b=10, thus no overlap between the two strategies.
-But if you have A_a=1 and K_a=5, then you will have A_b=2, and K_b=10 which implies an almost complete overlap.
-Besides this approach does not hold the differences constant.
-I thought you did:
-
-% \begin{equation}
-%\begin{aligned}
-%\label{eqOverlap}
-%A_b = A_a + (1-\lambda A_a)(K_a-A_a)\\
-%K_b = K_a + (1-\lambda A_a)(K_a-A_a)
-%\end{aligned}
-% \end{equation}
-
-%which would instead give, with lambda=0 this time, A_a=3 K_a=5 A_b=5, K_b=7 for the first example, and  A_a=1 K_a=5 A_b=5, K_b=9 for the second.
 
@@
 From $0.9 < \lambda < 0.55$ the dynamics remain remarkably similar, while for $\lambda = 0.5$ the effects of social learning decreases because of the large gap between $A$ and $K$ of both strategies.
-
-%Ok this reads more like my interpretation of lambda..
 

### 2015-06-05 manuscript/draft.tex
updated plots

--- 2015-06-02 manuscript/draft.tex
@@
 Here we explore the theoretical implications when payoff directly affects reproductive fitness and it is density dependent, that is its expected amount is a function of the number of individuals possessing the same trait.
-More specifically we consider situations where the density dependence is simultaneously direct and inverse, whereby the highest payoff is achieved at intermediate densities.This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits that are enhanced by cooperation or mutual facilitation, but are somehow constrained by some carrying capacity.
+More specifically we consider situations where the density dependence is simultaneously direct and inverse, whereby the highest payoff is achieved at intermediate densities.
+This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits that are enhanced by cooperation or mutual facilitation, but are somehow constrained by some carrying capacity.
 We  explores the evolutionary dynamics of a single trait model with two variants, each exhibiting a different Allee effect.

### 2015-06-05 experimentMN3/alleeTasks.py

@@ -9,23 +9,21 @@ def main():
     zValues = [0.05, 0.2]
     lambdaValues = [0.5, 0.75, 1]
     """
+    """
     # 2 - exploration lambda/z (relative figure)
     competitionValues = [0.0]
     zValues = arange(0,1.1,0.1)
-    lambdaValues = arange(0.5,1.1,0.05)
-
+    lambdaValues = arange(0.5,1.05,0.05)
     """
     # 3 - exploration competition/z 
-    competitionValues = [-0.1,0.1]
+    competitionValues = [0.1,0.25,0.5]
     zValues = [0.05, 0.2]
     lambdaValues = [0.75]
-    """
-
     """
     # 4 - z (figure 3)
     competitionValues = [0.0]
     zValues = [0,0.05]
-    lambdaValues = [1/3]
+    lambdaValues = [2/3]
     """
 
     tasksFile = open('allee.tasks', 'w')

### 2015-06-05 experimentMN3/run.R

@@ -14,8 +14,7 @@ lambda = as.numeric(args[4])
 # fixed values
 Am=0.2
 Km=0.8
-r.n=0.05
-r.m=r.n
+r.m=0.05
 x.lim=c(0,1.2)
 y.lim=c(0,1.2)
 
@@ -27,6 +26,7 @@ cKm=-competition
 diff = Km-Am
 An=Am + (1-lambda)*diff
 Kn=Km + (1-lambda)*diff
+r.n=r.m*EqualiseMaxGrowthRate(An,Am,Kn,Km)
 
 result <- basin(x.lim=x.lim,y.lim=x.lim,points=300,timeSteps=10000,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,z=z,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,verbose=F,multicore=TRUE,cores=16)
 

### 2015-06-05 manuscript/figures/figure6.R

@@ -12,6 +12,8 @@ getRawValues <- function(run)
     print(paste('loading:',fileName))
     load(fileName)
 
+    result <- subset(result, ini.m>=Am & ini.m<=Km & ini.n>=An & ini.n<=Kn)
+
     #Identify Stable Attractor Node
     Em=ExpStableNode(Kn,Km,cKn,cKm)$m
     En=ExpStableNode(Kn,Km,cKn,cKm)$n
@@ -46,14 +48,18 @@ getPercentages<- function(raw)
 }
 
 raw <- data.frame()
-for(run in 0:65)
+for(run in 0:77)
 {
     raw <- rbind(raw, getRawValues(run))
 }
 
 percentages <- getPercentages(raw)
+percentages <- subset(percentages, lambda<=1.0)    
+colorValues = c('#006400','#D3D3D3','#4169E1','#CD5C5C','gold2')
 
-plotted <- subset(percentages, lambda>=0.9 | lambda<=0.6)
+#ggplot(percentages, aes(x=z, y=value, fill=eq)) + geom_area(stat='identity') + scale_fill_manual(name='Equilibria', values=colorValues)+ facet_wrap(~lambda)
+
+plotted <- subset(percentages, lambda >=0.9 | lambda<=0.6 )
 plotted$lambda <- factor(plotted$lambda, levels=c(1,0.95,0.9,0.6,0.55,0.5))    
 
 # coexistence: #006400
@@ -61,7 +67,10 @@ plotted$lambda <- factor(plotted$lambda, levels=c(1,0.95,0.9,0.6,0.55,0.5))
 # nExtinct: #4169E1
 # mExtinct: #CD5C5C
 
+# ggplot(percentages, aes(x=z, y=value, fill=eq)) + geom_area(stat="identity") + facet_wrap(~lambda)
 pdf('figure6.pdf', width=10, height=6)
-ggplot(plotted, aes(x=z, y=value, fill=eq)) + geom_area(stat='identity') + scale_fill_manual(name='Equilibria', values=c('#006400','#D3D3D3','#4169E1','#CD5C5C','gold2'))+ facet_wrap(~lambda, nrow=2)
+
+ggplot(plotted, aes(x=z, y=value, fill=eq)) + geom_area(stat='identity') + scale_fill_manual(name='Equilibria', values=colorValues)+ facet_wrap(~lambda, nrow=2)
+
 dev.off()
 

### 2015-06-05 manuscript/figures/figure7.R

@@ -3,17 +3,9 @@ source("../../src/basin.R")
 source("../../src/utility.R")
 
 
-pdf('figure7.pdf', width=12, height=12)
-par(mfrow=c(2,2))
+pdf('figure7.pdf', width=12, height=9)
+par(mfrow=c(2,3))
 load('../../experimentMN3/03_competition/result_000.Rdata')
-title=substitute(paste(z == i, " ", c== j), list(i=z,j=competition))
-basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b")
-abline(h=c(An,Kn),lty=2,col="white")
-abline(v=c(Am,Km),lty=2,col="white")
-axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
-axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
-
-load('../../experimentMN3/03_competition/result_001.Rdata')
 title=substitute(paste(z == i, " ", c== j), list(i=z,j=competition))
 basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b")
 abline(h=c(An,Kn),lty=2,col="white")
@@ -29,7 +21,31 @@ abline(v=c(Am,Km),lty=2,col="white")
 axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
 axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
 
+load('../../experimentMN3/03_competition/result_004.Rdata')
+title=substitute(paste(z == i, " ", c== j), list(i=z,j=competition))
+basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b")
+abline(h=c(An,Kn),lty=2,col="white")
+abline(v=c(Am,Km),lty=2,col="white")
+axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
+axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
+
[…]

### 2015-06-05 manuscript/draft.tex
Minor Updates

@@
-%\documentclass[review,authoryear]{elsarticle}
- \documentclass[3p,authoryear,twocolumn]{elsarticle} %Just to give a feel of how it would look like
+\documentclass[review,authoryear]{elsarticle}
+% \documentclass[3p,authoryear,twocolumn]{elsarticle} %Just to give a feel of how it would look like
 
@@
 
-%Check if $R_{max}=\frac{r(A-K)^2}{4AK}$ this actually true, or is rather $R_{max}=\frac{r(K-A)2}{4AK}$
-
 The cultural transmission component of equation \eqref{eq1} is based on the \emph{proportional imitation rule} \citep{schlag1998}.
@@
 
-We investigate the equilibrium conditions for different initial population sizes of  $a$ and $b$ and different settings of $z$, $A$, and $K$, assuming that  $r_a=r_b$, and for convenience holding the condition $K_b \geq K_a$.
-Given that equations \ref{eq1}-\ref{eq4} have no analytical solution, we identified equilibrium conditions via simulation (see appendix 1 for details; source code can be found on the following repository %put link here
-as well as in the ESM).  
-%if the points we need to reiterate are short, perhaps a footnote will suffice instead of an appendix? 
-% 100,000 iterations 
-% basin plots resolution
+We investigate the equilibrium conditions for different initial population sizes of $a$ and $b$ and different settings of $z$, $A$, and $K$, ensuring that in all cases $K_b \geq K_a$.
+We assumed that two strategies differed only in their critical densities and carrying capacities but not on their intrinsic payoffs.
+To achieve this we set $r_b=\rho r_a$, where $\rho$ is a constant\footnote{Given two variants with the same baseline growth rate (i.e.
+$r_a=r_b$), equation \eqref{eq2} result into a lower $R$ for the variant with higher $A$ and $K$.
+In order to avoid this we increased the baseline of the strategy with lower $R$ by a factor $\rho$, given by the following equation:
+\begin{align*}
+\rho=\frac{A_bK_b(K_a-A_a)^2}{A_aK_a(A_b-K_b)^2}
+\end{align*}
+  } 
+that ensures that the two variants have always the same $R_{max}$.
+Given that equations \ref{eq1}-\ref{eq4} have no analytical solution, we identified equilibrium conditions via simulation (source codes of the simulation can be found on the following repository \textcolor{red}{PUT LINK HERE})
 
@@
 
-When $z=0$, equations \ref{eq1}-\ref{eq4} \emph{de facto} portray a standard cubic growth model with Allee effect, where the success (or failure) of a given focal trait is exclusively given by its own density.
+When $z=0$, equations \ref{eq1}-\ref{eq4} becomes a standard cubic growth model with Allee effect, where the success (or failure) of a given focal trait is exclusively given by its own density.
 For a generic cultural trait $x$, the model predicts three equilibrium points at $x=0$ (extinction), $x=A_x$, and $x=K_x$.
-The equilibrium $x=A_x$ is unstable in all cases, whereas the stability of the node at $x=K_x$ is a function of the per capita growth rate $r_x$.
+The equilibrium $x=A_x$ is unstable for all values of $r_x$, whereas the stability of the node at $x=K_x$ is a function of the per capita growth rate $r_x$.
 Increasing the latter decreases the stability of fixed points~\citep{scheuring_1999}, leading to the transition from point attractors (with equilibrium at $x=K_x$), to limit cycles and chaotic dynamics with islands of stability (see fig.
@@
   \centering
-      \includegraphics[width=0.5\textwidth]{./figures/figure1.jpg}
+      \includegraphics[width=1\textwidth]{./figures/figure1.jpg}
   \caption{Bifurcation diagram of a single population model with $A_x=0.2$ and $K_x=0.8$. For $0<r_x<0.6649747$, any initial value of $x$ greater than $A_x$ converges to $K_x$.}
@@
 
-Here we consider only scenarios where the node $x=K_x$ is stable.
+For the rest of the paper we will consider only scenarios where the node $x=K_x$ is stable.
 In this case the initial value of $x(t=0)$ will determine the ultimate equilibria of the system: extinction, if $x(t=0)<A_x$; unstable equilibrium, if $x(t=0)=A_x$; and stable equilibrium, if $x(t=0)>A_x$.
@@
   \centering
-      \includegraphics[width=0.5\textwidth]{./figures/figure2.pdf}
+      \includegraphics[width=1\textwidth]{./figures/figure2.pdf}
   \caption{Basins of attraction for the four equilibria (mutual extinction: grey; extinction of $a$: blue; extinction of $b$: red; stable coexistence: green).

### 2015-06-08 manuscript/draft.tex
Updates

@@
 When the magnitude of social learning is sufficiently high, the system is subject to effects similar to an increase in the per capita growth rate $r$, with the consequent decrease in the stability of the fixed points.
-Figure \ref{fig:bifurcationWithTransmission} shows the bifurcation diagram for three different initial settings of $a$ and $b$.
-In the first two cases, increase in $z$ leads to a shift of the equilibria, which remains stable even when $z$ is increased.
-The third case shows, around $z=0.5$ a transition from a point attractor to a system characterised by repeated shifts between the two transmission strategies.
-
-%This should be changed with the basin plot and the time-series; omit discussion on bifurcation
-
-\begin{figure*}
-  \centering
-      \includegraphics[width=1\textwidth]{./figures/figure4.jpg}
-  \caption{Bifurcation diagram for $a$ (red) and $b$ (blue) as a function of social learning ($z$): left: $a=0.11$, $b=0.4$; Middle: $a=0.5$, $b=0.21$, $a=0.4$, $b=0.4$.
-Model parameters:  $A_a=0.1$; $K_a=0.6$; $A_b=0.2$; $K_b=0.7$; $r_a=r_b=0.05$}
-    \label{fig:bifurcationWithTransmission}
-\end{figure*}
-
-High reliance on social learning thus generates a new basin of attraction ...
-%Not sure how to link this to the unstability figures.
-I think we already made the point with the analysis above, and I'm no longer sure what this new plot is ultimately adding (I know, I asked you to make it, sorry!)
-
-Small differences in initial conditions can have major consequences in the boundaries between the unstable region and the rest of equilibria.
-Figure \ref{fig:unstable} shows an example with high value of social learning $z=0.4$, where two scenarios ($j$ and $k$) generate different outcomes from very close initial conditions.
+Figure XXXX1 shows how an increase in $z$ generate time-series with shifts from one variant to the other.
+This regime is prone to occur when both variants are initialised with high densities (XXXX2/basin)/.
+In these conditions the total density is too high to ensure a stable coexistence, and the regulatory effect of negative fitness is avoided by the fast rates of cultural transmission (i.e.
+individual shift strategy before experiencing a population decline resulting from density being over $K$).
+Figure XXX2 also implies that high reliance on social learning leeds to the emergence of a new basin of attraction for these unstably cycles, with small differences in initial conditions leading again to profoundly different dynamics.
 
@@
 %  \caption{}
-%    \label{fig:bifurcationWithTransmission}
+%    \label{}
+%\end{figure}
+
+%\begin{figure}
+%  \centering
+%      \includegraphics[width=0.5\textwidth]{./figures/figureUnstable2.pdf}
+%  \caption{}
+%    \label{}
 %\end{figure}

### 2015-06-09 manuscript/draft.tex
Update
This is the last update for my part, there is lots of points raised in
the comments of the tex file and more stuff I wanted to do but I really
ran out of time. I’ll email you with more details on this.

@@
 
+%I attached the list of results that I sent you over the email that we need  to put in the text
+
+%1)With no Allee effect the strategy with the highest payoff wins  /// This is obvious we don?t need to prove this 
+%2) With carrying capacity, but no Allee effect we get possible cases of reversions.
+//Lake and Crema explores this, but the model has also uncertainty in the payoff signal so not sure if it is technically comparable?
+Should we do something with this?
+%
+%
+%** Expected pattern without transmission **
+%-3)We have our unstable nodes, stable nodes, and extinction. As long as the strategies are above their critical densities, they will survive. // DONE
+%
+%** Expected pattern with transmission **
+%4) There is a possibility for a trait to not get extinct even if its below its critical density.
+This happens only if: a) the alternative strategy is also below the critical density and its distance from A is further b) the alternative strategy is beyond carrying capacity.
+If the alternative strategy is in the positive payoff interval the focal strategy will go extinct /// This is shown with all the basin plots,  so no need to anything extra.
+%
+%5) If both traits are above their respective critical densities, we expect stable coexistence when there is no transmission, but when z>0, strategies that are closer to their critical densities have a non-zero probability to go extinct.
+This probability increases as the alternative population approaches its own K or A.
+/// This is shown by the basin plots, so its fine.
+%
+%6) As a corollary to point 5, with increasing z (but also as a function of lambda), the shared perimeter of the boundary of the basin of attractions of the two strategies increases in proximity to the both critical densities.
+This means that small changes in behaviour (z) or external fluctuations can promote the adoption of one or the other but not the both.
+// This is hard to prove I guess, and not sure if the basin plots are sufficient to state this.
+%
+%7) If both variants have initial conditions within their positive payoff density ranges, a decrease in lambda promotes the coexistence of both strategy for a given value of z.
+Thus with increasing ?distance?
+between the two strategies, its less likely that we observe the dominance of one or the other.
+We ought to find out why this is the case though, perhaps we need to look at individual time-series to have some clues.
+// We should be able to see this with the overlap analysis + possibly individual time-series to get some grips
+%
+%8) With increasing values of z, the basin of coexistence is decreased to the expense of limit cycles (can we check if it is non-chaotic using Lyapunov?) and extinction of one strategy or the other.
+The growth is however asymmetric (maybe), and the strategy with high A and K have will have a smaller proportion of initial conditions where it can dominate.
+// overlap analysis should, hopefully, still show this.
+%
+%9) If point 8 is true, we can then argue that transition to strategies with higher carrying capacity is difficult, and entails in many case stable or unstable forms of coexistence.
+The only exception is when the two strategies is interacting //competition analysis, with competition values and two transmission settings.
+%
+%10) or the critical densities are the same // just a single basin plot would suffice.
+
 We investigate the equilibrium conditions for different initial population sizes of $a$ and $b$ and different settings of $z$, $A$, and $K$, ensuring that in all cases $K_b \geq K_a$.
@@
 We identify two major changes in the equilibrium properties of the system as a result of cultural transmission.
-Firstly the shape of the basins of attraction are altered, and hence initial conditions closer to the boundaries between one basin and another will be strongly affected,  leading occasionally to very different evolutionary equilibria.
+
+First, the shape of the basins of attraction are altered, and hence initial conditions closer to the boundaries between one basin and another will be strongly affected,  leading occasionally to very different evolutionary equilibria.
 Figure \ref{fig:TransmissionBasin} shows an example of this.
@@
 
-% XAVI put high res figure
 \begin{figure*}
@@
 \end{figure*}
+%Is it possible to change this figure to the settings for lambda=0.75 in figure5.pdf, setting as z 0 and 0.2?
+This way we get a much stronger changes in the basin of attraction, and we could also mention here point 6 above?
 
-When the magnitude of social learning is sufficiently high, the system is subject to effects similar to an increase in the per capita growth rate $r$, with the consequent decrease in the stability of the fixed points.
-Figure XXXX1 shows how an increase in $z$ generate time-series with shifts from one variant to the other.
-This regime is prone to occur when both variants are initialised with high densities (XXXX2/basin)/.
-In these conditions the total density is too high to ensure a stable coexistence, and the regulatory effect of negative fitness is avoided by the fast rates of cultural transmission (i.e.
+Second, when the magnitude of social learning is high, the system is subject to effects similar to those observed with an increase in the per capita growth rate $r$ (cf.
+fig.
+\ref{fig:bifurcationDiagram}), that is a decrease in the stability of the fixed points.
+Figure XXXX shows an example where the increase in $z$ generates a transition from stable coexistence to cyclical fluctuations.
+%We need to make this figure as time-series, or actually use the bifurcation diagram.
+This regime is prone to occur when both variants are initialised with high densities.
+%This refers to the basin plot figureUnstable1.pdf that you've already made.
+Given the description below, do you think that those tiny "islands" of orange inside the green are actually specific values that are covered by the fluctuations in the time-series?
+Given that the system is deterministic, it wouldn't be a surprise, if anything there MUST be a combination of values for these...
+In these conditions the total density is too high to ensure a stable coexistence, and the regulatory effect of negative fitness is minimised by the fast rates of cultural transmission (i.e.
 individual shift strategy before experiencing a population decline resulting from density being over $K$).
@@
 \subsection{Differences in critical density and carrying capacities}
-% XAVI better title?
 
@@
 the $z$ parameter) has major effects on the final equilibria (c.f.
-fig.\ref{fig:bifurcationWithTransmission}).
+fig.
+XXX?).
 However, this will also depend on the key parameters defining the range of positive payoff for the two strategies (i.e.
@@
 While a comprehensive analysis is not possible, here we illustrate the relationship between these parameters and $z$ by introducing a new parameter $\lambda \in (0,1)$ adding the following constraints to equation \eqref{eq1}:
-%I slightly revised your intro here. Mainly because this reads as if changes depends on the intervals of positive payoff alone, while I think it actually depends on any possibly combination of values.
 
@@
 
-The dynamics of two strategies are similar for low values of $\lambda$.
-High values indicate that the minimum number of individuals needed to achieve positive fitness in $b$ gets closer to $K_a$, while the population size that $b$ can sustain also grows.
-
-As a consequence, the number of individuals switching strategies when $b$ has lower payoff is also higher and could potentially overload $a$ due to its lower $K_a$.
-
-Figure \ref{fig:overlap} shows how the degree of overlap affects the equilibria of our model with and without social learning.
-The relation between the two processes strongly affects the extent of the different basins in a non-dynamic fashion.
-%I've put "extent" instead of "shape" as we are not directly assessing the latter. Also I would omit "strongly".
-
-% XAVI put high res figure, and update left, middle, right order
+Equation \eqref{eqOverlap} ensures that the difference between carrying capacity and critical density is hold constant and equal between the two variants, so that $\lambda$ can be interpreted as an inverse  measure of overlap.
+Indeed when $\lambda=1$ the two variants are identical, and when $\lambda=0$, the critical density of one strategy becomes equal to the carrying capacity of the other (i.e.
+$A_b=K_a$).
+Figure \ref{fig:overlap} examines three different settings of $\lambda$  with low ($z=0.05$) and intermediate ($z=0.2$) degree of reliance on social learning.
+The plot suggests that, higher values of $z$ limits the number of initial conditions within the positive payoff zone (i.e.
+when the initial conditions for $a$ and $b$ are within the intervals $[A_a,K_a]$ and $[A_b,K_b]$) where a stable co-existence is achieved, but the magnitude of this effect is smaller for lower values of $\lambda$ (i.e.
+smaller overlap).
+We can further investigate this point by computing the relative areas of different basins of attraction within the positive payoff ozone for different parameter combinations of $z$ and $\lambda$.
+The results (fig.
+\ref{fig:percentages}) confirms that indeed with high overlap and high rates of social learning conditions enabling the stable coexistence is reduced, initially as an increase of conditions promoting the dominance of one variant or another, and eventually as the unstable fluctuations regime take over.
+With decreasing $\lambda$ unstable regimes vanish, and the relationship between $z$ and the size of the basin of attraction for stable coexistence becomes more or less constant for each value of $\lambda$.
+% I think we should use different parameter combinations for this second plot.
+Basically we have lambda=0.95 and lambda=0.9 being identical, and lambda=0.6~0.5 also identical.
+Ideally we should also think about slightly more refined interval for z as well?
 
@@
 \end{figure*}
-
-Figure \ref{fig:percentages} explores in deeper detail the effects of both parameters ($z$ and $lambda$) on the basin of each equilibria.
-An increase in $z$ decreases the possibility of having coexistence for high similarity between strategies $a$ and $b$ (i.e.
-$\lambda \geq 0.9$.
-From $0.9 < \lambda < 0.55$ the dynamics remain remarkably similar, while for $\lambda = 0.5$ the effects of social learning decreases because of the large gap between $A$ and $K$ of both strategies.
 
@@
 
+%This section needs to be revised:
 Figure \ref{fig:percentages} also suggest that with decreasing values of $\lambda$, the variant with lower carrying capacity/critical density expands its basin of attraction.
@@
 Larger values of any of the $c$ parameters dictates a stronger influence of the other population.
+
+%We need to mention that analytical equilibrium solution and put the equations here. Then discuss about the effects for different values of z.
 

### 2015-06-12 manuscript/draft.tex
updated plots

@@
   \centering
-      \includegraphics[width=2\columnwidth]{./figures/figure3.pdf}
+      \includegraphics[width=\columnwidth]{./figures/figure3.pdf}
   \caption{Basins of attraction for the four equilibria (mutual extinction: grey; extinction of $a$: blue; extinction of $b$: red; stable coexistence: green) with (right) and without (left) social learning.
-Model parameters: $A_a=0.2$; $K_a=0.8$; $A_b=0.4$; $K_b=1.0$; $r_a=r_b=0.05$; $z=0$ (left panel) and $z=0.05$ (right panel) }
+Model parameters: $A_a=0.2$; $K_a=0.8$; $A_b=0.35$; $K_b=0.95$; $r_a=0.05$; $z=0$ (left panel) and $z=0.2$ (right panel) }
     \label{fig:TransmissionBasin}
 \end{figure*}
-%Is it possible to change this figure to the settings for lambda=0.75 in figure5.pdf, setting as z 0 and 0.2?
-This way we get a much stronger changes in the basin of attraction, and we could also mention here point 6 above?
 
@@
 With decreasing $\lambda$ unstable regimes vanish, and the relationship between $z$ and the size of the basin of attraction for stable coexistence becomes more or less constant for each value of $\lambda$.
-% I think we should use different parameter combinations for this second plot.
-Basically we have lambda=0.95 and lambda=0.9 being identical, and lambda=0.6~0.5 also identical.
-Ideally we should also think about slightly more refined interval for z as well?
 
@@
   \caption{Proportion of initial conditions falling in each basin of attraction for different values of $\lambda$.
-X axis controls the parameter $z$ defining social learning.} %We need to add what is the range of initial conditions being explored and what are the initial values of A_a and K_a
+X axis controls the parameter $z$ defining social learning.
+Model parameters: $A_a=0.2$; $K_a=0.8$; $r_a=0.05$}
     \label{fig:percentages}

### 2015-06-12 experimentMN3/alleeTasks.py

@@ -9,21 +9,24 @@ def main():
     zValues = [0.05, 0.2]
     lambdaValues = [0.5, 0.75, 1]
     """
-    """
+    
     # 2 - exploration lambda/z (relative figure)
     competitionValues = [0.0]
-    zValues = arange(0,1.1,0.1)
-    lambdaValues = arange(0.5,1.05,0.05)
+    zValues = arange(0,0.55,0.05)
+    lambdaValues = arange(0.35,1.05,0.05)
+    
     """
     # 3 - exploration competition/z 
     competitionValues = [0.1,0.25,0.5]
     zValues = [0.05, 0.2]
     lambdaValues = [0.75]
     """
+
+    """
     # 4 - z (figure 3)
     competitionValues = [0.0]
-    zValues = [0,0.05]
-    lambdaValues = [2/3]
+    zValues = [0,0.2]
+    lambdaValues = [0.75]
     """
 
     tasksFile = open('allee.tasks', 'w')

### 2015-06-12 manuscript/figures/figure3.R

@@ -12,10 +12,10 @@ abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
 axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
 axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
-trajPlot(ini.m=0.22,ini.n=0.6,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
-trajPlot(ini.m=0.61,ini.n=0.42,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
-trajPlot(ini.m=0.95,ini.n=0.39,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
+trajPlot(ini.m=0.25,ini.n=0.5,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
 trajPlot(ini.m=0.5,ini.n=0.6,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
+trajPlot(ini.m=0.61,ini.n=0.36,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
+trajPlot(ini.m=0.95,ini.n=0.32,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
 
 text(0.2480076,0.6390929,labels="i",col=1,font=2)
 text(0.6483907,0.4705729,labels="k",col=1,font=2)

### 2015-06-12 manuscript/figures/figure6.R

@@ -28,7 +28,6 @@ getRawValues <- function(run)
     print(sprintf('total: %02d - should be: %02d',total, nrow(result)))
 
     resultdf <- data.frame(eq='coexistence', value=nrow(coexistence))
-    resultdf <- rbind(resultdf, data.frame(eq='extinction', value=nrow(extinction)))
     resultdf <- rbind(resultdf, data.frame(eq='a Extinct', value=nrow(mExtinct)))
     resultdf <- rbind(resultdf, data.frame(eq='b Extinct', value=nrow(nExtinct)))
     resultdf <- rbind(resultdf, data.frame(eq='unstable', value=(nrow(result)-total)))
@@ -48,19 +47,18 @@ getPercentages<- function(raw)
 }
 
 raw <- data.frame()
-for(run in 0:77)
+for(run in 0:153)
 {
     raw <- rbind(raw, getRawValues(run))
 }
 
 percentages <- getPercentages(raw)
-percentages <- subset(percentages, lambda<=1.0)    
-colorValues = c('#006400','#D3D3D3','#4169E1','#CD5C5C','gold2')
+colorValues = c('#006400','#4169E1','#CD5C5C','gold2')
 
 #ggplot(percentages, aes(x=z, y=value, fill=eq)) + geom_area(stat='identity') + scale_fill_manual(name='Equilibria', values=colorValues)+ facet_wrap(~lambda)
 
-plotted <- subset(percentages, lambda >=0.9 | lambda<=0.6 )
-plotted$lambda <- factor(plotted$lambda, levels=c(1,0.95,0.9,0.6,0.55,0.5))    
+plotted <- subset(percentages, lambda >=0.85 | lambda==0.5 | lambda==0.35 )
+plotted$lambda <- factor(plotted$lambda, levels=c(1,0.95,0.9,0.85,0.5,0.35))    
 
 # coexistence: #006400
 # extinct: #D3D3D3

### 2015-06-13 manuscript/figures/figure3.R
updated texts for fig3

@@ -14,13 +14,13 @@ axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
 axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
 trajPlot(ini.m=0.25,ini.n=0.5,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
 trajPlot(ini.m=0.5,ini.n=0.6,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
-trajPlot(ini.m=0.61,ini.n=0.36,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
+trajPlot(ini.m=0.61,ini.n=0.38,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
 trajPlot(ini.m=0.95,ini.n=0.32,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
 
-text(0.2480076,0.6390929,labels="i",col=1,font=2)
-text(0.6483907,0.4705729,labels="k",col=1,font=2)
-text(0.9674449,0.4328860,labels="l",col=1,font=2)
+text(0.26,0.48,labels="i",col=1,font=2)
 text(0.5,0.65,labels="j",col=1,font=2)
+text(0.58,0.39,labels="k",col=1,font=2)
+text(0.98,0.32,labels="l",col=1,font=2)
 
 load('../../experimentMN3/00_fig3/result_001.Rdata')
 title=substitute(paste(z == i), list(i=z))
@@ -29,14 +29,16 @@ abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
[…]

### 2015-06-17 manuscript/draft.tex
update

@@
 \begin{abstract}
-Several social learning strategies are guided by an evaluation of alternative cultural variants by means of the expected payoffs associated by these.
-The assumption of the social learner is that the payoff received by the transmitter while expressing a given cultural trait, is a proxy for for inferring the qualitative features of the trait itself.
-Here we explore the theoretical implications when payoff directly affects reproductive fitness and it is density dependent, that is its expected amount is a function of the number of individuals possessing the same trait.
-More specifically we consider situations where the density dependence is simultaneously direct and inverse, whereby the highest payoff is achieved at intermediate densities.
-This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits that are enhanced by cooperation or mutual facilitation, but are somehow constrained by some carrying capacity.
-We  explores the evolutionary dynamics of a single trait model with two variants, each exhibiting a different Allee effect.
+Several social learning strategies guide the selection of cultural variant to imitate through an evaluation of the expected benefits associated with each observed alternatives.
+The assumption of the social learner is that the payoff received by the transmitter while expressing a given cultural variant is a proxy for for inferring the qualitative features of the variant itself.
+Here we explore the theoretical implications when: 1) the payoff attributed to a given cultural variant directly affects the reproductive success of its bearer;  and 2) payoff is density dependent, that is its expected amount is a function of the number of individuals possessing the same variant.
+We consider situations where such density dependence is simultaneously direct and inverse, with minimum and maximum critical thresholds and a peak in the payoff achieved at intermediate densities.
+This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits that are enhanced by cooperation or mutual facilitation, but are somehow constrained by some form of carrying capacity.
+We explore the evolutionary dynamics of a single trait-two variants model, each exhibiting a different Allee effect.
 We used a modified Lotka-Volterra model of payoff-biased social learning, and show that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion, stable co-existence, and fixation to single variant.
@@
 Early works inspired by population biology \citep{cavallisforza_feldman_1981,boyd1985} have laid the foundation of a now mature trans-disciplinary field which unites the social and biological sciences \citep{mesoudi_etal_2006}.
-The rich cross-fertilisation between different fields has fomented the designing of new analytical techniques as well as new usage for existing methods, but more crucially have prompted the development of theoretical models of how culture is acquired and transmitted between individuals.
+The rich cross-fertilisation between different subfields has promoted the designing of new analytical techniques as well as new uses for existing methods, but more crucially have prompted the development of theoretical models of how culture is acquired and transmitted between individuals.
 %Add reference?
@@
 A central element, shared by most models of social learning, is the existence of some heuristics that promotes the selection of a given cultural variant over another \citep{laland2004}.
-This can be based on some properties of the variant itself (\emph{content bias}) or derived from contextual cues (\emph{context bias}), such as the commonness or the rareness of a trait (\emph{conformist bias} and \emph{anti-conformist bias}), or some features associated with the transmitter such as prestige, success, or similarity to the learner (\emph{context bias})\citep{henrich_mcelreath2003}.
+This can be based on some properties of the variant itself (\emph{content bias}) or derived from contextual cues (\emph{context bias}), such as the commonness or the rareness of a trait (\emph{conformist bias} and \emph{anti-conformist bias}), or  features associated with the transmitter such as prestige, success, or similarity to the learner (\emph{context bias})\citep{henrich_mcelreath2003}.
 
@@
 Here the probability of copying a cultural variant $i$ is directly proportional to a payoff $s$ observed amongst individuals possessing $i$.
-In other words the social learner assumes that the success of an individual (measured in $s$) is to some extent function of the observed behavioural trait, that is $s=f(i)$.
-Several variants of this learning strategies have been proposed and explored, including: \emph{imitate the best}, where the social learner identifies and copies the trait possessed by the individual with the highest payoff; \emph{compare means}, where the social learner compare the average payoff of different strategies and chooses the highest variant; and the \emph{proportional imitation rule}, where the probability of adopting a variant is proportional to the difference in payoff between the demonstrator and the learner~\citep{schlag1998,baldini2013}.
+In other words the social learner assumes that the success of an individual (measured in $s$) is a function of the observed behavioural trait, that is $s=f(i)$.
+Several variants of this learning strategies have been proposed and explored, including: \emph{imitate the best}, where the social learner identifies and copies the trait possessed by the individual with the highest payoff; \emph{compare means}, where the social learner compare the average payoff of different strategies and chooses the highest variant; and the \emph{proportional imitation rule}, where the probability of adopting a variant is proportional to the difference in payoff between the demonstrator and the learner~\citep{schlag1998,baldini2013,crema_lake_inpress}.
 
 Whilst most studies focused on how difference in these imitation rules can drive the dynamics of the system, less attention has been given to the relationship between target cultural traits and their payoff signals.
-Typical signal are most likely some indirect proxy of success (e.g.
-number of offspring, income, presence/absence of other traits associated with success or prestige, etc.)  that are assumed, by the social learner, to be more or less linked with a cultural trait.
-In most cases the causal link between the payoff and a cultural trait is just inferred, and the exact processes and variables leading the latter to the former is most likely unknown.
+Typical signals are most likely some indirect proxy of success (e.g.
+number of offspring, income, presence/absence of other traits associated with success or prestige, etc.)  that is assumed (by the social learner) to be more or correlated with intrinsic properties of the cultural variant.
+In most cases the causal link between the payoff and a cultural trait is just inferred, and the exact processes and variables linking the two is unknown.
 This implies that the strength of correlation between payoff and cultural trait can vary, introducing different degrees of uncertainty that can have a strong impact to the efficiency of specific learning rules.
-For example,~\citet{crema_lake_inpress} have shown that \emph{imitate the best} is becomes increasingly detrimental with high payoff uncertainty and a large pool of potential social teachers.
-This is because innovations are by definition rare in a population, and hence even if their average payoff are higher than extant suboptimal alternatives, their probability of signalling the highest payoff, and hence be selected, is low.
+For example,~\citet{crema_lake_inpress} have shown that \emph{imitate the best} is increasingly detrimental with high payoff uncertainty and a large pool of potential social teachers.
+This is because innovations are initially rare in a population, and hence even if their average payoff are higher than extant suboptimal alternatives, their probability of signalling the highest payoff, and hence being selected, is low.
 
-This uncertainty in the payoff signal is derived by the presence of interacting cultural and physical traits, as well as the environment where the trait is manifest.
+The uncertainty in the payoff signal is derived by the presence of interacting cultural and physical traits, as well as the environment where the trait is manifest.
 Thus the function $s=f(i)$ should be better described as $s=f(i,\epsilon)$, where $\epsilon$ represent all additional hidden and latent variables that contributes, along with the cultural variant of interest $i$, to the payoff signal $s$.
@@
 Several behavioural traits are expected to show such density dependence.
-In particular, behaviour linked to the exploitation of limited resources are expected to exhibit density dependence.
-Novel optimal variants are likely to spread rapidly when the trait is rare, but once a threshold frequency is exceeded (i.e.
-the carrying capacity),  beneficial effect will start to decline, the payoff will become smaller, and eventually rarer suboptimal variants can be preferred (see Lake and Crema \citeyear{lake_and_crema_2012} for an extensive analysis).
+For example, behaviour linked to the exploitation of some limited resources are expected to exhibit decline when the density exceeds a critical threshold (i.e carrying capacity).
+Thus novel optimal variants are likely to spread rapidly when the trait is rare, but as it approaches carrying capacity, the  beneficial effects will start to decline, the payoff will become smaller, and eventually rarer suboptimal variants might be preferred (see Lake and Crema \citeyear{lake_and_crema_2012} for an extensive analysis).
 

### 2015-06-18 manuscript/draft.tex
updated reference

--- 2015-06-12 manuscript/draft.tex
@@
 Early works inspired by population biology \citep{cavallisforza_feldman_1981,boyd1985} have laid the foundation of a now mature trans-disciplinary field which unites the social and biological sciences \citep{mesoudi_etal_2006}.
-The rich cross-fertilisation between different fields has fomented the designing of new analytical techniques as well as new usage for existing methods, but more crucially have prompted the development of theoretical models of how culture is acquired and transmitted between individuals.
-%Add reference?
+The rich cross-fertilisation between different fields has fomented the designing of new analytical techniques as well as new usage for existing methods, but more crucially have prompted the development of theoretical models of how culture is acquired and transmitted between individuals~\citep{mesoudi_cultural_2015}.
 

### 2015-06-18 manuscript/draft.tex

--- 2015-06-17 manuscript/draft.tex
@@
 Early works inspired by population biology \citep{cavallisforza_feldman_1981,boyd1985} have laid the foundation of a now mature trans-disciplinary field which unites the social and biological sciences \citep{mesoudi_etal_2006}.
-The rich cross-fertilisation between different subfields has promoted the designing of new analytical techniques as well as new uses for existing methods, but more crucially have prompted the development of theoretical models of how culture is acquired and transmitted between individuals.
-%Add reference?
+The rich cross-fertilisation between different subfields has promoted the designing of new analytical techniques as well as new uses for existing methods, but more crucially have prompted the development of theoretical models of how culture is acquired and transmitted between individuals~\citep{mesoudi_cultural_2015}.
 

### 2015-06-20 manuscript/draft.tex
Minor Update

--- 2015-06-17 manuscript/draft.tex
@@
 
-Several behavioural traits are expected to show such density dependence.
-For example, behaviour linked to the exploitation of some limited resources are expected to exhibit decline when the density exceeds a critical threshold (i.e carrying capacity).
-Thus novel optimal variants are likely to spread rapidly when the trait is rare, but as it approaches carrying capacity, the  beneficial effects will start to decline, the payoff will become smaller, and eventually rarer suboptimal variants might be preferred (see Lake and Crema \citeyear{lake_and_crema_2012} for an extensive analysis).
+Several behavioural traits are expected to show such a density dependence.
+For example, behaviour linked to the exploitation of limited resources will exhibit a decline in payoff when density exceeds a critical threshold (i.e carrying capacity).
+In this case a new optimal variant is likely to spread rapidly when the it is rare, but as the number of individuals possessing it approaches carrying capacity, its beneficial effects will start to decline, the payoff will decrease, and eventually rarer suboptimal variants might be preferred (see Lake and Crema \citeyear{lake_and_crema_2012} for an extensive analysis).
 

### 2015-06-20 manuscript/figures/figure4.R
Update
-Changed bifurcation plot with transmission

@@ -22,7 +22,7 @@ y.lim=c(0,1.2)
 
 bifurcationPlotSingle<-function(Zseq=seq(0,0.7,0.005),timeSteps=10000,...)
 {
-    plot(x=1,y=1,xlim=c(min(Zseq),max(Zseq)),type="n",ylab="a",xlab="z",ylim=c(0,1.2))
+    plot(x=1,y=1,xlim=c(min(Zseq),max(Zseq)),type="n",ylab="a",xlab="z",ylim=c(0.1,1.1))
     for (z in Zseq)
         {
             tmp=unique(allee(timeSteps=timeSteps,storeFinalOnly=FALSE,z=z,ini.m=0.4,ini.n=0.4,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm)[[1]][(timeSteps-500):timeSteps])
@@ -30,8 +30,11 @@ bifurcationPlotSingle<-function(Zseq=seq(0,0.7,0.005),timeSteps=10000,...)
                }
         }
 
-bifurcationPlotSingle(Zseq=seq(0,0.7,0.005),timeSteps=10000,ini.m=0.4,ini.n=0.4,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm)
-bifurcationPlotSingle(Zseq=seq(0,0.7,0.005),timeSteps=10000,ini.m=0.7,ini.n=0.4,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm)
+bifurcationPlotSingle(Zseq=seq(0.4,0.82,0.001),timeSteps=10000,ini.m=0.4,ini.n=0.4,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm)
+
+
+
+
 
 
 

### 2015-06-25 src/basin.R
Update on plot functions

@@ -90,14 +90,16 @@ trajPlot<-function(ini.m=1001,ini.n=1001,...)
         points(x=ini.m,y=ini.n,pch=20)
     }
 
-basePlot<-function(...)
+basePlot<-function(leg=TRUE,ymax=NA,...)
     {
         tmp=allee(...)
-        plot(tmp$m,type="l",ylim=c(0,max(unlist(tmp))),col="indianred",lwd=2,ylab="Population Size",xlab="Time")
-        axis(side=4,at=c(Km,Kn),labels=c(expression(K[m]),expression(K[n])),hadj=0.5,las=2)
+        if(is.na(ymax)){ymax=max(unlist(tmp))}
+        plot(tmp$m,type="l",ylim=c(0,ymax),col="indianred",lwd=2,ylab="Population Size",xlab="Time")
+        axis(side=4,at=c(Km,Kn),labels=c(expression(K[a]),expression(K[b])),hadj=0.5,las=2)
         abline(h=c(Km,Kn),lty=2)
         lines(tmp$n,col="royalblue",lwd=2)
-        legend("bottomright",legend=c("m","n"),lty=1,lwd=2,col=c("indianred","royalblue"))
+        if(leg==TRUE)
+        {legend("bottomright",legend=c("a","b"),lty=1,lwd=2,col=c("indianred","royalblue"))}
     }
 
 

### 2015-06-25 manuscript/figures/figure4.R
Figure 4 new (two) versions

@@ -20,48 +20,30 @@ y.lim=c(0,1.2)
 
 
 
-bifurcationPlotSingle<-function(Zseq=seq(0,0.7,0.005),timeSteps=10000,...)
+bifurcationPlotSingle<-function(Zseq=seq(0,0.7,0.005),timeSteps=10000,cexSetting=6,...)
 {
-    plot(x=1,y=1,xlim=c(min(Zseq),max(Zseq)),type="n",ylab="a",xlab="z",ylim=c(0.1,1.1))
+    plot(x=1,y=1,xlim=c(min(Zseq),max(Zseq)),type="n",ylab="",xlab="",ylim=c(0.1,1.1),cex.axis=cexSetting,cex.lab=cexSetting)
     for (z in Zseq)
         {
             tmp=unique(allee(timeSteps=timeSteps,storeFinalOnly=FALSE,z=z,ini.m=0.4,ini.n=0.4,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm)[[1]][(timeSteps-500):timeSteps])
             points(x=rep(z,length(tmp)),y=tmp,pch=20,cex=0.5)
-               }
+        }
+mtext(side=1,"z",line=6,cex=6)
+mtext(side=2,"a",line=6,cex=6)
         }
 
+par(mar=c(5, 6, 4, 2) + 2)
 bifurcationPlotSingle(Zseq=seq(0.4,0.82,0.001),timeSteps=10000,ini.m=0.4,ini.n=0.4,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm)
 
 
-
-
-
-
-
-
-cexSetting=10
-bifurcationPlotDouble<-function(Zseq=seq(0,0.7,0.005),timeSteps=10000,cexSetting,...)
-{
-    plot(x=1,y=1,xlim=c(min(Zseq),max(Zseq)),type="n",ylab="",xlab="z",ylim=c(0,1.2),cex.axis=cexSetting,cex.lab=cexSetting)
-    for (z in Zseq)
-        {
-            tmp=allee(timeSteps=timeSteps,storeFinalOnly=FALSE,z=z,...)
[…]

### 2015-06-25 manuscript/draft.tex
Complete Draft Update

@@
 \documentclass[review,authoryear]{elsarticle}
-% \documentclass[3p,authoryear,twocolumn]{elsarticle} %Just to give a feel of how it would look like
 
@@
 \title{Allee Effect and Cultural Evolution}
-%\title{The evolution of payoff-biased learning under Allee Effect} %Perhaps a
 
@@
 Several social learning strategies guide the selection of cultural variant to imitate through an evaluation of the expected benefits associated with each observed alternatives.
-The assumption of the social learner is that the payoff received by the transmitter while expressing a given cultural variant is a proxy for for inferring the qualitative features of the variant itself.
+The assumption of the social learner is that the payoff received by the transmitter while expressing a given cultural variant is a proxy for  inferring the qualitative features of the variant itself.
 Here we explore the theoretical implications when: 1) the payoff attributed to a given cultural variant directly affects the reproductive success of its bearer;  and 2) payoff is density dependent, that is its expected amount is a function of the number of individuals possessing the same variant.
-We consider situations where such density dependence is simultaneously direct and inverse, with minimum and maximum critical thresholds and a peak in the payoff achieved at intermediate densities.
+We consider situations where such density dependence is simultaneously direct and inverse, with minimum and maximum critical thresholds bounding the range of positive payoff, with a peak value at midpoint.
 This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits that are enhanced by cooperation or mutual facilitation, but are somehow constrained by some form of carrying capacity.
 We explore the evolutionary dynamics of a single trait-two variants model, each exhibiting a different Allee effect.
-We used a modified Lotka-Volterra model of payoff-biased social learning, and show that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion, stable co-existence, and fixation to single variant.
+We used a modified Lotka-Volterra model of payoff-biased social learning, and show that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion, stable co-existence, and fixation.
 
@@
 In other words the social learner assumes that the success of an individual (measured in $s$) is a function of the observed behavioural trait, that is $s=f(i)$.
-Several variants of this learning strategies have been proposed and explored, including: \emph{imitate the best}, where the social learner identifies and copies the trait possessed by the individual with the highest payoff; \emph{compare means}, where the social learner compare the average payoff of different strategies and chooses the highest variant; and the \emph{proportional imitation rule}, where the probability of adopting a variant is proportional to the difference in payoff between the demonstrator and the learner~\citep{schlag1998,baldini2013,crema_lake_inpress}.
+Several variants of this learning strategies have been proposed and explored, including: \emph{imitate the best}, where the social learner identifies and copies the trait possessed by the individual with the highest payoff; \emph{compare means}, where the social learner compare the average payoff of different strategies and chooses the highest; and the \emph{proportional imitation rule}, where the probability of adopting a variant is proportional to the difference in payoff between the demonstrator and the learner~\citep{schlag1998,baldini2013,crema_lake_inpress}.
 
-Whilst most studies focused on how difference in these imitation rules can drive the dynamics of the system, less attention has been given to the relationship between target cultural traits and their payoff signals.
+Whilst most studies focused on how difference in these imitation rules can drive the evolutionary dynamics of the cultural system, less attention has been given to the relationship between target cultural traits and their payoff signals.
 Typical signals are most likely some indirect proxy of success (e.g.
-number of offspring, income, presence/absence of other traits associated with success or prestige, etc.)  that is assumed (by the social learner) to be more or correlated with intrinsic properties of the cultural variant.
+number of offspring, income, presence/absence of other traits associated with success or prestige, etc.)  that is assumed (by the social learner) to be, to some extent,  correlated with intrinsic properties of the cultural variant.
 In most cases the causal link between the payoff and a cultural trait is just inferred, and the exact processes and variables linking the two is unknown.
@@
 The uncertainty in the payoff signal is derived by the presence of interacting cultural and physical traits, as well as the environment where the trait is manifest.
-Thus the function $s=f(i)$ should be better described as $s=f(i,\epsilon)$, where $\epsilon$ represent all additional hidden and latent variables that contributes, along with the cultural variant of interest $i$, to the payoff signal $s$.
+Thus the function $s=f(i)$ should be better described as $s=f(i,\epsilon)$, where $\epsilon$ represent all additional hidden and latent variables that contribute, along with the cultural variant of interest $i$, to the payoff signal $s$.
 A trivial example is a fisherman who might evaluate the performance of a new hook (the target cultural trait) using as a cue the number of fish captured (the payoff signal).
-The correlation between the two is partly derived by the efficiency of the hook, but also by the rod, the choice of the bait, the availability of the prey species, the choice of the fishing spot, the physical strength of the fisherman, and chance.
+The correlation between the two is partly derived by the efficiency of the hook, but also by the rod, the choice of the bait, the amount of available prey, the choice of the fishing spot, the physical strength of the fisherman, and chance.
 In many cases we should expect that the social teacher and learner share many of these interacting traits as well as the environment where these are displayed.
@@
 One of these are contexts where the success of a variant is strongly correlated with the absolute number of individuals $n_i$ possessing the same trait ($i$) within a population.
-This density dependence, which can be described by the payoff function $f(i,n_i)$, implies that the adoption (or abandonment) of a trait can alter its expected payoff-signal, possibly generating a variety of feedback mechanism.
+This density dependence, which can be described by the payoff function $f(i,n_i)$, implies that the adoption (or abandonment) of a trait can alter its expected payoff-signal, potentially causing a cascade of feedback effects.
 
@@
 For example, behaviour linked to the exploitation of limited resources will exhibit a decline in payoff when density exceeds a critical threshold (i.e carrying capacity).
-In this case a new optimal variant is likely to spread rapidly when the it is rare, but as the number of individuals possessing it approaches carrying capacity, its beneficial effects will start to decline, the payoff will decrease, and eventually rarer suboptimal variants might be preferred (see Lake and Crema \citeyear{lake_and_crema_2012} for an extensive analysis).
+In this case a new optimal variants are likely to spread rapidly when the they are rare, but as the number of individuals possessing these approach their respective carrying capacities, their beneficial effects will start to decline, the payoff  decrease, and eventually rarer suboptimal variants might be preferred (see Lake and Crema \citeyear{lake_and_crema_2012} for an extensive analysis).
 
-However, density dependence does not necessarily imply only a negative correlation between population size and payoff.
+Density dependence does not, however, necessarily imply only a negative correlation between population size and payoff.
 In many ecological contexts per capita growth rate is positively correlated with population density, thus resulting into an inverse density dependence.
@@
 promote positive niche construction \citep{vandermeer_2008}, and requires a critical mass \citep{rogers_2003} are likely to exhibit some degree of Allee effect.
-This is thus not limited necessarily to the domain of traits directly promoting reproductive fitness, and hence can include language~\citep{kandler2009}, organisational populations~\citep{caroll_and_hannan_1989} and communication technologies \citep{van_slyke_perceived_2007}.
+The implication is thus not limited necessarily to the domain of traits directly promoting reproductive fitness, and hence can include language~\citep{kandler2009}, organisational populations~\citep{caroll_and_hannan_1989} and communication technologies \citep{van_slyke_perceived_2007}.
 
@@
 
-The reproductive growth rate $R$ of each population is defined by the following set of equations:
+The reproductive growth rate $R$ of each population is defined by the following pair of equations:
 
@@
 \begin{aligned}
-R_{a,t}& = r_a (\frac{a_t}{A_a}-1)(1-\frac{a_t}{K_a})\\
-R_{b,t}& = r_b (\frac{b_t}{A_b}-1)(1-\frac{b_t}{K_b}) 
+R_{a,t}& = r_a \left(\frac{a_t}{A_a}-1\right)\left(1-\frac{a_t}{K_a}\right)\\
+R_{b,t}& = r_b \left(\frac{b_t}{A_b}-1\right)\left(1-\frac{b_t}{K_b}\right) 
 \label{eq2}
@@
 
-Equation \eqref{eq2} depicts the so called \emph{strong} Allee effect \citep{Wang_and_Kot_2001}, where the reproductive rate is negative when population density is either below $A$ or above $K$, and  in this case reach its maxima $R_{max}=\frac{r(A-K)^2}{4AK}$ when density is equal to $\frac{(K-A)}{2}$.
+Equation \eqref{eq2} depicts the so called \emph{strong} Allee effect \citep{Wang_and_Kot_2001}, where reproductive rate is negative when population density is either below $A$ or above $K$, and in this case reach its maxima $R_{max}=\frac{r(A-K)^2}{4AK}$ when density is equal to $\frac{(K-A)}{2}$.
 
@@
 
-%I attached the list of results that I sent you over the email that we need  to put in the text
-
-%1)With no Allee effect the strategy with the highest payoff wins  /// This is obvious we don?t need to prove this 
-%2) With carrying capacity, but no Allee effect we get possible cases of reversions.
-//Lake and Crema explores this, but the model has also uncertainty in the payoff signal so not sure if it is technically comparable?
-Should we do something with this?
-%
-%
-%** Expected pattern without transmission **
-%-3)We have our unstable nodes, stable nodes, and extinction. As long as the strategies are above their critical densities, they will survive. // DONE
-%
-%** Expected pattern with transmission **
-%4) There is a possibility for a trait to not get extinct even if its below its critical density.
-This happens only if: a) the alternative strategy is also below the critical density and its distance from A is further b) the alternative strategy is beyond carrying capacity.
-If the alternative strategy is in the positive payoff interval the focal strategy will go extinct /// This is shown with all the basin plots,  so no need to anything extra.
-%
-%5) If both traits are above their respective critical densities, we expect stable coexistence when there is no transmission, but when z>0, strategies that are closer to their critical densities have a non-zero probability to go extinct.
-This probability increases as the alternative population approaches its own K or A.
-/// This is shown by the basin plots, so its fine.
-%
-%6) As a corollary to point 5, with increasing z (but also as a function of lambda), the shared perimeter of the boundary of the basin of attractions of the two strategies increases in proximity to the both critical densities.
-This means that small changes in behaviour (z) or external fluctuations can promote the adoption of one or the other but not the both.
-// This is hard to prove I guess, and not sure if the basin plots are sufficient to state this.
-%
-%7) If both variants have initial conditions within their positive payoff density ranges, a decrease in lambda promotes the coexistence of both strategy for a given value of z.
-Thus with increasing ?distance?
-between the two strategies, its less likely that we observe the dominance of one or the other.
-We ought to find out why this is the case though, perhaps we need to look at individual time-series to have some clues.
-// We should be able to see this with the overlap analysis + possibly individual time-series to get some grips
-%
-%8) With increasing values of z, the basin of coexistence is decreased to the expense of limit cycles (can we check if it is non-chaotic using Lyapunov?) and extinction of one strategy or the other.
-The growth is however asymmetric (maybe), and the strategy with high A and K have will have a smaller proportion of initial conditions where it can dominate.
-// overlap analysis should, hopefully, still show this.
-%
-%9) If point 8 is true, we can then argue that transition to strategies with higher carrying capacity is difficult, and entails in many case stable or unstable forms of coexistence.
-The only exception is when the two strategies is interacting //competition analysis, with competition values and two transmission settings.
-%
-%10) or the critical densities are the same // just a single basin plot would suffice.
-
 We investigate the equilibrium conditions for different initial population sizes of $a$ and $b$ and different settings of $z$, $A$, and $K$, ensuring that in all cases $K_b \geq K_a$.
@@
   } 
-that ensures that the two variants have always the same $R_{max}$.
-Given that equations \ref{eq1}-\ref{eq4} have no analytical solution, we identified equilibrium conditions via simulation (source codes of the simulation can be found on the following repository \textcolor{red}{PUT LINK HERE})
+which ensures that the two variants have always the same $R_{max}$.
+Given that equations \ref{eq1}-\ref{eq4} have no analytical solution that we are aware of, we identified equilibrium conditions via simulation (source codes of the simulation can be found on the following repository \textcolor{red}{PUT LINK HERE})
 
@@
 When social learning is enabled (i.e.
-when $z>0$), the stability of the nodes are perturbed by individuals switching from one strategy to another.
-We identify two major changes in the equilibrium properties of the system as a result of cultural transmission.
+when $z>0$), the stability of the nodes are perturbed by individuals switching from one strategy to another, leading to two major changes in the equilibrium properties of the system.
 
-First, the shape of the basins of attraction are altered, and hence initial conditions closer to the boundaries between one basin and another will be strongly affected,  leading occasionally to very different evolutionary equilibria.
+First, the basins of attraction are transformed, and hence the system equilibrium at $z=0$ and $z>0$ can be different for the same initial conditions.
+This is particularly the case for combinations of $a$ and $b$ that are closer to the boundaries between two or more basins.
 Figure \ref{fig:TransmissionBasin} shows an example of this.
@@
   \caption{Basins of attraction for the four equilibria (mutual extinction: grey; extinction of $a$: blue; extinction of $b$: red; stable coexistence: green) with (right) and without (left) social learning.
-Model parameters: $A_a=0.2$; $K_a=0.8$; $A_b=0.35$; $K_b=0.95$; $r_a=0.05$; $z=0$ (left panel) and $z=0.2$ (right panel) }
+Model parameters: $A_a=0.2$; $K_a=0.8$; $A_b=0.35$; $K_b=0.95$; $r_a=0.05$;$r_b=0.1039063$; $z=0$ (left panel) and $z=0.2$ (right panel) }
     \label{fig:TransmissionBasin}
@@
 \ref{fig:bifurcationDiagram}), that is a decrease in the stability of the fixed points.
-Figure XXXX shows an example where the increase in $z$ generates a transition from stable coexistence to cyclical fluctuations.
-%We need to make this figure as time-series, or actually use the bifurcation diagram.
-This regime is prone to occur when both variants are initialised with high densities.
-%This refers to the basin plot figureUnstable1.pdf that you've already made.
-Given the description below, do you think that those tiny "islands" of orange inside the green are actually specific values that are covered by the fluctuations in the time-series?
-Given that the system is deterministic, it wouldn't be a surprise, if anything there MUST be a combination of values for these...
+Figure \ref{fig:BifurcationSocialLearning} shows the bifurcation diagram for the population size of $a$ as a function of $z$.
+Here initial the initial effects of $z$ does not alter the equilibrium point ($K_a=0.6$), but once $z>0.5$ we observe a sudden transition to unstable fluctuations, with an increasing range of different values for $a$.
+%We can keep more or less the same description showing instead \ref{fig:TSsocialLearning}
+
+The basin plot (fig.\ref{fig:basinunstable}) suggests that this regime is prone to occur when both variants are initialised with high densities in respect to their critical densities.
 In these conditions the total density is too high to ensure a stable coexistence, and the regulatory effect of negative fitness is minimised by the fast rates of cultural transmission (i.e.
 individual shift strategy before experiencing a population decline resulting from density being over $K$).
-Figure XXX2 also implies that high reliance on social learning leeds to the emergence of a new basin of attraction for these unstably cycles, with small differences in initial conditions leading again to profoundly different dynamics.
 
-%\begin{figure}
-%  \centering
-%      \includegraphics[width=0.5\textwidth]{./figures/figureUnstable1.pdf}
-%  \caption{}
-%    \label{}
-%\end{figure}
+\begin{figure}
+  \centering
+      \includegraphics[width=0.5\textwidth]{./figures/figure4.jpg}
+  \caption{}
+    \label{fig:BifurcationSocialLearning}
+\end{figure}
 
-%\begin{figure}
-%  \centering
-%      \includegraphics[width=0.5\textwidth]{./figures/figureUnstable2.pdf}
-%  \caption{}
-%    \label{}
-%\end{figure}
+\begin{figure}
+  \centering
+      \includegraphics[width=0.5\textwidth]{./figures/figure4b.pdf}
+  \caption{}
+    \label{fig:TSsocialLearning}
+\end{figure}
+
+\begin{figure}
+  \centering
+      \includegraphics[width=0.5\textwidth]{./figures/figureUnstable1.pdf}
+  \caption{}
+    \label{fig:basinunstable}
+\end{figure}
 
@@
 fig.
-XXX?).
+\ref{fig:BifurcationSocialLearning}).
+%or any of the three figure above.
 However, this will also depend on the key parameters defining the range of positive payoff for the two strategies (i.e.
@@
 
-Equation \eqref{eqOverlap} ensures that the difference between carrying capacity and critical density is hold constant and equal between the two variants, so that $\lambda$ can be interpreted as an inverse  measure of overlap.
-Indeed when $\lambda=1$ the two variants are identical, and when $\lambda=0$, the critical density of one strategy becomes equal to the carrying capacity of the other (i.e.
+Equation \eqref{eqOverlap} ensures that the difference between carrying capacity and critical density is hold constant and equal between the two variants, so that $\lambda$ can be interpreted as an inverse  measure of overlap: when $\lambda=1$ the two variants are identical, and when $\lambda=0$, the critical density of one strategy becomes equal to the carrying capacity of the other (i.e.
 $A_b=K_a$).
-Figure \ref{fig:overlap} examines three different settings of $\lambda$  with low ($z=0.05$) and intermediate ($z=0.2$) degree of reliance on social learning.
+Figure \ref{fig:overlap} examines three different settings of $\lambda$  with low ($z=0.05$) and intermediate ($z=0.2$) degrees of reliance on social learning.
 The plot suggests that, higher values of $z$ limits the number of initial conditions within the positive payoff zone (i.e.
@@
 The results (fig.
-\ref{fig:percentages}) confirms that indeed with high overlap and high rates of social learning conditions enabling the stable coexistence is reduced, initially as an increase of conditions promoting the dominance of one variant or another, and eventually as the unstable fluctuations regime take over.
-With decreasing $\lambda$ unstable regimes vanish, and the relationship between $z$ and the size of the basin of attraction for stable coexistence becomes more or less constant for each value of $\lambda$.
+\ref{fig:percentages}) confirms that indeed with high overlap and high rates of social learning conditions enabling the stable coexistence is reduced, initially as a result of an increase of conditions promoting the dominance of one variant or another, and eventually because the unstable fluctuations regime takes over.
+When  $\lambda$ is small, however, unstable regimes vanish, and the relationship between $z$ and the size of the basin of attraction for stable coexistence becomes more or less constant for each value of $\lambda$.
 
@@
 
-\begin{figure*}
+\begin{figure}
   \centering
@@
     \label{fig:percentages}
-\end{figure*}
+\end{figure}
 
@@
 
-%This section needs to be revised:
-Figure \ref{fig:percentages} also suggest that with decreasing values of $\lambda$, the variant with lower carrying capacity/critical density expands its basin of attraction.
-This implies that, with other things being equal, traits with higher $A$ and $K$ have a smaller number of favourable conditions that can promote its existence.
-Indeed in the case of the two-variants model examined here, the strategy with higher $A$ and $K$ can survive primarily when its initial density is comparatively high (e.g.
-above or equal its maximum payoff) and when the alternative strategy is close to its carrying capacity.
-The latter condition implies that novel strategies with higher carrying capacity but also higher critical densities are fully adopted only when the extant strategy with lower $A$ and $K$ has a density closer to its upper threshold.
-Thus innovations capable to sustain larger population density requires either a comparatively lower critical density that ensures a higher overlap with extant strategies, or an initial number of adoptees that are not guided by a payoff-based transmission.
-A third option, which we briefly explore here, is to relax the assumption that the payoff of each variant is exclusively a function of their intrinsic density.
+The results detailed in previous section indicates that, within the positive payoff region, fixation to one strategy or another can occur only for a comparatively narrow set of initial conditions close to the boundaries between different basins of attractions.
+For all other conditions we expect to observe either a stable coexistence or unstable fluctuations depending on how much individuals are prone to social learning.
+The results thus implies that if two variants have the similar payoff values, there is no intrinsic advantage of one strategy over the other.
+Thus the variant with higher critical thresholds $A$ and $K$ do not necessarily predominate over a wider range of initial population sizes, and instead its full adoption by the entire population $N$ is ensure only when its initial density is comparatively high (e.g.
+above or equal its maximum payoff) and when the alternative strategy is close to its critical threshold $A$.
+This also implies that if the variant with higher $A$ and $K$ is novel, some mechanism (e.g.
+an alternative form of social learning) must ensure that its density approaches its highest payoff, and at the same, the incumbent strategy should exhibit a comparatively low density.
+Alternatively, the novel variant should either exhibit a higher payoff (thus has some intrinsic advantage) or some form of competition between the two variant should be considered.
+Here we briefly explore the latter scenario, by relaxing the assumption of independence in the computation of the payoff of each variant.
 For instance several subsistence strategies do not have fully independent resource pools, and hence the success of a given variant might be detrimental to the payoff of the other irregardless of the effects provoked by cultural transmission.
@@
 
-where the four new parameters $c_{A_a}$, $c_{A_b}$, $c_{K_a}$ and $c_{K_b}$, measure respectively the negative (or positive) effect determined by the number of individuals of the “other” traits.
+where the four new parameters $c_{A_a}$, $c_{A_b}$, $c_{K_a}$ and $c_{K_b}$, measure respectively the negative (or positive) effect determined by the number of individuals of the €œother€ variant. 
 When $c_{A}>0$ the Allee threshold is essentially increased, while when $c_{K}>0$ the carrying capacity is diminished as function of the other population. 
 Larger values of any of the $c$ parameters dictates a stronger influence of the other population.
+While the unstable node at the critical density ($A$) is maintained, equation \ref{eqCompetition} does not have an equilibrium point at carrying capacity.
+Instead a new, competition-driven equilibrium can be identified by the following pair of equations:
 
-%We need to mention that analytical equilibrium solution and put the equations here. Then discuss about the effects for different values of z.
+\begin{equation}
+\begin{aligned}
+a=\frac{K_a-c_{K_b}K_b}{1-c_{K_b}c_{K_a}}\\
+b=\frac{c_{K_a}K_a-K_b}{c_{K_a}c_{K_b}-1}
+\label{eqCompetitionSolution}
+\end{aligned}
+\end{equation}
 
-%what if competition
-%show new stable points
-%show dynamics?
+Notice that when all competition parameters are set to 0, equation \ref{eqCompetitionSolution} points to $K_a$ and $K_b$ as the stable equilibrium node.
 
 Figure \ref{fig:competition} shows a scenario where strategy $a$ is dominant over strategy $b$.
+Here the combination of high reliance on social learning and competitive advantage can have a drastic effect to the system, with a wide combination of parameters leading to a fixation to $b$, and a stable coexistence ensured only for when $a$ has densities close to its carrying capacity.
 
@@
 
-% XAVI some brief discussion on results here, and update the dashed lines
-
-% Do we need more? Perhaps another iteration of the overlap analysis with the competition promoting the variant with higher critical density and carrying capacity?
+% TODO : Do we need more? Perhaps another iteration of the overlap analysis with the competition promoting the variant with higher critical density and carrying capacity?
 
@@
 
-% Cooperation / Facilitation / and Niche Construction review ?
-
 Our analysis examined the impact of payoff-based social learning strategies in the evolutionary dynamics of cultural traits with direct and inverse density dependence.
 The results showed that social learning can alter the basins of attraction of stable nodes, leading to different dynamics and equilibria from the very same initial conditions.
-When the reliance on social learning is comparatively high, and  the two variants have a substantial overlap in their range of positive payoff (i.e.
-they share similar critical densities and carrying capacities), we also observe the emergence of a chaotic behaviour where the system fluctuates without reaching a stable equilibria.
-Our results imply that evolutionary trajectories that are generally expected from a standard two population Allee models can exhibit unexpected outcomes.
-Variants below critical density can survive and a stable coexistence of strategies above this threshold is not necessarily guaranteed.
-The effects of social learning can also be observed in the trajectories of these evolutionary dynamics .
-Reversions to lower densities do not exclusively occur only when the variant is above its carrying capacity, and when the reliance on social learning is high these episodes becomes more frequent and extreme in its magnitude.
+If there is no reliance on social learning, the critical density separates the ultimate equilibrium of the system: if the initial condition is below this threshold the variant is destined to go extinct, whilst if it is above this value it will eventually reach carrying capacity.
+When social learning is present this boundary shifts, variants can survive even if its initial condition is below critical density and equally can reach extinction even if above this threshold.
+This scenario is further enhanced when the two variants are not independent but are in direct competition.
+In this case, social learning can further enhance the advantage of one variant over the other.
 
-%Not sure where to put competition, maybe the whole thing should be in the appendix?
+When the reliance on social learning is comparatively high, and  he two variants have a substantial overlap in their range of positive payoff densities (i.e.
+they share similar critical densities and carrying capacities), the system does not always reach an stable equilibrium, but can instead exhibit cyclical fluctuations between the two variants.
+In other words the consequences of social learning is not absorbed by the system, and instead a negative feedback process leads to continuous switch from one variant to the other.
 
-The implications of our model can provide some insights in the evolutionary history of subsistence traits in human history.
-Many prime examples of keystone innovations are closely related to the introduction of new subsistence strategies that enabled a larger carrying capacity, often coupled with the cost of higher critical densities.
+% This is a bullet list of the main conclusion as an alternative to the paragraph above.
+
+%\begin{itemize}
+%\item{In the absence of social learning and competition, both variants survive as long as they are above their respective critical densities (fig.\ref{fig:NoTransmissionBasin}),}
+%\item{If social learning is present, variants can be survive even when it is below its critical density. This scenario can emerge if: 
+%1) the alternative variant is also below the critical density but its gap from $A$ is larger;
+%2) the alternative strategy is beyond its carrying capacity (fig. \ref{fig:TransmissionBasin})}
+%\item{However, social learning result also into a non-zero probability of an extinction even when a variant is above its critical density.
+This occurs when the variant is close to its critical density and is maximised when the alternative variant has intermediate densities (i.e.
+highest payoff) (fig.
+\ref{fig:TransmissionBasin})}
+%\item{Increasing reliance on social learning and high similarity in the critical density and carrying capacity of the two variants leads to an increase in the length of boundary between the basin of attractions for the extinction of either variants at low densities.
+This implies that, under these settings, a small change in behaviour or an external fluctuations can promote the extinction of variant or the other but reduces the chances of a coexistence.
+(fig.
+\ref{fig:overlap})}
+%\item{Increasing reliance on social learning and high similarity in the critical density and carrying capacity also leads to a decline in the stable coexistence of the two strategies and the emergence of cyclical fluctuations between the two variants.}
+%\item{Competition between the two variants can lead to a wider range of initial conditions where the subdominant variant is likely to go extinct, especially when its density is not sufficiently high}
+%\end{itemize}
+
+The implications of our model can provide some insights in the evolutionary history of cultural traits exhibiting density dependence.
+Perhaps the most notable example is the spread of keystone innovations in subsistence strategies which led to the adoption of variants enabling larger carrying capacity, often at the cost of higher critical densities.
 The spread of behavioural traits such as the know-how of the domestication process~\citep{barker2006} the use of new subsistence tools~\citep{petraglia_population_2009} and processing techniques~\citep{molleson1993}, as well as the inclusion/exclusion and extensification/intensification of different prey species are just some of the most notable examples.
-Both ethnographic and archaeological evidence show many instances where these critical transitions where not smooth, often leading to the coexistence of alternative strategies as well as episodes of a return to a previously abandoned strategies.
-%PUT HERE LIST of EXAMPLES.
-While in most cases these are more anecdotal evidence than testable empirical cases, the insights of our models suggests that different levels of reliance on social learning could in fact profoundly drive the system into different trajectories.
-These insights can possibly contribute to existing models of subsistence shifts proposed in human behavioural ecology\citep{smith1992,bird2006,kennett2006}, combining its core theoretical framework to the one developed by cultural evolution studies.
+In many of these examples the initial spread of the innovation generated a positive feedback process, either through enhanced possibility of direct cooperation, or some form of mutual facilitation often indirectly emerging from positive niche construction.
+At the same time, ethnographic and archaeological evidence show how many instances these critical transitions where not smooth, often leading to the coexistence of alternative strategies as well as episodes of a return to a previously abandoned cultural traits.
+Several ethnographic examples showcase small-scale societies between alternating between different subsistence modes over a comparatively brief period of time \citep{layton1991,mace_1993,oota_et_al_2005},
+ while increasing archaeological evidence suggest that the transition from hunting-gathering to farming was not unidirectional, but with instances of temporary reversions  \citep{rowley2001,stevens_fuller_2012} %ideally we need a couple of more citations here?
+While in most cases these are  anecdotal evidence rather than testable empirical cases, the insights of our models suggests that different levels of reliance on social learning could alone portray some of the evolutionary trajectories observed in these systems.
 

### 2015-06-26 manuscript/draft.tex
errors fixed

@@
 
-where the four new parameters $c_{A_a}$, $c_{A_b}$, $c_{K_a}$ and $c_{K_b}$, measure respectively the negative (or positive) effect determined by the number of individuals of the €œother€ variant. 
+where the four new parameters $c_{A_a}$, $c_{A_b}$, $c_{K_a}$ and $c_{K_b}$, measure respectively the negative (or positive) effect determined by the number of individuals of the other variant. 
 When $c_{A}>0$ the Allee threshold is essentially increased, while when $c_{K}>0$ the carrying capacity is diminished as function of the other population. 

### 2015-07-07 manuscript/draft.tex
small fixes

@@
 We explore the evolutionary dynamics of a single trait-two variants model, each exhibiting a different Allee effect.
-We used a modified Lotka-Volterra model of payoff-biased social learning, and show that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion, stable co-existence, and fixation.
+We use a modified Lotka-Volterra model of payoff-biased social learning, and show that different degrees of reliance on social learning can exhibit a variety of equilibria, including episodes of reversion, stable co-existence, and fixation.
 
@@
 For example, behaviour linked to the exploitation of limited resources will exhibit a decline in payoff when density exceeds a critical threshold (i.e carrying capacity).
-In this case a new optimal variants are likely to spread rapidly when the they are rare, but as the number of individuals possessing these approach their respective carrying capacities, their beneficial effects will start to decline, the payoff  decrease, and eventually rarer suboptimal variants might be preferred (see Lake and Crema \citeyear{lake_and_crema_2012} for an extensive analysis).
+In this case new optimal variants are likely to spread rapidly when they are rare, but as the number of individuals possessing them approach their respective carrying capacities, their beneficial effects will start to decline, the payoff will decrease, and eventually rarer suboptimal variants might be preferred (see Lake and Crema \citeyear{lake_and_crema_2012} for an extensive analysis).
 
@@
 For the rest of the paper we will consider only scenarios where the node $x=K_x$ is stable.
-In this case the initial value of $x(t=0)$ will determine the ultimate equilibria of the system: extinction, if $x(t=0)<A_x$; unstable equilibrium, if $x(t=0)=A_x$; and stable equilibrium, if $x(t=0)>A_x$.
+In this case the initial value of $x_{t=0}$ will determine the ultimate equilibria of the system: extinction, if $x_{t=0}<A_x$; unstable equilibrium, if $x_{t=0}=A_x$; and stable equilibrium, if $x_{t=0}>A_x$.
 Thus in a 2-trait model, the coexistence of both variants is guaranteed only when the initial density of both traits are above their respective critical densities, otherwise we should expect that either (i.e.
-$a(t=0) \geq A_a$ and $b(t=0)<A_b$ or $a(t=0)<A_a$ and $b(t=0) \geq A_b$) or both (i.e.
-when $a(t=0)<A_a$ and $b(t=0)<A_b$) variants are ultimately extinct (see fig.
+$a_{t=0} \geq A_a$ and $b_{t=0}<A_b$ or $a_{t=0}<A_a$ and $b_{t=0} \geq A_b$) or both (i.e.
+when $a_{t=0}<A_a$ and $b_{t=0}<A_b$) variants are ultimately extinct (see fig.
 \ref{fig:NoTransmissionBasin}).
@@
 Figure \ref{fig:TransmissionBasin} shows an example of this.
-When social learning is not present (fig.~\ref{fig:TransmissionBasin}a), the initial conditions $i$, $j$, and $k$ result into a stable equilibrium at $a=K_a$ and $b=K_b$, whilst the starting point $l$ leads to the extinction of $b$.
-When $z=0.05$ (fig.~\ref{fig:TransmissionBasin}b), the basins of attractions are modified, resulting into different trajectories and in some cases different equilibria for the same initial conditions.
+When social learning is not present (fig.~\ref{fig:TransmissionBasin} left), the initial conditions $i$, $j$, and $k$ result into a stable equilibrium at $a=K_a$ and $b=K_b$, whilst the starting point $l$ leads to the extinction of $b$.
+When $z=0.05$ (fig.~\ref{fig:TransmissionBasin} right), the basins of attractions are modified, resulting into different trajectories and in some cases different equilibria for the same initial conditions.
 Thus $i$ and $k$ are no longer leading to the coexistence of both traits but to the demise of one or the other, whilst $l$ now ensures the coexistence of both variants.
@@
 Figure \ref{fig:BifurcationSocialLearning} shows the bifurcation diagram for the population size of $a$ as a function of $z$.
-Here initial the initial effects of $z$ does not alter the equilibrium point ($K_a=0.6$), but once $z>0.5$ we observe a sudden transition to unstable fluctuations, with an increasing range of different values for $a$.
+Here the initial effects of $z$ do not alter the equilibrium point ($K_a=0.6$), but once $z>0.5$ we observe a sudden transition to unstable fluctuations, with an increasing range of different values for $a$.
 %We can keep more or less the same description showing instead \ref{fig:TSsocialLearning}
 
+% TODO XRC I thought we were removing this figure
 The basin plot (fig.\ref{fig:basinunstable}) suggests that this regime is prone to occur when both variants are initialised with high densities in respect to their critical densities.
@@
 
+% TODO XRC choose figure4 or figure4b
 \begin{figure}
@@
 For all other conditions we expect to observe either a stable coexistence or unstable fluctuations depending on how much individuals are prone to social learning.
-The results thus implies that if two variants have the similar payoff values, there is no intrinsic advantage of one strategy over the other.
-Thus the variant with higher critical thresholds $A$ and $K$ do not necessarily predominate over a wider range of initial population sizes, and instead its full adoption by the entire population $N$ is ensure only when its initial density is comparatively high (e.g.
+The results thus implies that if two variants have similar payoff values, there is no intrinsic advantage of one strategy over the other.
+Thus the variant with higher critical thresholds $A$ and $K$ do not necessarily predominate over a wider range of initial population sizes, and instead its full adoption by the entire population $N$ is ensured only when its initial density is comparatively high (e.g.
 above or equal its maximum payoff) and when the alternative strategy is close to its critical threshold $A$.
@@
   \caption{The impact of competition on the basins of attraction for the different equilibria.
-The Parameters used are $z=0.05$ (left) and $z=0.2$ (right) for social learning, and $c=-0.1$ (top) and $c=0.1$ (bottom) for competition.}
+The Parameters used are $z=0.05$ (up) and $z=0.2$ (bottom) for social learning, and $c=0$ (left), $c=0.1$ (center) and $c=0.2$ (right) for competition.}
     \label{fig:competition}
@@
 The results showed that social learning can alter the basins of attraction of stable nodes, leading to different dynamics and equilibria from the very same initial conditions.
-If there is no reliance on social learning, the critical density separates the ultimate equilibrium of the system: if the initial condition is below this threshold the variant is destined to go extinct, whilst if it is above this value it will eventually reach carrying capacity.
+If there is no reliance on social learning, the critical density separates the ultimate equilibrium of the system: if the initial condition is below this threshold the variant is destined to go extinct, whilst if it is above this value will eventually reach carrying capacity.
 When social learning is present this boundary shifts, variants can survive even if its initial condition is below critical density and equally can reach extinction even if above this threshold.
@@
 
-When the reliance on social learning is comparatively high, and  he two variants have a substantial overlap in their range of positive payoff densities (i.e.
+When the reliance on social learning is comparatively high, and the two variants have a substantial overlap in their range of positive payoff densities (i.e.
 they share similar critical densities and carrying capacities), the system does not always reach an stable equilibrium, but can instead exhibit cyclical fluctuations between the two variants.
@@
 In many of these examples the initial spread of the innovation generated a positive feedback process, either through enhanced possibility of direct cooperation, or some form of mutual facilitation often indirectly emerging from positive niche construction.
-At the same time, ethnographic and archaeological evidence show how many instances these critical transitions where not smooth, often leading to the coexistence of alternative strategies as well as episodes of a return to a previously abandoned cultural traits.
-Several ethnographic examples showcase small-scale societies between alternating between different subsistence modes over a comparatively brief period of time \citep{layton1991,mace_1993,oota_et_al_2005},
- while increasing archaeological evidence suggest that the transition from hunting-gathering to farming was not unidirectional, but with instances of temporary reversions  \citep{rowley2001,stevens_fuller_2012} %ideally we need a couple of more citations here?
-While in most cases these are  anecdotal evidence rather than testable empirical cases, the insights of our models suggests that different levels of reliance on social learning could alone portray some of the evolutionary trajectories observed in these systems.
+At the same time, ethnographic and archaeological evidence show how many instances of critical transitions where not smooth, often leading to the coexistence of alternative strategies as well as episodes of a return to a previously abandoned cultural traits.
+Several ethnographic examples showcase small-scale societies alternating between different subsistence modes over a comparatively brief period of time \citep{layton1991,mace_1993,oota_et_al_2005}, while increasing archaeological evidence suggest that the transition from hunting-gathering to farming was not unidirectional, but with instances of temporary reversions \citep{rowley2001,stevens_fuller_2012} %ideally we need a couple of more citations here?
+While in most cases these are anecdotal evidence rather than testable empirical cases, the insights of our model suggests that different levels of reliance on social learning could alone portray some of the evolutionary trajectories observed in these systems.
 

### 2015-07-09 manuscript/draft.tex
acquisition cost
unstability section

@@
 where $z$ is a measure of reliance in social learning (effectively equivalent to highest possible transmission rate), $\Delta$ is the maximum between $R_{max(a,t)}$ and $R_{max(b,t)}$, and $m$ is a calibration parameter that measures the perception of the difference in the payoffs (i.e.
-small values of $m$ determines a higher rate by which the theoretical maximum transmission rate is reached).
+small values of $m$ determines a higher rate by which the theoretical maximum transmission rate is reached), thus portraying potential acquisition costs of the new variant~\citep{henrich2001cultural}.
 
@@
   \caption{Basins of attraction for the four equilibria (mutual extinction: grey; extinction of $a$: blue; extinction of $b$: red; stable coexistence: green) with (right) and without (left) social learning.
-Model parameters: $A_a=0.2$; $K_a=0.8$; $A_b=0.35$; $K_b=0.95$; $r_a=0.05$;$r_b=0.1039063$; $z=0$ (left panel) and $z=0.2$ (right panel) }
+Model parameters: $A_a=0.2$; $K_a=0.8$; $A_b=0.35$; $K_b=0.95$; $r_a=0.05$; $r_b=0.1039063$; $z=0$ (left panel) and $z=0.2$ (right panel) }
     \label{fig:TransmissionBasin}
@@
 \ref{fig:bifurcationDiagram}), that is a decrease in the stability of the fixed points.
-Figure \ref{fig:BifurcationSocialLearning} shows the bifurcation diagram for the population size of $a$ as a function of $z$.
-Here the initial effects of $z$ do not alter the equilibrium point ($K_a=0.6$), but once $z>0.5$ we observe a sudden transition to unstable fluctuations, with an increasing range of different values for $a$.
-%We can keep more or less the same description showing instead \ref{fig:TSsocialLearning}
-
-% TODO XRC I thought we were removing this figure
-The basin plot (fig.\ref{fig:basinunstable}) suggests that this regime is prone to occur when both variants are initialised with high densities in respect to their critical densities.
-In these conditions the total density is too high to ensure a stable coexistence, and the regulatory effect of negative fitness is minimised by the fast rates of cultural transmission (i.e.
+Figure \ref{fig:TSsocialLearning} shows fluctuations in population size of the system over time for different values of $z$.
+Here low values of $z$ do not change the equilibrium point at ($a=K_a$ and $b=K_b$).
+However, at $z \geq 0.6$ unstable dynamics can be observed leading the system towards limit cycles.
+In this configuration the regulatory effect of negative fitness is minimised by the fast rates of cultural transmission (i.e.
 individual shift strategy before experiencing a population decline resulting from density being over $K$).
-
-% TODO XRC choose figure4 or figure4b
-\begin{figure}
-  \centering
-      \includegraphics[width=0.5\textwidth]{./figures/figure4.jpg}
-  \caption{}
-    \label{fig:BifurcationSocialLearning}
-\end{figure}
 
@@
   \centering
-      \includegraphics[width=0.5\textwidth]{./figures/figure4b.pdf}
-  \caption{}
+      \includegraphics[width=1\textwidth]{./figures/figure4b.pdf}
+  \caption{Time series of population size for $z=0$ (left), $z=0.3$ (center) and $z=0.6$ (right).
+Increasing values of $z$ shift the system from the co-existence equilibrium towards a limit cycle with abrupt population changes.
+Model parameters: $A_a=0.1$; $A_b=0.2$; $K_a=0.6$; $K_b=0.7$; $r_a=0.05$}
     \label{fig:TSsocialLearning}
-\end{figure}
-
-\begin{figure}
-  \centering
-      \includegraphics[width=0.5\textwidth]{./figures/figureUnstable1.pdf}
-  \caption{}
-    \label{fig:basinunstable}
 \end{figure}
@@
 fig.
-\ref{fig:BifurcationSocialLearning}).
-%or any of the three figure above.
+\ref{fig:TransmissionBasin}).
 However, this will also depend on the key parameters defining the range of positive payoff for the two strategies (i.e.

### 2015-08-03 manuscript/draft.tex
missed note
minor edits
I still need to edit the discussion section

@@
 \begin{abstract}
-Several social learning strategies guide the selection of cultural variant to imitate through an evaluation of the expected benefits associated with each observed alternatives.
-The assumption of the social learner is that the payoff received by the transmitter while expressing a given cultural variant is a proxy for  inferring the qualitative features of the variant itself.
+Several social learning strategies guide the selection of cultural variant through an evaluation of the expected benefits associated with each observed alternatives.
+The assumption of the social learner is that the payoff received by the transmitter, while displaying a given variant, is a proxy for  inferring the qualitative features of the variant itself.
 Here we explore the theoretical implications when: 1) the payoff attributed to a given cultural variant directly affects the reproductive success of its bearer;  and 2) payoff is density dependent, that is its expected amount is a function of the number of individuals possessing the same variant.
-We consider situations where such density dependence is simultaneously direct and inverse, with minimum and maximum critical thresholds bounding the range of positive payoff, with a peak value at midpoint.
-This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits that are enhanced by cooperation or mutual facilitation, but are somehow constrained by some form of carrying capacity.
+We consider situations where such density dependence is both direct and inverse, with minimum and maximum critical thresholds bounding the range of positive payoff and a peak value at midpoint.
+This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits that are enhanced by cooperation or mutual facilitation, but are  constrained by some form of carrying capacity.
 We explore the evolutionary dynamics of a single trait-two variants model, each exhibiting a different Allee effect.
@@
 The last three decades have witnessed a steady growth of formal models of cultural evolution.
-Early works inspired by population biology \citep{cavallisforza_feldman_1981,boyd1985} have laid the foundation of a now mature trans-disciplinary field which unites the social and biological sciences \citep{mesoudi_etal_2006}.
+Early works inspired by population biology \citep{cavallisforza_feldman_1981,boyd1985} have laid the foundation of a now mature trans-disciplinary field which unites the social and biological sciences \citep{mesoudi_etal_2006, mcelreath_and_boyd_2007}.
 The rich cross-fertilisation between different subfields has promoted the designing of new analytical techniques as well as new uses for existing methods, but more crucially have prompted the development of theoretical models of how culture is acquired and transmitted between individuals~\citep{mesoudi_cultural_2015}.
+%We need probably a list of references, say about 5-7 papers?
 
-A central element, shared by most models of social learning, is the existence of some heuristics that promotes the selection of a given cultural variant over another \citep{laland2004}.
+At the core of most models of social learning there is some heuristics that promotes the selection of a given cultural variant over another \citep{laland2004}.
 This can be based on some properties of the variant itself (\emph{content bias}) or derived from contextual cues (\emph{context bias}), such as the commonness or the rareness of a trait (\emph{conformist bias} and \emph{anti-conformist bias}), or  features associated with the transmitter such as prestige, success, or similarity to the learner (\emph{context bias})\citep{henrich_mcelreath2003}.
@@
 This paper examines a group of social learning strategies generally referred to as \emph{payoff-based transmission} \citep{schlag1998,kendal_etal_2009,lake_and_crema_2012,baldini2013,kandler_and_laland_2013,crema_lake_inpress}.
-Here the probability of copying a cultural variant $i$ is directly proportional to a payoff $s$ observed amongst individuals possessing $i$.
+Here the probability of copying a cultural variant $i$ is directly proportional to a payoff $s$ exhibited by individuals possessing $i$.
 In other words the social learner assumes that the success of an individual (measured in $s$) is a function of the observed behavioural trait, that is $s=f(i)$.
@@
 
-Whilst most studies focused on how difference in these imitation rules can drive the evolutionary dynamics of the cultural system, less attention has been given to the relationship between target cultural traits and their payoff signals.
+Whilst most studies focus on how difference in these imitation rules can drive the evolutionary dynamics of the cultural system, less attention has been given to the relationship between target  traits and their payoff signals.
 Typical signals are most likely some indirect proxy of success (e.g.
-number of offspring, income, presence/absence of other traits associated with success or prestige, etc.)  that is assumed (by the social learner) to be, to some extent,  correlated with intrinsic properties of the cultural variant.
-In most cases the causal link between the payoff and a cultural trait is just inferred, and the exact processes and variables linking the two is unknown.
-This implies that the strength of correlation between payoff and cultural trait can vary, introducing different degrees of uncertainty that can have a strong impact to the efficiency of specific learning rules.
-For example,~\citet{crema_lake_inpress} have shown that \emph{imitate the best} is increasingly detrimental with high payoff uncertainty and a large pool of potential social teachers.
-This is because innovations are initially rare in a population, and hence even if their average payoff are higher than extant suboptimal alternatives, their probability of signalling the highest payoff, and hence being selected, is low.
+number of offspring, income, presence/absence of other traits associated with success or prestige, etc.)  that is assumed (by the social learner) to be correlated with intrinsic properties of the cultural variant.
+In most cases the causal link between  payoff and cultural trait is just inferred, and the exact processes and variables linking the two is unknown.
+Furthermore the strength of correlation between payoff and cultural trait can vary, introducing different degrees of uncertainty that can hinder the correct evaluation of a trait depending on the form of social learning.
+For example,~\citet{crema_lake_inpress} have shown that \emph{imitate the best} can even be detrimental in some situations where both payoff uncertainty  and the pool of potential social teachers are high.
+This is because beneficial innovations are initially rare in a population, and hence even if their average payoff are higher than extant suboptimal alternatives, their probability of displaying the highest payoff, and hence being selected, is low.
 
@@
 In other words, cultural and ecological inheritance should in part reduce the impact played by $\epsilon$, and consequently the uncertainty in the payoff-signal.
-There are however exceptions.
-One of these are contexts where the success of a variant is strongly correlated with the absolute number of individuals $n_i$ possessing the same trait ($i$) within a population.
+There are however situations where even with most variables hold constant the payoff is subject to high variability.
+Perhaps one of the most relevant examples are
+are contexts where the success of a variant is strongly correlated with the absolute number of individuals $n_i$ possessing the same trait ($i$) within a population.
 This density dependence, which can be described by the payoff function $f(i,n_i)$, implies that the adoption (or abandonment) of a trait can alter its expected payoff-signal, potentially causing a cascade of feedback effects.
@@
 
-Although a systematic review and a shared theoretical framework akin to those proposed in population ecology \citep{kramer_etal_2009} is not present in the social sciences, there are several lines of evidence supporting the existence of both direct and inverse density dependence in cultural and behavioural traits.
-For example several ethnographic studies on small scale societies show how the per-capita return of hunting groups show higher yield when the group size is not too small nor too big~\citep{hill_and_hawkes_1983,janssen_and_hill_2014}.
+While the concept of Allee effect is well established in population ecology (see Kramer et al \citeyear{kramer_etal_2009} for a systematic review), there is no concept with direct equivalence in the social sciences.
+That said, there are multiple, established lines of evidence that support  the existence of both direct and inverse density dependence in cultural and behavioural traits.
+For example several ethnographic studies on small scale societies show how the per-capita return of hunting groups show higher yield when the group size is not too small nor too big~ \citep{hill_and_hawkes_1983,janssen_and_hill_2014}.
 Other instances of inverse density dependence include population level activities such as the use of anthropogenic fire~\citep{bird2013} or selective hunting~\citep{dods_2002} which often require a population size above a certain critical density in order to promote sufficient niche enhancement and a consequent increase in payoff~\citep{rowley-conwy_and_layton_2011}.
 In a broader sense any cultural trait that benefits from direct or indirect cooperation,
-%Add reference \citep{}
 promote positive niche construction \citep{vandermeer_2008}, and requires a critical mass \citep{rogers_2003} are likely to exhibit some degree of Allee effect.
-The implication is thus not limited necessarily to the domain of traits directly promoting reproductive fitness, and hence can include language~\citep{kandler2009}, organisational populations~\citep{caroll_and_hannan_1989} and communication technologies \citep{van_slyke_perceived_2007}.
+The implication is thus not limited to traits that directly enhance reproductive fitness, and can include language~\citep{kandler2009}, organisational populations~\citep{caroll_and_hannan_1989} and communication technologies \citep{van_slyke_perceived_2007}.
 
-Here we explore the consequences of Allee effect in the evolutionary dynamics of cultural transmission, assuming that: 1) the social learning strategy is payoff-based; and 2)  the payoff is equivalent to the reproductive fitness of each individual.
-We will demonstrate, by means of a difference equation, that direct and inverse density dependence can strongly affect the evolutionary trajectory of the system, leading to changes in the conditions required to reach specific equilibria, as well as to the emergence of unstable patterns characterised by temporary or permanent episodes of shifts between alternative cultural variants.
+Here we explore a difference equation model of cultural transmission with traits exhibiting an Allee effect.
+We assume that: 1) the social learning strategy is payoff-based; and 2)  the payoff is equivalent to the reproductive fitness of each individual.
+We demonstrate that direct and inverse density dependence can strongly affect the evolutionary trajectory of the system, leading to changes in the conditions required to reach specific equilibria, as well as to the emergence of unstable patterns characterised by temporary or permanent episodes of shifts between alternative cultural variants.
 
@@
 where $r$ is the basic growth rate, $A$ is the critical density and $K$ is the carrying capacity, following the condition $0<A < K$.
+Notice that here $R$ is also the payoff of each variant for a given population size.
 
@@
 The cultural transmission component of equation \eqref{eq1} is based on the \emph{proportional imitation rule} \citep{schlag1998}.
-In this case, the probability of an individual with lower growth rate to adopt the trait of an individual with higher growth rate is proportional to the difference in the growth rates.
+In this case, the probability of an individual with lower payoff to adopt the alternative, optimal variant is proportional to the difference in the growth rates.
 This is formally portrayed as follows:
@@
 where $z$ is a measure of reliance in social learning (effectively equivalent to highest possible transmission rate), $\Delta$ is the maximum between $R_{max(a,t)}$ and $R_{max(b,t)}$, and $m$ is a calibration parameter that measures the perception of the difference in the payoffs (i.e.
-small values of $m$ determines a higher rate by which the theoretical maximum transmission rate is reached), thus portraying potential acquisition costs of the new variant~\citep{henrich2001cultural}.
+small values of $m$ determines a higher rate by which the theoretical maximum transmission rate is reached), thus portraying also potential acquisition costs of the new variant (c.f.
+"threshold of evidence', Hernich, \citeyear{henrich2001cultural}: 994)
 
@@
 To achieve this we set $r_b=\rho r_a$, where $\rho$ is a constant\footnote{Given two variants with the same baseline growth rate (i.e.
-$r_a=r_b$), equation \eqref{eq2} result into a lower $R$ for the variant with higher $A$ and $K$.
+$r_a=r_b$), equation \eqref{eq2} returns a lower $R$ for the variant with higher $A$ and $K$.
 In order to avoid this we increased the baseline of the strategy with lower $R$ by a factor $\rho$, given by the following equation:
@@
   } 
-which ensures that the two variants have always the same $R_{max}$.
+that ensures that the two variants have always the same $R_{max}$.
 Given that equations \ref{eq1}-\ref{eq4} have no analytical solution that we are aware of, we identified equilibrium conditions via simulation (source codes of the simulation can be found on the following repository \textcolor{red}{PUT LINK HERE})
@@
 
-For the rest of the paper we will consider only scenarios where the node $x=K_x$ is stable.
+For the rest of the paper we will consider only scenarios where the node $x=K_x$ is stable when $z=0$.
 In this case the initial value of $x_{t=0}$ will determine the ultimate equilibria of the system: extinction, if $x_{t=0}<A_x$; unstable equilibrium, if $x_{t=0}=A_x$; and stable equilibrium, if $x_{t=0}>A_x$.
@@
 First, the basins of attraction are transformed, and hence the system equilibrium at $z=0$ and $z>0$ can be different for the same initial conditions.
-This is particularly the case for combinations of $a$ and $b$ that are closer to the boundaries between two or more basins.
+This is particularly the case for combinations of $a$ and $b$ that are closer to the boundaries between two or more basins when $z=0$.
 Figure \ref{fig:TransmissionBasin} shows an example of this.
@@
 In this configuration the regulatory effect of negative fitness is minimised by the fast rates of cultural transmission (i.e.
-individual shift strategy before experiencing a population decline resulting from density being over $K$).
+individuals switch to the alternative trait before the decrease in payoff lead to a population decline).
 
@@
 
-The analyses of the previous section suggest that maximum proportion of population that can switch strategies (i.e.
-the $z$ parameter) has major effects on the final equilibria (c.f.
+The analyses of the previous section suggest that the number of individuals engaging into social learning within a given temporal interval can have major impacts in the qualitative state of the final equilibrium (c.f.
 fig.
 \ref{fig:TransmissionBasin}).
-However, this will also depend on the key parameters defining the range of positive payoff for the two strategies (i.e.
-$[A_a,K_a]$ and $[A_b,K_b]$).
-While a comprehensive analysis is not possible, here we illustrate the relationship between these parameters and $z$ by introducing a new parameter $\lambda \in (0,1)$ adding the following constraints to equation \eqref{eq1}:
+This number is dependent on the propensity to engage into cultural transmission ($z$) and the potential difference in the payoff of the two available cultural variants.
+The latter can be determined by difference in the baseline reproductive rate ($r$), but also by the degree of overlap in the range of positive payoff for the two variants ($[A_a,K_s]$ and $[A_b,K_b]$) alone.
+We explore this second scenario and its relationship to $z$ using the following constraints for our parameters:
 
@@
 
-Equation \eqref{eqOverlap} ensures that the difference between carrying capacity and critical density is hold constant and equal between the two variants, so that $\lambda$ can be interpreted as an inverse  measure of overlap: when $\lambda=1$ the two variants are identical, and when $\lambda=0$, the critical density of one strategy becomes equal to the carrying capacity of the other (i.e.
+Equation \eqref{eqOverlap} ensures that the difference between carrying capacity and critical density is hold constant and equal between the two variants, so that the parameter $\lambda$ can be interpreted as an inverse  measure of overlap: when $\lambda=1$ the two variants are identical, and when $\lambda=0$, the critical density of one strategy becomes equal to the carrying capacity of the other (i.e.
 $A_b=K_a$).
 Figure \ref{fig:overlap} examines three different settings of $\lambda$  with low ($z=0.05$) and intermediate ($z=0.2$) degrees of reliance on social learning.
-The plot suggests that, higher values of $z$ limits the number of initial conditions within the positive payoff zone (i.e.
-when the initial conditions for $a$ and $b$ are within the intervals $[A_a,K_a]$ and $[A_b,K_b]$) where a stable co-existence is achieved, but the magnitude of this effect is smaller for lower values of $\lambda$ (i.e.
+The plot suggests that, higher values of $z$ limits the number of initial conditions where a stable co-existence is achieved within the positive payoff zone (i.e.
+when the initial conditions for $a$ and $b$ are within the intervals $[A_a,K_a]$ and $[A_b,K_b]$), although the magnitude of this effect is smaller for lower values of $\lambda$ (i.e.
 smaller overlap).
-We can further investigate this point by computing the relative areas of different basins of attraction within the positive payoff ozone for different parameter combinations of $z$ and $\lambda$.
+We can further investigate this point by computing the relative areas of different basins of attraction within the positive payoff zone for different parameter combinations of $z$ and $\lambda$.
 The results (fig.
-\ref{fig:percentages}) confirms that indeed with high overlap and high rates of social learning conditions enabling the stable coexistence is reduced, initially as a result of an increase of conditions promoting the dominance of one variant or another, and eventually because the unstable fluctuations regime takes over.
-When  $\lambda$ is small, however, unstable regimes vanish, and the relationship between $z$ and the size of the basin of attraction for stable coexistence becomes more or less constant for each value of $\lambda$.
+\ref{fig:percentages}) confirms that indeed with high overlap and high rates of social learning conditions enabling the stable coexistence is reduced, initially as a result of an increase of conditions promoting the dominance of one variant, and eventually because the unstable fluctuations regime takes over.
+When  $\lambda$ is small, this pattern disappears, and the relationship between $z$ and the size of the basin of attraction for stable coexistence becomes more or less stable for each value of $\lambda$.
 
@@
 The results detailed in previous section indicates that, within the positive payoff region, fixation to one strategy or another can occur only for a comparatively narrow set of initial conditions close to the boundaries between different basins of attractions.
-For all other conditions we expect to observe either a stable coexistence or unstable fluctuations depending on how much individuals are prone to social learning.
+For all other conditions we expect to observe either a stable coexistence or unstable fluctuations depending on the propensity on social learning and the ease by which a switch from one variant to another is possible.
 The results thus implies that if two variants have similar payoff values, there is no intrinsic advantage of one strategy over the other.
@@
 \begin{aligned}
-R_{a,t}& = r_a (\frac{a_t-c_{A_a}b_t}{A_a}-1)(1-\frac{a_t-c_{K_a}b_t}{K_a})\\
-R_{b,t}& = r_b (\frac{b_t-c_{A_b}a_t}{A_b}-1)(1-\frac{b_t-c_{K_b}a_t}{K_b}) 
+R_{a,t}& = r_a \left(\frac{a_t-c_{A_a}b_t}{A_a}-1\right)\left(1-\frac{a_t-c_{K_a}b_t}{K_a}\right)\\
+R_{b,t}& = r_b  \left(\frac{b_t-c_{A_b}a_t}{A_b}-1\right)\left(1-\frac{b_t-c_{K_b}a_t}{K_b}\right) 
 \label{eqCompetition}
@@
 Larger values of any of the $c$ parameters dictates a stronger influence of the other population.
-While the unstable node at the critical density ($A$) is maintained, equation \ref{eqCompetition} does not have an equilibrium point at carrying capacity.
+While the unstable node at the critical density ($A$) is maintained, equation \eqref{eqCompetition} does not have an equilibrium point at carrying capacity.
 Instead a new, competition-driven equilibrium can be identified by the following pair of equations:
@@
 
-Notice that when all competition parameters are set to 0, equation \ref{eqCompetitionSolution} points to $K_a$ and $K_b$ as the stable equilibrium node.
+Notice that when all competition parameters are set to 0, equation \eqref{eqCompetitionSolution} points to $K_a$ and $K_b$ as the stable equilibrium node.
 
@@
 
-% TODO : Do we need more? Perhaps another iteration of the overlap analysis with the competition promoting the variant with higher critical density and carrying capacity?
-
 \section{Discussion}
 
-Our analysis examined the impact of payoff-based social learning strategies in the evolutionary dynamics of cultural traits with direct and inverse density dependence.
-The results showed that social learning can alter the basins of attraction of stable nodes, leading to different dynamics and equilibria from the very same initial conditions.
+Our analysis examined the impact of direct and inverse density dependence in payoff-based social learning strategies.
+The results showed that the coexistence of both positive and negative feedback mechanism can generate a wide range of patterns where small changes in behaviour and initial conditions can both deeply affect the evolutionary trajectories of the system.
+
+social learning can alter the basins of attraction of stable nodes, leading to different dynamics and equilibria from the very same initial conditions.
 If there is no reliance on social learning, the critical density separates the ultimate equilibrium of the system: if the initial condition is below this threshold the variant is destined to go extinct, whilst if it is above this value will eventually reach carrying capacity.
@@
 In other words the consequences of social learning is not absorbed by the system, and instead a negative feedback process leads to continuous switch from one variant to the other.
-
-% This is a bullet list of the main conclusion as an alternative to the paragraph above.
-
-%\begin{itemize}
-%\item{In the absence of social learning and competition, both variants survive as long as they are above their respective critical densities (fig.\ref{fig:NoTransmissionBasin}),}
-%\item{If social learning is present, variants can be survive even when it is below its critical density. This scenario can emerge if: 
-%1) the alternative variant is also below the critical density but its gap from $A$ is larger;
-%2) the alternative strategy is beyond its carrying capacity (fig. \ref{fig:TransmissionBasin})}
-%\item{However, social learning result also into a non-zero probability of an extinction even when a variant is above its critical density.
-This occurs when the variant is close to its critical density and is maximised when the alternative variant has intermediate densities (i.e.
-highest payoff) (fig.
-\ref{fig:TransmissionBasin})}
-%\item{Increasing reliance on social learning and high similarity in the critical density and carrying capacity of the two variants leads to an increase in the length of boundary between the basin of attractions for the extinction of either variants at low densities.
-This implies that, under these settings, a small change in behaviour or an external fluctuations can promote the extinction of variant or the other but reduces the chances of a coexistence.
-(fig.
-\ref{fig:overlap})}
-%\item{Increasing reliance on social learning and high similarity in the critical density and carrying capacity also leads to a decline in the stable coexistence of the two strategies and the emergence of cyclical fluctuations between the two variants.}
-%\item{Competition between the two variants can lead to a wider range of initial conditions where the subdominant variant is likely to go extinct, especially when its density is not sufficiently high}
-%\end{itemize}
 
@@
 
-\section{Appendix I}
-
--50,000 iterations
--floating point arithmetic
-
 \end{document}

### 2015-09-15 manuscript/draft.tex
(Semi)-Final Draft

@@
 We consider situations where such density dependence is both direct and inverse, with minimum and maximum critical thresholds bounding the range of positive payoff and a peak value at midpoint.
-This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits that are enhanced by cooperation or mutual facilitation, but are  constrained by some form of carrying capacity.
+This non-linear relationship, known as Allee effect, portrays a large number of behavioural traits that are enhanced by cooperation or mutual facilitation, but are constrained by finite amount of resources.
 We explore the evolutionary dynamics of a single trait-two variants model, each exhibiting a different Allee effect.
@@
 Early works inspired by population biology \citep{cavallisforza_feldman_1981,boyd1985} have laid the foundation of a now mature trans-disciplinary field which unites the social and biological sciences \citep{mesoudi_etal_2006, mcelreath_and_boyd_2007}.
-The rich cross-fertilisation between different subfields has promoted the designing of new analytical techniques as well as new uses for existing methods, but more crucially have prompted the development of theoretical models of how culture is acquired and transmitted between individuals~\citep{mesoudi_cultural_2015}.
-%We need probably a list of references, say about 5-7 papers?
+The rich cross-fertilisation between different subfields has promoted the designing of new analytical techniques as well as new uses for existing methods, but more crucially have prompted the development of theoretical models of how culture is acquired and transmitted between individuals~\citep{henrich_mcelreath2003,laland2004,mcelreath_and_boyd_2007,shennan_2009_patternprocess,Kempe_and_Mesoudi_2014,mesoudi_cultural_2015}.
 
@@
 Here the probability of copying a cultural variant $i$ is directly proportional to a payoff $s$ exhibited by individuals possessing $i$.
-In other words the social learner assumes that the success of an individual (measured in $s$) is a function of the observed behavioural trait, that is $s=f(i)$.
+In other words the social learner assumes that the success of an individual (measured in $s$) is a function (of which details are often unknown) of the observed behavioural trait, that is $s=f(i)$.
 Several variants of this learning strategies have been proposed and explored, including: \emph{imitate the best}, where the social learner identifies and copies the trait possessed by the individual with the highest payoff; \emph{compare means}, where the social learner compare the average payoff of different strategies and chooses the highest; and the \emph{proportional imitation rule}, where the probability of adopting a variant is proportional to the difference in payoff between the demonstrator and the learner~\citep{schlag1998,baldini2013,crema_lake_inpress}.
@@
 
-The uncertainty in the payoff signal is derived by the presence of interacting cultural and physical traits, as well as the environment where the trait is manifest.
+The uncertainty in the payoff signal is derived by the presence of interacting cultural and physical traits, as well as the environment where the trait is expressed.
 Thus the function $s=f(i)$ should be better described as $s=f(i,\epsilon)$, where $\epsilon$ represent all additional hidden and latent variables that contribute, along with the cultural variant of interest $i$, to the payoff signal $s$.
@@
 In other words, cultural and ecological inheritance should in part reduce the impact played by $\epsilon$, and consequently the uncertainty in the payoff-signal.
-There are however situations where even with most variables hold constant the payoff is subject to high variability.
-Perhaps one of the most relevant examples are
-are contexts where the success of a variant is strongly correlated with the absolute number of individuals $n_i$ possessing the same trait ($i$) within a population.
-This density dependence, which can be described by the payoff function $f(i,n_i)$, implies that the adoption (or abandonment) of a trait can alter its expected payoff-signal, potentially causing a cascade of feedback effects.
+There are however situations the payoff is highly variable even holding these variables constant.
+One of the most relevant examples are showcased in situations where the performance of a given  cultural variant $i$ is partly determined by the number of individuals $n_i$ expressing $i$ within a population.
+This \emph{density dependence}, which can be described by the payoff function $f(i,n_i)$, implies that the adoption (or abandonment) of a trait can alter its expected payoff-signal, potentially causing a cascade of feedback loops.
 
-Several behavioural traits are expected to show such a density dependence.
+Several behavioural traits are expected to show such a pattern.
 For example, behaviour linked to the exploitation of limited resources will exhibit a decline in payoff when density exceeds a critical threshold (i.e carrying capacity).
@@
 
-Density dependence does not, however, necessarily imply only a negative correlation between population size and payoff.
-In many ecological contexts per capita growth rate is positively correlated with population density, thus resulting into an inverse density dependence.
-This phenomena, known as \emph{Allee} effect \citep{allee1958,courchamp_etal_1999}, is generally explained by a range of processes including cooperation, mutual facilitation, and environmental conditioning; all processes that we expect to observe  in many cultural traits.
-The Allee effect has been extensively studied  from  empirical and theoretical standpoints, and its implications assessed for a variety of phenomena, including habitat selection \citep{greene_and_stamps_2001}, resource competition \citep{jang2013},  dispersal \citep{steele_2009}, shifts in settlement pattern \citep{crema_2014}, colonisation events,  \citep{kennet_etal_2006}, genetic diversity \citep{roques_etal_2012} and extinction of languages \citep{sutherland_2003} .
-Yet, despite this wide array of applications, we are unaware of any theoretical studies exploring how the simultaneous presence of direct and inverse density dependence in the payoff signal can affect the evolutionary dynamics of cultural transmission.
+Density dependence does not necessarily imply only a negative correlation between population size and payoff.
+In many ecological contexts per capita growth rate is positively correlated with population density, thus resulting into an inverse density dependence generally linked with processes such as cooperation, mutual facilitation, and environmental conditioning.
+This phenomena, known as \emph{Allee} effect \citep{allee1958,courchamp_etal_1999}, has been extensively studied  from both empirical and theoretical standpoints, and its consequences recognised in a wide variety of contexts, including habitat selection \citep{greene_and_stamps_2001}, resource competition \citep{jang2013},  dispersal \citep{steele_2009}, shifts in settlement pattern \citep{crema_2014}, colonisation events  \citep{kennet_etal_2006}, genetic diversity \citep{roques_etal_2012}, and extinction of languages \citep{sutherland_2003}.
+Yet, despite its ubiquity, we are unaware of any theoretical studies exploring how the simultaneous presence of direct and inverse density dependence in the payoff signal can affect the evolutionary dynamics of cultural transmission.
 
-While the concept of Allee effect is well established in population ecology (see Kramer et al \citeyear{kramer_etal_2009} for a systematic review), there is no concept with direct equivalence in the social sciences.
-That said, there are multiple, established lines of evidence that support  the existence of both direct and inverse density dependence in cultural and behavioural traits.
+Indeed, while the concept of Allee effect is well established in population ecology (see Kramer et al \citeyear{kramer_etal_2009} for a systematic review), there is no direct equivalent in the social sciences.
+Yet, there are multiple, established lines of evidence that support  the existence of both direct and inverse density dependence in human cultural and behavioural traits.
 For example several ethnographic studies on small scale societies show how the per-capita return of hunting groups show higher yield when the group size is not too small nor too big~ \citep{hill_and_hawkes_1983,janssen_and_hill_2014}.
@@
 
-Here we explore a difference equation model of cultural transmission with traits exhibiting an Allee effect.
-We assume that: 1) the social learning strategy is payoff-based; and 2)  the payoff is equivalent to the reproductive fitness of each individual.
+Here we explore the broad consequences of the Allee effect by means of a theoretical equation-based model of cultural transmission.
+For the context of this study we narrow our scope by assuming that: 1) the social learning strategy is payoff-based; and 2)  the payoff is equivalent to the reproductive fitness of each individual.
 We demonstrate that direct and inverse density dependence can strongly affect the evolutionary trajectory of the system, leading to changes in the conditions required to reach specific equilibria, as well as to the emergence of unstable patterns characterised by temporary or permanent episodes of shifts between alternative cultural variants.
@@
 We consider the evolutionary dynamics of two mutually exclusive cultural variants which payoff functions are characterised by an Allee effect.
-We assume a non-fixed population of size $N$, subdivided into two subpopulations with sizes $a$ and $b$, each associated with the two cultural variants.
+We assume a non-fixed population of size $N$, subdivided into two subpopulations with sizes $a$ and $b$, each associated with a different cultural variant.
 We further assume that $a$ and $b$ can change as a result of intrinsic population growth/decline as well as shifts from one to the other as a result of cultural transmission.
@@
 Given that equations \ref{eq1}-\ref{eq4} have no analytical solution that we are aware of, we identified equilibrium conditions via simulation (source codes of the simulation can be found on the following repository \textcolor{red}{PUT LINK HERE})
+%%% We Need to clean the code and make a github repository for this %%%
 
@@
       \includegraphics[width=\textwidth]{./figures/figure5}
-  \caption{Basins of attraction for $z=0.05$ (top) and $z=0.2$ (bottom); $\lambda=1$ (left), $\lambda=0.75$ (middle) and $\lambda=0.5$}
+  \caption{Basins of attraction for $z=0.05$ (top) and $z=0.2$ (bottom); $\lambda=1$ (left), $\lambda=0.75$ (middle) and $\lambda=0.5$ (right)}
     \label{fig:overlap}
@@
 Our analysis examined the impact of direct and inverse density dependence in payoff-based social learning strategies.
-The results showed that the coexistence of both positive and negative feedback mechanism can generate a wide range of patterns where small changes in behaviour and initial conditions can both deeply affect the evolutionary trajectories of the system.
-
-social learning can alter the basins of attraction of stable nodes, leading to different dynamics and equilibria from the very same initial conditions.
+The results showed that Allee effect lead to the emergence of both positive and negative feedback mechanism, and can generate a wide range of patterns where small changes in the initial conditions can both deeply affect the evolutionary trajectories of the system.
+Our theoretical model indicated that social learning can in fact alter the basins of attraction of stable nodes, leading to different dynamics and equilibria from the very same initial conditions.
 If there is no reliance on social learning, the critical density separates the ultimate equilibrium of the system: if the initial condition is below this threshold the variant is destined to go extinct, whilst if it is above this value will eventually reach carrying capacity.
@@
 they share similar critical densities and carrying capacities), the system does not always reach an stable equilibrium, but can instead exhibit cyclical fluctuations between the two variants.
-In other words the consequences of social learning is not absorbed by the system, and instead a negative feedback process leads to continuous switch from one variant to the other.
+In other words the impact of social learning is not absorbed by the system, and instead a negative feedback process leads to continuous switch from one variant to the other.
 
@@
 In many of these examples the initial spread of the innovation generated a positive feedback process, either through enhanced possibility of direct cooperation, or some form of mutual facilitation often indirectly emerging from positive niche construction.
-At the same time, ethnographic and archaeological evidence show how many instances of critical transitions where not smooth, often leading to the coexistence of alternative strategies as well as episodes of a return to a previously abandoned cultural traits.
-Several ethnographic examples showcase small-scale societies alternating between different subsistence modes over a comparatively brief period of time \citep{layton1991,mace_1993,oota_et_al_2005}, while increasing archaeological evidence suggest that the transition from hunting-gathering to farming was not unidirectional, but with instances of temporary reversions \citep{rowley2001,stevens_fuller_2012} %ideally we need a couple of more citations here?
-While in most cases these are anecdotal evidence rather than testable empirical cases, the insights of our model suggests that different levels of reliance on social learning could alone portray some of the evolutionary trajectories observed in these systems.
+At the same time, ethnographic and archaeological evidence show how many instances of critical transitions where not smooth, often leading to the coexistence of alternative strategies as well as episodes of a return to a previously abandoned cultural traits (see Day and Walter \citeyear{day_and_walter_1989} for similar insights in economics).
+Several ethnographic examples showcase small-scale societies alternating between different subsistence modes over a comparatively brief period of time \citep{layton1991,mace_1993,oota_et_al_2005}, while increasing archaeological evidence suggest that the transition from hunting-gathering to farming was not unidirectional, but with instances of temporary reversions \citep{rowley2001,stevens_fuller_2012}.
+Formal cross-cultural analysis of these episodes are still rare (but see Ullah et al.
+\citeyear{ullah_etal_2015}) but the insights of our model suggests that different levels of reliance on social learning could alone portray some of the evolutionary trajectories observed in these systems.
 

### 2015-09-27 manuscript/figures/figure3.R
updated font size

@@ -7,38 +7,38 @@ par(mfrow=c(1,2))
 
 load('../../experimentMN3/00_fig3/result_000.Rdata')
 title=substitute(paste(z == i), list(i=z))
-basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b")
+basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b", cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
 abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
-axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
-axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
+axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8, cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
+axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2, , cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
 trajPlot(ini.m=0.25,ini.n=0.5,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
 trajPlot(ini.m=0.5,ini.n=0.6,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
 trajPlot(ini.m=0.61,ini.n=0.38,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
 trajPlot(ini.m=0.95,ini.n=0.32,Am=Am,Km=Km,An=An,Kn=Kn,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,r.m=r.m,r.n=r.n,z=z)
 
-text(0.26,0.48,labels="i",col=1,font=2)
[…]

### 2015-09-27 manuscript/figures/figure4.R

@@ -1,6 +1,6 @@
-source("./src/alleeIGC.R")
-source("./src/basin.R")
-source("./src/utility.R")
+source("../../src/alleeIGC.R")
+source("../../src/basin.R")
+source("../../src/utility.R")
 
 
 lambda=0.1
@@ -45,5 +45,5 @@ basePlot(timeSteps=200,storeFinalOnly=FALSE,z=0.3,ini.m=0.4,ini.n=0.4,r.n=r.n,r.
 title("z=0.3")
 basePlot(timeSteps=200,storeFinalOnly=FALSE,z=0.6,ini.m=0.4,ini.n=0.4,r.n=r.n,r.m=r.m,Kn=Kn,Km=Km,An=An,Am=Am,cAn=cAn,cAm=cAm,cKn=cKn,cKm=cKm,leg=FALSE,ymax=0.9)
 title("z=0.6")
-dev.print(device=pdf,"~/github/CultureAllee/manuscript/figures/figure4b.pdf")
+dev.print(device=pdf,"figure4b.pdf")
 

### 2015-09-27 manuscript/figures/figure5.R

@@ -7,51 +7,51 @@ pdf('figure5.pdf', width=15, height=10)
 par(mfrow=c(2,3))
 load('../../experimentMN3/01_lambda/result_002.Rdata')
 title=substitute(paste(z == i, "  ", lambda == j), list(i=z,j=lambda))
-basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b")
+basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b", cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
 abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
-axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
-axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
+axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8, cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
+axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2, cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
 
 load('../../experimentMN3/01_lambda/result_001.Rdata')
 title=substitute(paste(z == i, "  ", lambda == j), list(i=z,j=lambda))
-basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b")
+basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b", cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
 abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
-axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
[…]

### 2015-09-27 manuscript/figures/figure7.R

@@ -7,51 +7,51 @@ pdf('figure7.pdf', width=12, height=9)
 par(mfrow=c(2,3))
 load('../../experimentMN3/03_competition/result_000.Rdata')
 title=substitute(paste(z == i, " ", c== j), list(i=z,j=competition))
-basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b")
+basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b", cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
 abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
-axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
-axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
+axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8, cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
+axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2, cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
 
 load('../../experimentMN3/03_competition/result_002.Rdata')
 title=substitute(paste(z == i, " ", c== j), list(i=z,j=competition))
-basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b")
+basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b", cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
 abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
-axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8)
-axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2)
[…]

### 2015-09-27 src/basin.R

@@ -94,12 +94,12 @@ basePlot<-function(leg=TRUE,ymax=NA,...)
     {
         tmp=allee(...)
         if(is.na(ymax)){ymax=max(unlist(tmp))}
-        plot(tmp$m,type="l",ylim=c(0,ymax),col="indianred",lwd=2,ylab="Population Size",xlab="Time")
-        axis(side=4,at=c(Km,Kn),labels=c(expression(K[a]),expression(K[b])),hadj=0.5,las=2)
+        plot(tmp$m,type="l",ylim=c(0,ymax),col="indianred",lwd=2,ylab="Population Size",xlab="Time", cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)
+        axis(side=4,at=c(Km,Kn),labels=c(expression(K[a]),expression(K[b])),hadj=0.5,las=2, cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)
         abline(h=c(Km,Kn),lty=2)
         lines(tmp$n,col="royalblue",lwd=2)
         if(leg==TRUE)
-        {legend("bottomright",legend=c("a","b"),lty=1,lwd=2,col=c("indianred","royalblue"))}
+        {legend("bottomright",legend=c("a","b"),lty=1,lwd=2,col=c("indianred","royalblue"), cex=1.5)}
     }
 
 

### 2015-09-27 manuscript/figures/figure7.R
fixed error

@@ -50,8 +50,8 @@ title=substitute(paste(z == i, " ", c== j), list(i=z,j=competition))
 basinPlot(res=result,Am=Am,Km=Km,An=An,Kn=Kn,cKn=cKn,cKm=cKm,main=title,xlab="a",ylab="b", cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
 abline(h=c(An,Kn),lty=2,col="white")
 abline(v=c(Am,Km),lty=2,col="white")
-axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])),padj=0.8), cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2
-axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])),hadj=0.5,las=2, cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
+axis(side=3,at=c(Am,Km),labels=c(expression(A[a]),expression(K[a])), padj=0.8, cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
+axis(side=4,at=c(An,Kn),labels=c(expression(A[b]),expression(K[b])), hadj=0.5,las=2, cex.lab=2, cex.axis=2, cex.main=2, cex.sub=2)
 
 dev.off()
 

### 2015-09-29 manuscript/draft.tex
Reference to ESM
Reference ESM for the simulation code example

@@
 that ensures that the two variants have always the same $R_{max}$.
-Given that equations \ref{eq1}-\ref{eq4} have no analytical solution that we are aware of, we identified equilibrium conditions via simulation (source codes of the simulation can be found on the following repository \textcolor{red}{PUT LINK HERE})
-%%% We Need to clean the code and make a github repository for this %%%
+Given that equations \ref{eq1}-\ref{eq4} have no analytical solution that we are aware of, we identified equilibrium conditions via simulation (examples codes are provided on the ESM).
 

### 2015-10-01 manuscript/draft.tex
typo fixed

@@
 
-\newcommand{\memo}[2]{\textcolor{#1}{#2}}
-\newcommand{\xavi}[1]{\memo{orange}{xavi: #1\\}}
-\newcommand{\enrico}[1]{\memo{blue}{enrico: #1\\}}
+\usepackage{lineno}
+    \linenumbers
 
@@
 \begin{keyword}
-Social Learning\sep Subsistence\sep Allee Effect\sep Cooperation\sep Carrying Capacity
+Social Learning\sep Cultural Evolution\sep Allee Effect\sep Cooperation\sep Carrying Capacity
 \end{keyword}
@@
 Early works inspired by population biology \citep{cavallisforza_feldman_1981,boyd1985} have laid the foundation of a now mature trans-disciplinary field which unites the social and biological sciences \citep{mesoudi_etal_2006, mcelreath_and_boyd_2007}.
-The rich cross-fertilisation between different subfields has promoted the designing of new analytical techniques as well as new uses for existing methods, but more crucially have prompted the development of theoretical models of how culture is acquired and transmitted between individuals~\citep{henrich_mcelreath2003,laland2004,mcelreath_and_boyd_2007,shennan_2009_patternprocess,Kempe_and_Mesoudi_2014,mesoudi_cultural_2015}.
+The rich cross-fertilisation between different subfields has encouraged the designing of new analytical techniques as well as new uses for existing methods, but more crucially have prompted the development of theoretical models of how culture is acquired and transmitted between individuals~\citep{henrich_mcelreath2003,laland2004,mcelreath_and_boyd_2007,shennan_2009_patternprocess,Kempe_and_Mesoudi_2014,mesoudi_cultural_2015}.
 
 At the core of most models of social learning there is some heuristics that promotes the selection of a given cultural variant over another \citep{laland2004}.
-This can be based on some properties of the variant itself (\emph{content bias}) or derived from contextual cues (\emph{context bias}), such as the commonness or the rareness of a trait (\emph{conformist bias} and \emph{anti-conformist bias}), or  features associated with the transmitter such as prestige, success, or similarity to the learner (\emph{context bias})\citep{henrich_mcelreath2003}.
+This can be based on some properties of the variant itself (\emph{content bias}) or derived from contextual cues (\emph{context bias}), such as the commonness or the rareness of a trait (\emph{conformist bias} and \emph{anti-conformist bias}), or  features associated with the transmitter such as prestige, success, or similarity to the learner (\emph{context bias}; see Henrich and McElreath \citeyear{henrich_mcelreath2003} for a review of these heuristics).
 
@@
 Whilst most studies focus on how difference in these imitation rules can drive the evolutionary dynamics of the cultural system, less attention has been given to the relationship between target traits and their payoff signals.
-Typical signals are most likely some indirect proxy of success (e.g.
+Typically these are most likely some proxy of success (e.g.
 number of offspring, income, presence/absence of other traits associated with success or prestige, etc.)  that is assumed (by the social learner) to be correlated with intrinsic properties of the cultural variant.
@@
 Density dependence does not necessarily imply only a negative correlation between population size and payoff.
-In many ecological contexts per capita growth rate is positively correlated with population density, thus resulting into an inverse density dependence generally linked with processes such as cooperation, mutual facilitation, and environmental conditioning.
+In many ecological contexts per-capita growth rate is positively correlated with population density, thus resulting into an inverse density dependence generally linked with processes such as cooperation, mutual facilitation, and environmental conditioning.
 This phenomena, known as \emph{Allee} effect \citep{allee1958,courchamp_etal_1999}, has been extensively studied  from both empirical and theoretical standpoints, and its consequences recognised in a wide variety of contexts, including habitat selection \citep{greene_and_stamps_2001}, resource competition \citep{jang2013},  dispersal \citep{steele_2009}, shifts in settlement pattern \citep{crema_2014}, colonisation events  \citep{kennet_etal_2006}, genetic diversity \citep{roques_etal_2012}, and extinction of languages \citep{sutherland_2003}.
@@
 small values of $m$ determines a higher rate by which the theoretical maximum transmission rate is reached), thus portraying also potential acquisition costs of the new variant (c.f.
-"threshold of evidence', Hernich, \citeyear{henrich2001cultural}: 994)
+"threshold of evidence', Henrich, \citeyear{henrich2001cultural}: 994)
 
@@
 
-When $z=0$, equations \ref{eq1}-\ref{eq4} becomes a standard cubic growth model with Allee effect, where the success (or failure) of a given focal trait is exclusively given by its own density.
+When $z=0$, equations \ref{eq1}-\ref{eq4} becomes a standard cubic growth model with Allee effect, where the success (or failure) of a given focal trait is primarily given by its own density.
 For a generic cultural trait $x$, the model predicts three equilibrium points at $x=0$ (extinction), $x=A_x$, and $x=K_x$.
-The equilibrium $x=A_x$ is unstable for all values of $r_x$, whereas the stability of the node at $x=K_x$ is a function of the per capita growth rate $r_x$.
+The equilibrium $x=A_x$ is unstable for all values of $r_x$, whereas the stability of the node at $x=K_x$ is a function of the per-capita growth rate $r_x$.
 Increasing the latter decreases the stability of fixed points~\citep{scheuring_1999}, leading to the transition from point attractors (with equilibrium at $x=K_x$), to limit cycles and chaotic dynamics with islands of stability (see fig.
@@
 
-Second, when the magnitude of social learning is high, the system is subject to effects similar to those observed with an increase in the per capita growth rate $r$ (cf.
+Second, when the magnitude of social learning is high, the system is subject to effects similar to those observed with an increase in the per-capita growth rate $r$ (cf.
 fig.
@@
 Increasing values of $z$ shift the system from the co-existence equilibrium towards a limit cycle with abrupt population changes.
-Model parameters: $A_a=0.1$; $A_b=0.2$; $K_a=0.6$; $K_b=0.7$; $r_a=0.05$}
+Model parameters: $A_a=0.1$; $A_b=0.2$; $K_a=0.6$; $K_b=0.7$; $r_a=r_b=0.05$}
     \label{fig:TSsocialLearning}
@@
 Here we briefly explore the latter scenario, by relaxing the assumption of independence in the computation of the payoff of each variant.
-For instance several subsistence strategies do not have fully independent resource pools, and hence the success of a given variant might be detrimental to the payoff of the other irregardless of the effects provoked by cultural transmission.
+For instance many subsistence strategies do not have fully independent resource pools, and hence the success of a given variant might be detrimental to the payoff of the other irregardless of the effects provoked by cultural transmission.
 We can revise our core model by including new terms portraying the competition effect between the two resources:
@@
 Our analysis examined the impact of direct and inverse density dependence in payoff-based social learning strategies.
-The results showed that Allee effect lead to the emergence of both positive and negative feedback mechanism, and can generate a wide range of patterns where small changes in the initial conditions can both deeply affect the evolutionary trajectories of the system.
+The results showed that Allee effect lead to the emergence of both positive and negative feedback mechanisms, and can generate a wide range of patterns where small changes in the initial conditions can deeply affect the evolutionary trajectories of the system.
 Our theoretical model indicated that social learning can in fact alter the basins of attraction of stable nodes, leading to different dynamics and equilibria from the very same initial conditions.
 If there is no reliance on social learning, the critical density separates the ultimate equilibrium of the system: if the initial condition is below this threshold the variant is destined to go extinct, whilst if it is above this value will eventually reach carrying capacity.
-When social learning is present this boundary shifts, variants can survive even if its initial condition is below critical density and equally can reach extinction even if above this threshold.
-This scenario is further enhanced when the two variants are not independent but are in direct competition.
+When social learning is present this boundary shifts; variants can survive even if its initial condition is below critical density and equally can reach extinction even if above this threshold.
+This scenario is also observed when the two variants are not independent but are in direct competition.
 In this case, social learning can further enhance the advantage of one variant over the other.
@@
 When the reliance on social learning is comparatively high, and the two variants have a substantial overlap in their range of positive payoff densities (i.e.
-they share similar critical densities and carrying capacities), the system does not always reach an stable equilibrium, but can instead exhibit cyclical fluctuations between the two variants.
+they share similar critical densities and carrying capacities), the system does not always reach an stable equilibrium, but can instead exhibit cyclical fluctuations.
 In other words the impact of social learning is not absorbed by the system, and instead a negative feedback process leads to continuous switch from one variant to the other.
@@
 Perhaps the most notable example is the spread of keystone innovations in subsistence strategies which led to the adoption of variants enabling larger carrying capacity, often at the cost of higher critical densities.
-The spread of behavioural traits such as the know-how of the domestication process~\citep{barker2006} the use of new subsistence tools~\citep{petraglia_population_2009} and processing techniques~\citep{molleson1993}, as well as the inclusion/exclusion and extensification/intensification of different prey species are just some of the most notable examples.
+The spread of behavioural traits such as the know-how of the domestication process~\citep{barker2006}, the use of new subsistence tools~\citep{petraglia_population_2009} and processing techniques~\citep{molleson1993}, as well as the inclusion/exclusion and extensification/intensification of different prey species are just some of the most notable examples.
 In many of these examples the initial spread of the innovation generated a positive feedback process, either through enhanced possibility of direct cooperation, or some form of mutual facilitation often indirectly emerging from positive niche construction.
-At the same time, ethnographic and archaeological evidence show how many instances of critical transitions where not smooth, often leading to the coexistence of alternative strategies as well as episodes of a return to a previously abandoned cultural traits (see Day and Walter \citeyear{day_and_walter_1989} for similar insights in economics).
-Several ethnographic examples showcase small-scale societies alternating between different subsistence modes over a comparatively brief period of time \citep{layton1991,mace_1993,oota_et_al_2005}, while increasing archaeological evidence suggest that the transition from hunting-gathering to farming was not unidirectional, but with instances of temporary reversions \citep{rowley2001,stevens_fuller_2012}.
+At the same time, ethnographic and archaeological evidence show how many instances of critical transitions where not smooth, often leading to the coexistence of alternative strategies as well as episodes of a return to previously abandoned cultural traits (see Day and Walter \citeyear{day_and_walter_1989} for similar insights in economics).
+Several ethnographic examples showcase small-scale societies alternating between different subsistence modes over a comparatively brief period of time \citep{layton1991,mace_1993,oota_et_al_2005}, while increasing archaeological evidence suggests that the transition from hunting-gathering to farming was not unidirectional, but with instances of temporary reversions \citep{rowley2001,stevens_fuller_2012}.
 Formal cross-cultural analysis of these episodes are still rare (but see Ullah et al.

### 2015-10-01 manuscript/draft.tex
title

@@
 
-\title{Allee Effect and Cultural Evolution}
+\title{Allee Effect in Cultural Evolution}
 
@@
 
-\address[label1]{CaSEs - Complexity and Socio-Ecological Dynamics Research Group, Barcelona}
+\address[label1]{CaSEs - Complexity and Socio-Ecological Dynamics Research Group, Universitat Pompeu Fabra, Barcelona}
 \address[label2]{UCL Institute of Archaeology}
-\address[label3]{BSC - Barcelona Supercomputing Center}
+\address[label3]{CASE - Computer Applications in Science and Engineering, Barcelona Supercomputing Center}
 

### 2015-10-01 manuscript/draft.tex
acknowledgements

@@
 
+\section{Acknowledgements}
+
+ERC was funded by a Marie-Curie Intra-European Fellowship (grant 625863).
+XRC is part of the SimulPast Project (CSD2010-00034) funded by the CONSOLIDER-INGENIO2010 program of the Ministry of Science and Innovation—Spain.
+Experiments were run using the MareNostrum 3 supercomputer at the Barcelona Supercomputing Centre facility.
+
 \section{References}
